### 2020-09-08 docs/resources/CogSci_paper/paper.tex
last update

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
% Modified : Roger Levy (rplevy@mit.edu)     12/31/2018

%% Change "letterpaper" in the following line to "a4paper" if you must.

\documentclass[10pt,letterpaper]{article}
\usepackage{cogsci}
\usepackage{dsfont}
\usepackage{bm}
\usepackage{amssymb}

%\cogscifinalcopy % Uncomment this line for the final submission

\usepackage{pslatex}
\usepackage{apacite}
\usepackage{float} % Roger Levy added this and changed figure/table
                   % placement to [H] for conformity to Word template,
                   % though floating tables and figures to top is
                   % still generally recommended!

%\usepackage[none]{hyphenat} % Sometimes it can be useful to turn off
%hyphenation for purposes such as spell checking of the resulting
%PDF.  Uncomment this block to turn off hyphenation.

%\setlength\titlebox{4.5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 4.5cm (the original size).
%%If you do, we reserve the right to require you to change it back in
%%the camera-ready version, which could interfere with the timely
%%appearance of your paper in the Proceedings.

\title{How to Make a Proceedings Paper Submission}

\author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\
  Department of Psychology, 1202 W. Johnson Street \\
  Madison, WI 53706 USA
  \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\
  Department of Educational Psychology, 1025 W. Johnson Street \\
  Madison, WI 53706 USA}

\begin{document}

\maketitle

\begin{abstract}

\textbf{Keywords:} 
\end{abstract}

\section{Introduction}

Identifying the causes of an event is a very important ability that people largely rely on in common life. Even if reasonning about causes doesn't generally pose problems and comes up with clear answers about what caused an event, it is sometimes much more difficult to disambiguous an event and answer such a question. When divergent causal intuitions are involved it can be hard to decide for one candidate cause rather than an other. This can lead to very different and serious implications, especially legal ones. For example a well-known story in the literature is the "Irish case", a real-world example of a court case: [include here Irish case]. In such legal contexts determining the main causes of an event is an extremely important task that involves finding who or what is at fault in dramatic accidents, property damage, etc. and can pose a veritable challenge. Here massive fines, jail sentences or death penalties can be the outcomes of the inferred causal history of an event. Reasonning about causes, however, is often implied in much more naive and common circonstances and doesn't necessary pose any difficulty --e.g. identifying what is the cause of the bottle shattering. But even in very simple situations like that, it is sometimes not so easy to theoretically explain and predict intuitive responses that people give. [Include here Susy vs Bob example]. In all these common-life situations we want to know what the \textit{actual cause} of an event is, in a specific context. That is we don't want here to infer some general causal laws responsible for the occurrence of an event but understand what is the main or most intuive singular causes in particular cases. The current paper focuses on causal judgements of the latter type, that is on the concept of \textit{actual} and not \textit{general} causation\footnote{also called "token" or "singular" causation}, following a distinction that has been frequently made in the litterature. Causal judgements have been widely studied and there are existing models of actual causation that aim to capture people's inferences about the causes of a single event. Current research on actual causation mainly relies on two different interpretations of causality : the counterfactual account (CF) or the physical process one (PP).

According to the first interpretation, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B. In other words if B would have been different without A, then A is an actual cause. More precisely A is a cause of B if and only if A and B are true, and A hadn't occurred B wouldn't have occurred. [Example of Suzy only] This type of explanation calls upon counterfactual scenario or possible worlds where the presumed cause of the effect is removed from the system while all other relevant factors are kept as unchanged as possible compared to the actual world. However, a first well-known challenge that the CF account encounters is that it is not clear which counterfactual world we have to take into account in order to justify a singular cause, and how to accurately compare those counterfactuals. Again let's say that Suzy throws a rock at a bottle but this time Bob as well throws a rock at the same bottle with the same precision as Suzy. Suzy's rock reaches the bottle just a second before Bob's and breaks it. According to the CF account if Suzy hadn't thrown the rock, the bottle would have still shattered since Bob's rock would have hit it regardless. So following the general idea mentionned above, Suzy's throw cannot be the cause of the bottle shattering -- which is counter-intuitive. This is an example of a general type of cases we call \textit{late preemption} and refers to the fact that the final event -- i.e. the effect -- is "overdetermined" [Hall]. A first common attempt to face the problem is to say that the bottle-shattering event that results from Suzy's throw is qualitatively different from the the bottle-shattering event that follows from Bob's throw. In other words A is a cause of B if and only if A and B occur, and A hadn't occurred B wouldn't have occurred at all or would have been delayed. Thus Suzy's throw is indeed the cause of the bottle-shattering event that occurs at the time it did.

[Halpern and Pearl solution : graphical models, static, pb of modelling, without time, etc.] 
[Then, more general problem with counterfactual dependance].

The physical process account of causation gives a completely different interpretation of causality and states that an event A causes B when there is a physical connection between them. Several theories have been proposed to caracterise this physical connection. In one of its latest and probably most convincing formulation, we have to distinguish two things : causal process and causation or causal interaction. A causal process is a physical process involving an object which conserves a certain quantity accross space and time. The conserved quantity is typically a measurable physical property of the object like the mass-energy, etc. and is precisely what scientific theories are meant to formalize. In that sence both Suzy's and Bob's throw are causal processes. A causal interaction or causation is an exchange of that conserved quantity, for example the mass-energy of the rock to the glass of the bottle. According to this definition, only Suzy's throw involves a causal interaction with the bottle. However the PP account of causation doesn't consider as genuine causation some very common cases that people intuitively judge as proper example of causation. For example if I hold the head of my ennemy under water and make him die, I'm not genuinely the cause of his death; rather I'm actually preventing the possibility of a genuine causation which is the physical process of breathing oxygen in order to live. The main problem is that, as a consequence, the PP account not only goes sometimes against some of our best causal intuitions, but also needs the counterfactual analysis as well to ground cases of \textit{quasi-causation} by omission or prevention like the latter one.

[?? A third account : probability raising ??]

\section{Theoretical proposition}

As a result of the preceeding analysis it seems that none of the current main theories of actual causation gives a satisfying explanation of the causal judgements that people actually do. This paper aims at solving the problem by interpreting causality in a completely different way. In the CF framework events are represented as mere propostions and causation is thought as a relation between static states. The PP account is meant to give an epistemological definition of the concept of causation and not an explanation of how people actually rely on temporal informations to make causal judgements. By contrast this paper focuses on the role of temporal information for inferring causal relationship between events qua changes of states over time. In other words we hypothesised that people's causal judgements are influenced by not only the values of all relevant variables at the time the effect occurs, but also the temporal order in which these variables took their values. [?? Add a concrete example ??] This approach has some precedent in the literature but has not been developed carefully and has never been tested experimentally as we intend to do. We add further that people mainly identify causation with a \textit{continuous} sequence of changes of states over time. The underlying intuition is that people trace back the history of changes from the immediate one that directly brought about the occurrence of the effect, up to the root change in the system that initiated the series of changes along the path. As we think that in common life it is really rare to see simultaneously two or more events occurring at the very same time and producing some common effect, we also want to postulate a \textit{no coincidence principle}. Relying on this principle we can split the system's time frame in units of time that are small enough to have no more than one event\footnote{Again we insist on the definition of \textit{event} as \textit{change of state}.} per unit. Lets represent by $\mu^{\emptyset}_{z_{i\rightarrow j}}(t)$ a change of state at time $t$, from variable $Z=z_i$ to $Z=z_j$, $\forall i,j\in \mathds{R}_+$. Here $Z$ is the variable we want to reason about and has no children (represented by the empty set $\{\emptyset\}$). Formalizing our above mentioned idea we first have to find $\mathcal{U}^Z_{\bm{y}_{i\rightarrow j}}(t-1)$, that is the set of the immediate changes in the parents $\bm{Y}$ of the variable $Z$ at time $t-1$ that led to the observed change in $Z$ at time $t$. So we want to find $\mathcal{U}$ such that $\mathcal{U}^Z_{\bm{y}_{i\rightarrow j}}(t-1)\rightarrow \mu^{\emptyset}_{z_{i\rightarrow j}}(t)$. According to our \textit{no coincidence principle} $\mathcal{U}$ is either empty or a singleton -- including only one parent whose value changed. Then we want to find the set of changes such that $\mathcal{U}^Y_{\bm{x}_{i\rightarrow j}}(t-2)\rightarrow \mu^{Z}_{y_{i\rightarrow j}}(t-1)$, that is the set of immediate previous changes in the parents $\bm{X}$ of the variable $Y$ at time $t-2$ that led to the observed change in $Y$ at time $t-1$. Following the same logic we suggest that if $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)\rightarrow \mu^{Y}_{x_{i\rightarrow j}}(t-2)$ is such that $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)=\emptyset$, then it means that $\mu^{Y}_{x_{i\rightarrow j}}(t-2)$ represents the chronologically first change along the path, occurring in $X$, and we suggest that this change is identified as being the main cause of $Z$.

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{CogSci_Template}

\end{document}

### 2020-09-09 docs/resources/CogSci_paper/paper.tex
paper modified

@@
 \usepackage{amssymb}
+\usepackage{indentfirst}
 
@@
 According to our \textit{no coincidence principle} $\mathcal{U}$ is either empty or a singleton -- including only one parent whose value changed.
+Lets say that at $t-1$ we find a change in a parent variable $Y$, so $\mathcal{U}^Z_{\bm{y}_{i\rightarrow j}}(t-1)=\{\mu^{Z}_{y_{i\rightarrow j}}(t-1)\}$.
 Then we want to find the set of changes such that $\mathcal{U}^Y_{\bm{x}_{i\rightarrow j}}(t-2)\rightarrow \mu^{Z}_{y_{i\rightarrow j}}(t-1)$, that is the set of immediate previous changes in the parents $\bm{X}$ of the variable $Y$ at time $t-2$ that led to the observed change in $Y$ at time $t-1$.
 Following the same logic we suggest that if $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)\rightarrow \mu^{Y}_{x_{i\rightarrow j}}(t-2)$ is such that $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)=\emptyset$, then it means that $\mu^{Y}_{x_{i\rightarrow j}}(t-2)$ represents the chronologically first change along the path, occurring in $X$, and we suggest that this change is identified as being the main cause of $Z$.
+
+\section{Experiments}
+
+To test our hypothesis that the main cause of an effect is the root change that initiates a continuous sequences of changes until the occurrence of the effect, participants were presented animations showing activation spreading over networks of nodes up to the final node.
+We run three different experiments which shared the same plot that we wanted as intuitive as possible: participants were told that they were working in a nuclear control room and that their job was to monitor networks of particule detectors.
+It the experiments when a detector, depicted by a square or circle (see below), absorbs a radioactive particule it becomes active and turns black, transmitting the activation across the links of the network so that an active detector activates the next one in the chain and so forth.
+All the networks included a special component, called 'Gauge of Critical Moment', that becomes active only if all of its input from the detectors it is connected to are active.
+The Gauge of Critical Moment was always represented by a square with "GCM" or "G" above.
+At the end of an activation sequence, participants had to click on the detector(s) they considered as the main cause(s) of the activation of the Gauge of Critical Moment.
+
+[Add some stuff... For example why choosing this setting, the underlying idea to make something related to physical intuitions, no high order cognition etc. ??]
+
+\subsection{Experiment 1}
+
+\textbf{Participants}. Paragraph...
+
+\textbf{Method}. Paragraph...
+
+\textbf{Results}. Paragraph...
+
+\textbf{Discussion}. Paragraph...
+
+\subsection{Experiment 2}
+
+\textbf{Participants}. Paragraph...
+
+\textbf{Method}. Paragraph...
+
+\textbf{Results}. Paragraph...
+
+\textbf{Discussion}. Paragraph...
+
+\subsection{Experiment 3}
+
+\textbf{Participants}. Paragraph...
+
+\textbf{Method}. Paragraph...
+
+\textbf{Results}. Paragraph...
+
+\textbf{Discussion}. Paragraph...
+
+\section{General discussion}
 

### 2020-09-23 docs/resources/CogSci_paper/paper (copy).tex
last update

--- 2020-09-08 docs/resources/CogSci_paper/paper.tex
@@
 \usepackage{amssymb}
+\usepackage{indentfirst}
+\usepackage{graphicx}
+\usepackage{subfig}
+\usepackage{multirow}
 
@@
 This can lead to very different and serious implications, especially legal ones.
-For example a well-known story in the literature is the "Irish case", a real-world example of a court case: [include here Irish case].
+For example a well-known story in the literature is the ``Irish case'', a real-world example of a court case: [include here Irish case].
 In such legal contexts determining the main causes of an event is an extremely important task that involves finding who or what is at fault in dramatic accidents, property damage, etc.
@@
 That is we don't want here to infer some general causal laws responsible for the occurrence of an event but understand what is the main or most intuive singular causes in particular cases.
-The current paper focuses on causal judgements of the latter type, that is on the concept of \textit{actual} and not \textit{general} causation\footnote{also called "token" or "singular" causation}, following a distinction that has been frequently made in the litterature.
+The current paper focuses on causal judgements of the latter type, that is on the concept of \textit{actual} and not \textit{general} causation\footnote{also called ``token'' or ``singular'' causation}, following a distinction that has been frequently made in the litterature.
 Causal judgements have been widely studied and there are existing models of actual causation that aim to capture people's inferences about the causes of a single event.
 Current research on actual causation mainly relies on two different interpretations of causality : the counterfactual account (CF) or the physical process one (PP).
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=6cm]{intro}
+\end{center}
+\caption{This is a figure.} 
+\label{fig:1}
+\end{figure}
 
@@
 This is an example of a general type of cases we call \textit{late preemption} and refers to the fact that the final event -- i.e.
-the effect -- is "overdetermined" [Hall].
+the effect -- is ``overdetermined'' [Hall].
 A first common attempt to face the problem is to say that the bottle-shattering event that results from Suzy's throw is qualitatively different from the the bottle-shattering event that follows from Bob's throw.
@@
 According to our \textit{no coincidence principle} $\mathcal{U}$ is either empty or a singleton -- including only one parent whose value changed.
+Lets say that at $t-1$ we find a change in a parent variable $Y$, so $\mathcal{U}^Z_{\bm{y}_{i\rightarrow j}}(t-1)=\{\mu^{Z}_{y_{i\rightarrow j}}(t-1)\}$.
 Then we want to find the set of changes such that $\mathcal{U}^Y_{\bm{x}_{i\rightarrow j}}(t-2)\rightarrow \mu^{Z}_{y_{i\rightarrow j}}(t-1)$, that is the set of immediate previous changes in the parents $\bm{X}$ of the variable $Y$ at time $t-2$ that led to the observed change in $Y$ at time $t-1$.
 Following the same logic we suggest that if $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)\rightarrow \mu^{Y}_{x_{i\rightarrow j}}(t-2)$ is such that $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)=\emptyset$, then it means that $\mu^{Y}_{x_{i\rightarrow j}}(t-2)$ represents the chronologically first change along the path, occurring in $X$, and we suggest that this change is identified as being the main cause of $Z$.
+
+\section{Experiments}
+
+To test our hypothesis that the main cause of an effect is the root change that initiates a continuous sequences of changes until the occurrence of the effect, participants were presented animations showing activation spreading over networks of nodes up to the final node.
+We run three different experiments which shared the same plot that we wanted as intuitive as possible: participants were told that they were working in a nuclear control room and that their job was to monitor networks of particule detectors.
+The instructions said that when a detector, depicted by a square or circle (see below), absorbs a radioactive particule it becomes active and turns black, transmitting the activation across the links of the network so that an active detector activates the next one in the chain and so forth.
+All the networks include a special component, called 'Gauge of Critical Moment', that becomes active only if all of its input from the detectors it is connected to are active.
+The Gauge of Critical Moment is always represented by a square with ``GCM'' (Experiment 1) or ``G'' (Experiment 2) above.
+At the end of an activation sequence, it is asked to click on the detector(s) they considered as the main cause(s) of the activation of the Gauge of Critical Moment.
+Examples of network activations were shown in the instructions and participants had to answer to answer a survey at the end to see if they correctly understood the instructions.
+If they made a mistake they had to read again all the instructions and answer again the survey.
+
+In both experiments we included only two types of input-output logic, namely single input (chains) and AND-Gate (branches where the effect needs the activation of two inputs to occur).
+When people were presented chains, we first wanted to see if they identified the main cause of the effect with a change of state that occurred in the chain the furthest away from effect (\textit{root change}) or the closest to it (\textit{immediate change}).
+When they were presented AND-Gates, we wanted to see if they identified the main cause of the effect with the root change that occurred chronologically first (\textit{1\textsuperscript{st} root change}, labeled ``R1'' henceforth), the root change that occurred chronologically last (\textit{2\textsuperscript{nd} root change}, ``R2'' henceforth), the immediate change that occurred chronologically first (\textit{1\textsuperscript{st} immediate change}, ``I1'') or the immediate change that occurred chronologically last (\textit{2\textsuperscript{nd} immediate change}, ``I1'').
+
+\subsection{Experiment 1}
+
+The objective of this experiment was to see whether people's causal judgement were influenced by: the length of the sequences of changes (for both chains and AND-Gates), and the delay between the 1\textsuperscript{st} immediate change and the 2\textsuperscript{nd} root change, that is between the end of the first sequence of changes and the beginning of the second sequence of changes (for AND-Gates only).
+
+\textbf{Participants}.
+The experiment was hosted by Google Cloud App Engine and 30 participants were recruited via Amazon Mahchanical Turk.
+There were 10 females (average age: 37.3) and 20 males (average age: 37.6).
+28 participants were english speakers, 1 were italian speaker and 1 marathi speaker.
+
+\textbf{Method}.
+Each participant was presented 15 different networks of square detectors.
+Square detectors maintain their activation through time.
+Participants were first presented the network in its initial and static state and were asked to click the ``Run'' button to observe an activation sequence over the network.
+They waited 5000ms to see a first change of state occurring in a detector and the activation delay between any two successive detectors was set on 100ms (see Fig.\ref{fig:1}).
+
+Three stimuli were chains that differed in length of activation sequence, that is including 2, 4 or 7 activated squares (with the GCM).
+These lengths were labeled ``Short'', ``Medium'' and ``Long''.
+The AND-Gates were similarily categorised ``Short'', ``Medium'' (like in Fig.\ref{fig:1}) and ``Long'' with both branches being the same length.
+In each length category participants were shown three similar AND-Gates that differed uniquely in the delays between the 1\textsuperscript{st} immediate change and the 2\textsuperscript{nd} root change.
+The delays were 2000ms, 4000ms and 6000ms.
+
+All the stimuli were presented in a random order.
+Two groups, ``Left'' and ``Right'', were designed and partcipants were randomly assigned to either group at the begining of the experiment.
+In the ``Left'' group all the networks were shifted towards left (the GCM being on the left) and in the ``Right'' group all the networks were shifted towards right.
+For each AND-Gate participants saw randomly either the top branch or the bottom branch activating first.
+
+After the activation of the GCM (the end of the animation) participants had to wait 1000ms before the squares became clickable and the following intruction appeared: ``In this sequence what caused the activation of the GCM?
+Respond by clicking on a detector''.
+They had the option to run again the animation (no more than 9 times) or to go to the next network assuming they clicked on one detector -- they couldn't select more than one detector.
+When selected the edges of a detector turned red.
+
+\textbf{Results}. Paragraph...
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=8cm]{results_E1}
+\end{center}
+\caption{This is a figure.} 
+\label{fig:2}
+\end{figure}
+
+\textbf{Discussion}. Paragraph...
+
+\subsection{Experiment 2}
+
+\textbf{Participants}.
+The experiment was hosted by Google App Engine and 89 participants were recruited via Amazon Mechanical Turk.
+There were 29 females (average age: 39.7), 59 males (average age: 37.6) and 1 declared ``other''.
+All the participants were english speakers.
+
+\textbf{Method}.
+Like in Experiment 1 each participant was presented 15 different networks.
+However this time we introduced round detectors that, contrarily to square detectors, don't maintain their activation through time.
+By adding this feature we were able to include networks that have a loop of detectors (see Fig.\ref{fig:3b}).
+1 stimulus was a chain of square detectors and 1 stimulus was a chain of round detectors were included (for the latter one, only ``G'' remained a square).
+The other 13 stimuli were AND-Gates and were sorted into three main categories: 4 stimuli were Rolled networks, i.e.
+with a loop, as depicted in Fig.\ref{fig:3b}, 3 were Unrolled networks with square and round detectors as in Fig.\ref{fig:3a}, labeled ``Unrolled circles'' and 6 stimuli were Unrolled with square detectors in place of the round one (not represented here), labeled ``Unrolled squares''.
+For each category of AND-Gates all the stimuli were sorted according to two crossed dimensions: ``Event'' \textit{vs} ``State''and ``R1 cont'' \textit{vs} ``R2 cont''.
+In condition ``Event'' none of the detectors is initially active or -- if the branch is a loop -- is continuously (re-)activated.
+In condition ``State'' one of the two branches contains initially detectors that are active or is continuously (re-)activated.
+In condition ``R1 cont'' (represented in Fig.\ref{fig:3a}) the activation is such that there is a continuous sequence of changes going from R1 to I2 that leads to the activation of ``G''.
+In condition ``R2 cont'' (not represented here) this is R2 that is continously connected to I2 that leads to the activation of ``G''.
+The design of the experiment for the AND-Gates is given in Tab.\ref{tab:1}.
+For the AND-Gates (Rolled and Unrolled) the delay between I1 and I2 in condition ``Event'' was maintained at 11 nodes (i.e.1100 ms) for almost all the networks.
+There was only one Unrolled squares network for which the delay was set at 8 nodes (800ms).
+
+\begin{table}[H]
+\begin{center} 
+\vskip 0.12in
+\begin{tabular}{c|c|c|c|c} 
+\hline
+\multicolumn{2}{c}{}&\multicolumn{2}{|c|}{Unrolled} & Rolled\\
+\cline{3-5}
+\multicolumn{2}{c|}{}& Squares    &  Circles & \\
+\cline{1-4}
+\multirow{2}{*}{Event} & R1 cont & 2 & 1 & 1\\
+ & R2 cont & 2 & 1 & 1\\
+\hline
+\multirow{2}{*}{State} & R1 cont & N/A & N/A & 1\\
+ & R2 cont & 2 & 1 & 1\\
+\hline
+\end{tabular} 
+\end{center}
+\caption{}
+\label{tab:1}
+\end{table}
+
+As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network.
+Participants waited 5000ms to see a first change of state occurring in a detector, except for the Rolled network case with activation already going around the loop (state condition) -- in this case they had to wait between 3800ms and 5700ms before activation starts in the other branch.
+Activation delay between any two successive squares was set on 100ms as well.
+
+\begin{figure}[ht]
+  \centering
+  \subfloat[Unrolled circles.]{\includegraphics[width=4.2cm]{stim_E2a}\label{fig:3a}}
+  \hfill
+  \subfloat[Rolled.]{\includegraphics[width=4.2cm]{stim_E2b}\label{fig:3b}}
+  \caption{Networks.}
+\end{figure}
+
+\textbf{Results}. Paragraph...
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=8cm]{results_E2}
+\end{center}
+\caption{This is a figure.} 
+\label{fig:4}
+\end{figure}
+
+\textbf{Discussion}. Paragraph...
+
+\section{General discussion}
 

### 2020-09-23 docs/resources/CogSci_paper/paper.tex
last update

--- 2020-09-08 docs/resources/CogSci_paper/paper.tex
@@
 
-Identifying the causes of an event is a very important ability that people largely rely on in common life.
-Even if reasonning about causes doesn't generally pose problems and comes up with clear answers about what caused an event, it is sometimes much more difficult to disambiguous an event and answer such a question.
-When divergent causal intuitions are involved it can be hard to decide for one candidate cause rather than an other.
-This can lead to very different and serious implications, especially legal ones.
-For example a well-known story in the literature is the "Irish case", a real-world example of a court case: [include here Irish case].
-In such legal contexts determining the main causes of an event is an extremely important task that involves finding who or what is at fault in dramatic accidents, property damage, etc.
-and can pose a veritable challenge.
-Here massive fines, jail sentences or death penalties can be the outcomes of the inferred causal history of an event.
-Reasonning about causes, however, is often implied in much more naive and common circonstances and doesn't necessary pose any difficulty --e.g.
-identifying what is the cause of the bottle shattering.
-But even in very simple situations like that, it is sometimes not so easy to theoretically explain and predict intuitive responses that people give.
-[Include here Susy vs Bob example].
-In all these common-life situations we want to know what the \textit{actual cause} of an event is, in a specific context.
-That is we don't want here to infer some general causal laws responsible for the occurrence of an event but understand what is the main or most intuive singular causes in particular cases.
-The current paper focuses on causal judgements of the latter type, that is on the concept of \textit{actual} and not \textit{general} causation\footnote{also called "token" or "singular" causation}, following a distinction that has been frequently made in the litterature.
-Causal judgements have been widely studied and there are existing models of actual causation that aim to capture people's inferences about the causes of a single event.
-Current research on actual causation mainly relies on two different interpretations of causality : the counterfactual account (CF) or the physical process one (PP).
+Identifying the causes of an event involves not only logical rules but also different senses of ``causation'' that people might have.
+Let's imagine that, to mitigate false alarms rate and improve the detection of fires, Lilly decides to install at home a dual alarm system that goes off if and only if two chains of ionization and photoelectric detectors are triggered (Fig.\ref{fig:1}a).
+Lilly wants to test her new system by exposing the detectors to some smoke.
+At some point a ionization detector starts beeping first while measuring a change in the electrical conductivity of air.
+It activates the next detector in the sequence which then activates the last one (Fig.\ref{fig:1}b).
+At this stage Lilly notices that the alarm remains silent.
+After a while a photoelectric detector starts beeping as well while measuring a change in the transparency of air.
+It activates the next detector in the sequence which then activates the last one and right after the alarm goes off (Fig.\ref{fig:1}c).
 
-According to the first interpretation, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B.
-In other words if B would have been different without A, then A is an actual cause.
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=6cm]{intro}
+\end{center}
+\caption{This is a figure.} 
+\label{fig:1}
+\end{figure}
+
+What would Lilly think caused the alarm to go off?
+Several answers can be considered.
+Based on logical dependencies, people's causal judgements can state that both the furthest activated ionization detector (first root change ``R1'') and the furthest activated photoelectric detector (second root change ``R2'') are equally the causes of the alarm activation.
+However a more intuitive answer would be that the provided temporal information influences people's choice toward one cause more than the other one.
+An important selection criterium could be how recent a candidate cause is compared to another one and would lead for instance to the following preferences: R2 $\succ$ R1, and I2 $\succ$ I1 (``I2'' and ``I1'' being respectively the second and first immediate changes).
+Nonetheless we can imagine other cases where people potentially rely in part on another type of temporal criterium: let's imagine that the activation of the ionization detectors starts before and finishes after the activation of the photoelectric detectors (such that R1 and I2 belong to the same sequence and \textit{idem} for R2 and I1).
+In that case the activation of the furthest photoelectric detector is still the most recent but it doesn't initiate a sequence that leads to the activation of the alarm anymore, whereas this is the opposite for the furthest ionization detector.
+In such situations we want to know what people would consider as the main cause of the system.
+
+The alarm case can be seen as an example of ``singular causation''\footnote{Also called ``token'' or ``actual causation'' by opposition with ``general causation'' -- following a distinction that has been frequently made in the litterature.} where we want to understand, in a singular context, what caused a particular event to occurr at the time it did.
+More generally it instantiates a class of causal systems where an event is a common effect of distinct sequences of necessary causes.
+[[Those kind of causal systems are very common in real life and can involve finding who or what is at fault in dramatic accidents, property damage, crimes, etc..
+Understanding why people, in a specific situation, decide for one candidate cause rather than another one can lead to better support important decisions (e.g.
+legal ones) and help us better predict causal judgements that are made in specific contexts.]] Causal judgements have been widely studied in the litterature, however we think that the temporal order in which each variable changes its value is an important feature of causal systems that have been hardly studied nor tested experimentally as we intended to do.
+Indeed current research on actual causation mainly relies on two different interpretations of causality that don't ground causal judgements on temporal information: the counterfactual (CF) and the physical process (PP) accounts.
+
+According to the CF account, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B.
 More precisely A is a cause of B if and only if A and B are true, and A hadn't occurred B wouldn't have occurred.
-[Example of Suzy only] This type of explanation calls upon counterfactual scenario or possible worlds where the presumed cause of the effect is removed from the system while all other relevant factors are kept as unchanged as possible compared to the actual world.
-However, a first well-known challenge that the CF account encounters is that it is not clear which counterfactual world we have to take into account in order to justify a singular cause, and how to accurately compare those counterfactuals.
-Again let's say that Suzy throws a rock at a bottle but this time Bob as well throws a rock at the same bottle with the same precision as Suzy.
-Suzy's rock reaches the bottle just a second before Bob's and breaks it.
-According to the CF account if Suzy hadn't thrown the rock, the bottle would have still shattered since Bob's rock would have hit it regardless.
-So following the general idea mentionned above, Suzy's throw cannot be the cause of the bottle shattering -- which is counter-intuitive.
-This is an example of a general type of cases we call \textit{late preemption} and refers to the fact that the final event -- i.e.
-the effect -- is "overdetermined" [Hall].
-A first common attempt to face the problem is to say that the bottle-shattering event that results from Suzy's throw is qualitatively different from the the bottle-shattering event that follows from Bob's throw.
-In other words A is a cause of B if and only if A and B occur, and A hadn't occurred B wouldn't have occurred at all or would have been delayed.
-Thus Suzy's throw is indeed the cause of the bottle-shattering event that occurs at the time it did.
+This interpretation calls upon counterfactual scenario or possible worlds where the presumed cause of the effect is removed from the system while all other relevant factors are kept as unchanged as possible compared to the actual world.
+According to us the main problem of this interpretation is that its best models explicitely rejects the idea of grounding causation on temporal information.
+In that framework events are represented as mere propostions and causation is thought as a relation between static states.
 
-[Halpern and Pearl solution : graphical models, static, pb of modelling, without time, etc.] 
-[Then, more general problem with counterfactual dependance].
-
-The physical process account of causation gives a completely different interpretation of causality and states that an event A causes B when there is a physical connection between them.
-Several theories have been proposed to caracterise this physical connection.
-In one of its latest and probably most convincing formulation, we have to distinguish two things : causal process and causation or causal interaction.
-A causal process is a physical process involving an object which conserves a certain quantity accross space and time.
-The conserved quantity is typically a measurable physical property of the object like the mass-energy, etc.
-and is precisely what scientific theories are meant to formalize.
-In that sence both Suzy's and Bob's throw are causal processes.
-A causal interaction or causation is an exchange of that conserved quantity, for example the mass-energy of the rock to the glass of the bottle.
-According to this definition, only Suzy's throw involves a causal interaction with the bottle.
-However the PP account of causation doesn't consider as genuine causation some very common cases that people intuitively judge as proper example of causation.
-For example if I hold the head of my ennemy under water and make him die, I'm not genuinely the cause of his death; rather I'm actually preventing the possibility of a genuine causation which is the physical process of breathing oxygen in order to live.
-The main problem is that, as a consequence, the PP account not only goes sometimes against some of our best causal intuitions, but also needs the counterfactual analysis as well to ground cases of \textit{quasi-causation} by omission or prevention like the latter one.
-
-[?? A third account : probability raising ??]
+According to the PP account, an event A causes B when there is a physical connection between them.
+In one of its latest and probably most convincing formulation, this theory distinguishes two things : causal process and causal interaction (i.e.
+causation).
+A causal process is a physical process involving an object which conserves a certain quantity, like mass-energy, accross space and time.
+A causal interaction or causation is an exchange of that conserved quantity.
+However the theory goes sometimes against some of our best causal intuitions and to makes implicitely use of the CF analysis to explain our judgements: if I hold the head of my ennemy under water and make him die, I'm not genuinely the cause of his death; rather I'm actually preventing the possibility of a genuine causation which is breathing oxygen in order to live.
+Moreover, the PP account is meant to give an epistemological definition of the concept of causation and not an explanation of how people actually rely on temporal informations to make causal judgements.
 
@@
 
-As a result of the preceeding analysis it seems that none of the current main theories of actual causation gives a satisfying explanation of the causal judgements that people actually do.
-This paper aims at solving the problem by interpreting causality in a completely different way.
-In the CF framework events are represented as mere propostions and causation is thought as a relation between static states.
-The PP account is meant to give an epistemological definition of the concept of causation and not an explanation of how people actually rely on temporal informations to make causal judgements.
-By contrast this paper focuses on the role of temporal information for inferring causal relationship between events qua changes of states over time.
-In other words we hypothesised that people's causal judgements are influenced by not only the values of all relevant variables at the time the effect occurs, but also the temporal order in which these variables took their values.
-[??
-Add a concrete example ??] This approach has some precedent in the literature but has not been developed carefully and has never been tested experimentally as we intend to do.
+As a result of the preceeding analysis it seems that none of the current main theories of actual causation has focused on the role of temporal information for inferring causal relationship between events qua changes of states over time.
+In contrast we hypothesised that people's causal judgements are influenced by not only the values of all relevant variables at the time the effect occurs, but also the temporal order in which these variables took their values.
+This approach has some precedent in the literature but has not been developed carefully and has never been tested experimentally as we intend to do.
 We add further that people mainly identify causation with a \textit{continuous} sequence of changes of states over time.
@@
 According to our \textit{no coincidence principle} $\mathcal{U}$ is either empty or a singleton -- including only one parent whose value changed.
+Lets say that at $t-1$ we find a change in a parent variable $Y$, so $\mathcal{U}^Z_{\bm{y}_{i\rightarrow j}}(t-1)=\{\mu^{Z}_{y_{i\rightarrow j}}(t-1)\}$.
 Then we want to find the set of changes such that $\mathcal{U}^Y_{\bm{x}_{i\rightarrow j}}(t-2)\rightarrow \mu^{Z}_{y_{i\rightarrow j}}(t-1)$, that is the set of immediate previous changes in the parents $\bm{X}$ of the variable $Y$ at time $t-2$ that led to the observed change in $Y$ at time $t-1$.
 Following the same logic we suggest that if $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)\rightarrow \mu^{Y}_{x_{i\rightarrow j}}(t-2)$ is such that $\mathcal{U}^X_{\bm{w}_{i\rightarrow j}}(t-3)=\emptyset$, then it means that $\mu^{Y}_{x_{i\rightarrow j}}(t-2)$ represents the chronologically first change along the path, occurring in $X$, and we suggest that this change is identified as being the main cause of $Z$.
+
+\section{Experiments}
+
+To test our hypothesis that the main cause of an effect is the root change that initiates a continuous sequences of changes until the occurrence of the effect, participants were presented animations showing activation spreading over networks of nodes up to the final node.
+We run three different experiments which shared the same plot that we wanted as intuitive as possible: participants were told that they were working in a nuclear control room and that their job was to monitor networks of particule detectors.
+The instructions said that when a detector, depicted by a square or circle (see below), absorbs a radioactive particule it becomes active and turns black, transmitting the activation across the links of the network so that an active detector activates the next one in the chain and so forth.
+All the networks include a special component, called 'Gauge of Critical Moment', that becomes active only if all of its input from the detectors it is connected to are active.
+The Gauge of Critical Moment is always represented by a square with ``GCM'' (Experiment 1) or ``G'' (Experiment 2) above.
+At the end of an activation sequence, it is asked to click on the detector(s) they considered as the main cause(s) of the activation of the Gauge of Critical Moment.
+Examples of network activations were shown in the instructions and participants had to answer to answer a survey at the end to see if they correctly understood the instructions.
+If they made a mistake they had to read again all the instructions and answer again the survey.
+
+In both experiments we included only two types of input-output logic, namely single input (chains) and AND-Gate (branches where the effect needs the activation of two inputs to occur).
+When people were presented chains, we first wanted to see if they identified the main cause of the effect with a change of state that occurred in the chain the furthest away from effect (\textit{root change}) or the closest to it (\textit{immediate change}).
+When they were presented AND-Gates, we wanted to see if they identified the main cause of the effect with the root change that occurred chronologically first (\textit{1\textsuperscript{st} root change}, labeled ``R1'' henceforth), the root change that occurred chronologically last (\textit{2\textsuperscript{nd} root change}, ``R2'' henceforth), the immediate change that occurred chronologically first (\textit{1\textsuperscript{st} immediate change}, ``I1'') or the immediate change that occurred chronologically last (\textit{2\textsuperscript{nd} immediate change}, ``I1'').
+
+\subsection{Experiment 1}
+
+The objective of this experiment was to see whether people's causal judgement were influenced by: the length of the sequences of changes (for both chains and AND-Gates), and the delay between the 1\textsuperscript{st} immediate change and the 2\textsuperscript{nd} root change, that is between the end of the first sequence of changes and the beginning of the second sequence of changes (for AND-Gates only).
+
+\textbf{Participants}.
+The experiment was hosted by Google Cloud App Engine and 30 participants were recruited via Amazon Mahchanical Turk.
+There were 10 females (average age: 37.3) and 20 males (average age: 37.6).
+28 participants were english speakers, 1 were italian speaker and 1 marathi speaker.
+
+\textbf{Method}.
+Each participant was presented 15 different networks of square detectors.
+Square detectors maintain their activation through time.
+Participants were first presented the network in its initial and static state and were asked to click the ``Run'' button to observe an activation sequence over the network.
+They waited 5000ms to see a first change of state occurring in a detector and the activation delay between any two successive detectors was set on 100ms (see Fig.\ref{fig:1}).
+
+Three stimuli were chains that differed in length of activation sequence, that is including 2, 4 or 7 activated squares (with the GCM).
+These lengths were labeled ``Short'', ``Medium'' and ``Long''.
+The AND-Gates were similarily categorised ``Short'', ``Medium'' (like in Fig.\ref{fig:1}) and ``Long'' with both branches being the same length.
+In each length category participants were shown three similar AND-Gates that differed uniquely in the delays between the 1\textsuperscript{st} immediate change and the 2\textsuperscript{nd} root change.
+The delays were 2000ms, 4000ms and 6000ms.
+
+All the stimuli were presented in a random order.
+Two groups, ``Left'' and ``Right'', were designed and partcipants were randomly assigned to either group at the begining of the experiment.
+In the ``Left'' group all the networks were shifted towards left (the GCM being on the left) and in the ``Right'' group all the networks were shifted towards right.
+For each AND-Gate participants saw randomly either the top branch or the bottom branch activating first.
+
+After the activation of the GCM (the end of the animation) participants had to wait 1000ms before the squares became clickable and the following intruction appeared: ``In this sequence what caused the activation of the GCM?
+Respond by clicking on a detector''.
+They had the option to run again the animation (no more than 9 times) or to go to the next network assuming they clicked on one detector -- they couldn't select more than one detector.
+When selected the edges of a detector turned red.
+
+\textbf{Results}. Paragraph...
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=8cm]{results_E1}
+\end{center}
+\caption{This is a figure.} 
+\label{fig:2}
+\end{figure}
+
+\textbf{Discussion}. Paragraph...
+
+\subsection{Experiment 2}
+
+The objective of the experiment was this time to see if people's causal judgements are
+
+\textbf{Participants}.
+The experiment was hosted by Google App Engine and 89 participants were recruited via Amazon Mechanical Turk.
+There were 29 females (average age: 39.7), 59 males (average age: 37.6) and 1 declared ``other''.
+All the participants were english speakers.
+
+\textbf{Method}.
+Like in Experiment 1 each participant was presented 15 different networks.
+However this time we introduced round detectors that, contrarily to square detectors, don't maintain their activation through time.
+By adding this feature we were able to include networks that have a loop of detectors (see Fig.\ref{fig:3b}).
+1 stimulus was a chain of square detectors and 1 stimulus was a chain of round detectors were included (for the latter one, only ``G'' remained a square).
+The other 13 stimuli were AND-Gates and were sorted into three main categories: 4 stimuli were Rolled networks, i.e.
+with a loop, as depicted in Fig.\ref{fig:3b}, 3 were Unrolled networks with square and round detectors as in Fig.\ref{fig:3a}, labeled ``Unrolled circles'' and 6 stimuli were Unrolled with square detectors in place of the round one (not represented here), labeled ``Unrolled squares''.
+For each category of AND-Gates all the stimuli were sorted according to two crossed dimensions: ``Event'' \textit{vs} ``State''and ``R1 cont'' \textit{vs} ``R2 cont''.
+In condition ``Event'' none of the detectors is initially active or -- if the branch is a loop -- is continuously (re-)activated.
+In condition ``State'' one of the two branches contains initially detectors that are active or is continuously (re-)activated.
+In condition ``R1 cont'' (represented in Fig.\ref{fig:3a}) the activation is such that there is a continuous sequence of changes going from R1 to I2 that leads to the activation of ``G''.
+In condition ``R2 cont'' (not represented here) this is R2 that is continously connected to I2 that leads to the activation of ``G''.
+The design of the experiment for the AND-Gates is given in Tab.\ref{tab:1}.
+For the AND-Gates (Rolled and Unrolled) the delay between I1 and I2 in condition ``Event'' was maintained at 11 nodes (i.e.1100 ms) for almost all the networks.
+There was only one Unrolled squares network for which the delay was set at 8 nodes (800ms).
+
+\begin{table}[ht]
+\begin{center} 
+\vskip 0.12in
+\begin{tabular}{c|c|c|c|c} 
+\hline
+\multicolumn{2}{c}{}&\multicolumn{2}{|c|}{Unrolled} & Rolled\\
+\cline{3-5}
+\multicolumn{2}{c|}{}& Squares    &  Circles & \\
+\cline{1-4}
+\multirow{2}{*}{Event} & R1 cont & 2 & 1 & 1\\
+ & R2 cont & 2 & 1 & 1\\
+\hline
+\multirow{2}{*}{State} & R1 cont & N/A & N/A & 1\\
+ & R2 cont & 2 & 1 & 1\\
+\hline
+\end{tabular} 
+\end{center}
+\caption{}
+\label{tab:1}
+\end{table}
+
+As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network.
+Participants waited 5000ms to see a first change of state occurring in a detector, except for the Rolled network case with activation already going around the loop (state condition) -- in this case they had to wait between 3800ms and 5700ms before activation starts in the other branch.
+Activation delay between any two successive squares was set on 100ms as well.
+
+\begin{figure}[ht]
+  \centering
+  \subfloat[Unrolled circles.]{\includegraphics[width=4.2cm]{stim_E2a}\label{fig:3a}}
+  \hfill
+  \subfloat[Rolled.]{\includegraphics[width=4.2cm]{stim_E2b}\label{fig:3b}}
+  \caption{Networks.}
+\end{figure}
+
+\textbf{Results}. Paragraph...
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=8cm]{results_E2}
+\end{center}
+\caption{This is a figure.} 
+\label{fig:4}
+\end{figure}
+
+\textbf{Discussion}. Paragraph...
+
+\section{General discussion}
 

### 2020-09-23 docs/resources/CogSci_paper/paper.tex
last update

@@
 
-The objective of the experiment was this time to see if people's causal judgements are
+The objective of the experiment was this time to see which event is seen as the main cause in cases where R1 and I2 are located on the same sequence.
+In other words we wanted to test the hypothesis that the main cause is the event that initiates a continous sequence of changes over time.
+We wanted to test it with different types of networks including networks with loop of detectors that don't maintain their activation through time and see if the type of networks influence as well causal judgements.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \begin{center}

### 2020-09-24 experiments/Experiment4_NO_CIRCLES/100ms/source/js/plugins/jspsych-causal-graph3.js
E4 added

@@ -1,0 +1,428 @@
+jsPsych.plugins['causal-graph3'] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'causal-graph1',
+		description: '',
+		parameters: {
+			stim_GIF_A: {
+			type: jsPsych.plugins.parameterType.IMAGE,
+			pretty_name: 'Stimulus',
+			default: undefined,
+			description: 'Image to be displayed.'
+			},
+			stim_GIF_B: {
+			type: jsPsych.plugins.parameterType.IMAGE,
+			pretty_name: 'Stimulus',
+			default: null,
+			array: true,
+			description: 'Image to be displayed.'
+			},
+			stim_GIF_C: {
+			type: jsPsych.plugins.parameterType.IMAGE,
+			pretty_name: 'Stimulus',
+			default: undefined,
+			description: 'Image to be displayed.'
+			},
+			loop_state : {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'Loop state',
+			default: null,
+			description: 'Whether the loop is in the state condition or not.'
+			},
+			stim_height: {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'Stimulus height',
+			default: 100,
+			description: 'Height of images in pixels.'
+			},
+			stim_width: {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'Stimulus width',
+			default: 100,
+			description: 'Width of images in pixels'
+			},
+			GIF_B_duration: {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'GIF duration',
+			default: null,
+			array: true,
+			description: 'How long the animation lasts.'
+			},
+			status: {
+			type: jsPsych.plugins.parameterType.STRING,
+			pretty_name: 'The status of the plugin',
+			default: 'active',
+			description: 'Whether the causal graph is active or not.'
+			},
+			A_coord: {
+			type: jsPsych.plugins.parameterType.OBJECT,
+			pretty_name: 'Nodes coordinates',
+			default: null,
+			array: true,
+			description: 'Coordinates of the round detectors.'
+			},
+			B_coord: {
+			type: jsPsych.plugins.parameterType.OBJECT,
+			pretty_name: 'Nodes coordinates',
+			default: null,
+			array: true,
+			description: 'Coordinates of the square detectors.'
+			},
+			prompt: {
[…]

### 2021-01-13 docs/resources/CogSci_paper/paper.tex
last update

--- 2020-09-23 docs/resources/CogSci_paper/paper.tex
@@
 
-\title{How to Make a Proceedings Paper Submission}
+\title{Actual Causes as Changes of State in Continuous Time}
 
@@
 \begin{center}
-\includegraphics[width=6cm]{intro}
+\includegraphics[width=7cm]{intro}
 \end{center}
-\caption{This is a figure.} 
+\caption{An example of an AND-Gate causal system.
+In an AND-Gate for an effect to occur it needs the occurence of all the causes it is linked to.
+Here for the smoke alarm to go off, both the ionization and photoelectric detectors it is connected to have to be activated.
+In this scenario the detectors activation spreads over the top chain first (b) and then over the bottom chain leading to the alarm activation (c).}
 \label{fig:1}
@@
 An important selection criterium could be how recent a candidate cause is compared to another one and would lead for instance to the following preferences: R2 $\succ$ R1, and I2 $\succ$ I1 (``I2'' and ``I1'' being respectively the second and first immediate changes).
-Nonetheless we can imagine other cases where people potentially rely in part on another type of temporal criterium: let's imagine that the activation of the ionization detectors starts before and finishes after the activation of the photoelectric detectors (such that R1 and I2 belong to the same sequence and \textit{idem} for R2 and I1).
-In that case the activation of the furthest photoelectric detector is still the most recent but it doesn't initiate a sequence that leads to the activation of the alarm anymore, whereas this is the opposite for the furthest ionization detector.
+Nonetheless we can imagine other cases where people potentially rely in part on another type of temporal criterium: let's imagine that the activation of the ionization detectors starts before and finishes after the activation of the photoelectric detectors, such that R1 and I2 are linked by a \textit{continuous sequence of changes}.
+In that case this is the activation of the oldest root change R1 and not the most recent root change R2 that initiate a continuous sequence of changes until the activation of the alarm.
 In such situations we want to know what people would consider as the main cause of the system.
@@
 The alarm case can be seen as an example of ``singular causation''\footnote{Also called ``token'' or ``actual causation'' by opposition with ``general causation'' -- following a distinction that has been frequently made in the litterature.} where we want to understand, in a singular context, what caused a particular event to occurr at the time it did.
-More generally it instantiates a class of causal systems where an event is a common effect of distinct sequences of necessary causes.
-[[Those kind of causal systems are very common in real life and can involve finding who or what is at fault in dramatic accidents, property damage, crimes, etc..
-Understanding why people, in a specific situation, decide for one candidate cause rather than another one can lead to better support important decisions (e.g.
-legal ones) and help us better predict causal judgements that are made in specific contexts.]] Causal judgements have been widely studied in the litterature, however we think that the temporal order in which each variable changes its value is an important feature of causal systems that have been hardly studied nor tested experimentally as we intended to do.
+Here specifically it instantiates a class of causal systems where an event is a common effect of distinct sequences of necessary causes.
+Causal judgements have been widely studied in the litterature, however the temporal order in which each variable changes its value is an important feature of causal systems that have been hardly studied nor tested experimentally as we intended to do.
 Indeed current research on actual causation mainly relies on two different interpretations of causality that don't ground causal judgements on temporal information: the counterfactual (CF) and the physical process (PP) accounts.
+On the one hand, in the CF framework, events are represented as mere propostions and causation is thought as a relation between static states (A is a cause of B if and only if A and B are true, and A hadn't occurred B wouldn't have occurred).
+On the other hand the PP account is meant to give an epistemological definition of the concept of causation and not an explanation of how people actually rely on temporal informations to make causal judgements.
 
-According to the CF account, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B.
-More precisely A is a cause of B if and only if A and B are true, and A hadn't occurred B wouldn't have occurred.
-This interpretation calls upon counterfactual scenario or possible worlds where the presumed cause of the effect is removed from the system while all other relevant factors are kept as unchanged as possible compared to the actual world.
-According to us the main problem of this interpretation is that its best models explicitely rejects the idea of grounding causation on temporal information.
-In that framework events are represented as mere propostions and causation is thought as a relation between static states.
-
-According to the PP account, an event A causes B when there is a physical connection between them.
-In one of its latest and probably most convincing formulation, this theory distinguishes two things : causal process and causal interaction (i.e.
-causation).
-A causal process is a physical process involving an object which conserves a certain quantity, like mass-energy, accross space and time.
-A causal interaction or causation is an exchange of that conserved quantity.
-However the theory goes sometimes against some of our best causal intuitions and to makes implicitely use of the CF analysis to explain our judgements: if I hold the head of my ennemy under water and make him die, I'm not genuinely the cause of his death; rather I'm actually preventing the possibility of a genuine causation which is breathing oxygen in order to live.
-Moreover, the PP account is meant to give an epistemological definition of the concept of causation and not an explanation of how people actually rely on temporal informations to make causal judgements.
+In this paper we present two experiments that show the importance of taking into account temporal information to explain the causal judgements people make.
+In particular we relied on the intuition behind the example above that the chronological order and the temporal relationship between distinct causes can bias people toward an event rather than another in considering which one is the main cause of the effect.
 
@@
 
-As a result of the preceeding analysis it seems that none of the current main theories of actual causation has focused on the role of temporal information for inferring causal relationship between events qua changes of states over time.
-In contrast we hypothesised that people's causal judgements are influenced by not only the values of all relevant variables at the time the effect occurs, but also the temporal order in which these variables took their values.
+So far it seems that none of the current main theories of actual causation has focused on the role of temporal information for inferring causal relationship between events qua changes of states over time.
+In contrast we hypothesised that people's causal judgements are influenced not only by the values of all relevant variables at the time the effect occurs, but also the temporal order in which these variables took their values.
 This approach has some precedent in the literature but has not been developed carefully and has never been tested experimentally as we intend to do.
@@
 To test our hypothesis that the main cause of an effect is the root change that initiates a continuous sequences of changes until the occurrence of the effect, participants were presented animations showing activation spreading over networks of nodes up to the final node.
-We run three different experiments which shared the same plot that we wanted as intuitive as possible: participants were told that they were working in a nuclear control room and that their job was to monitor networks of particule detectors.
+We run two experiments which shared the same plot that we wanted as intuitive as possible: participants were told that they were working in a nuclear control room and that their job was to monitor networks of particule detectors.
 The instructions said that when a detector, depicted by a square or circle (see below), absorbs a radioactive particule it becomes active and turns black, transmitting the activation across the links of the network so that an active detector activates the next one in the chain and so forth.
@@
 
-The objective of this experiment was to see whether people's causal judgement were influenced by: the length of the sequences of changes (for both chains and AND-Gates), and the delay between the 1\textsuperscript{st} immediate change and the 2\textsuperscript{nd} root change, that is between the end of the first sequence of changes and the beginning of the second sequence of changes (for AND-Gates only).
+The objective of this experiment was to see whether people's causal judgement were influenced by: the length of the sequences of changes (for chains and AND-Gates), and -- for AND-Gates only -- the delay between the last activation of the first branch (I1) and the first activation of the second branch (R2).
+In this experiment activation starts to spread over the second branch after the activation reached the end of the first branch (like in Fig.\ref{fig:1}).
 
@@
 Participants were first presented the network in its initial and static state and were asked to click the ``Run'' button to observe an activation sequence over the network.
-They waited 5000ms to see a first change of state occurring in a detector and the activation delay between any two successive detectors was set on 100ms (see Fig.\ref{fig:1}).
+They waited 5000ms to see a first change of state occurring in a detector and the activation delay between any two successive detectors was set on 100ms.
 
@@
 These lengths were labeled ``Short'', ``Medium'' and ``Long''.
-The AND-Gates were similarily categorised ``Short'', ``Medium'' (like in Fig.\ref{fig:1}) and ``Long'' with both branches being the same length.
-In each length category participants were shown three similar AND-Gates that differed uniquely in the delays between the 1\textsuperscript{st} immediate change and the 2\textsuperscript{nd} root change.
-The delays were 2000ms, 4000ms and 6000ms.
+The AND-Gates were similarily categorised ``Short'', ``Medium'' (the one in Fig.\ref{fig:1}) and ``Long'' with both branches being the same length.
+In each length category participants were shown three similar AND-Gates that differed uniquely in the delays between I1 and R2.
+The delays were 2s, 4s and 6s.
 
@@
 
-\textbf{Results}. Paragraph...
+\textbf{Results}.
+First of all we observed no significant effect [STATISTICS?] of neither the orientation of the networks (left \textit{vs} right) nor the location where the activation starts first (top \textit{vs} bottom).
+On average participants took 14 min to complete the experiment and overall 96\% of the time participants run an animation only once.
+Second of all the results show two main findings: first the length of the activation sequences (one, three or six detectors in each branch of the AND-Gates) has no effect on the response pattern; second neither the delay between I1 and R2 (2000ms, 4000ms and 6000ms) nor the fact that some detectors were already activated in State condition have a significant effect on causal judgement.
+So overall people massively chose the last root change R2 as the main cause of the GCM activation for medium and long networks (and R2/I2 for the short networks since both were identical).
+For the short networks (state and event types combined) 25\% \textit{vs} 65\% of participants chose respectively R1/I1 \textit{vs} R2/I2 as the main cause; for the medium networks (state and event combined) we got 23\% \textit{vs} 53\% for respectively R1 and R2 and 20\% \textit{vs} 55\% for the long networks (Fig.\ref{fig:2}).
+
+[WARNING: THE GRAPH BELOW SHOW THE RESULTS FOR SHORT, MEDIUM, LONG \textbf{LENGTHS} (ALL THE DELAYS MIXED) SO NOT THE DIFFERENT DELAYS BETWEEN I1 AND R2.
+BUT AS OUR FOCUS IS ON TEMPORAL INFORMATION, MAYBE WE SHOULD GRAPHICALLY SHOW THAT THE DELAY HAS NO EFFECT??
+BUT IT MEANS THAT WE HAVE TO ADD TWO MORE COLUMNS FOR EACH DETECTOR CATEGORY]
 
@@
 \end{center}
-\caption{This is a figure.} 
+\caption{Comparison of the results for different length of activation in AND-Gates.
+Short: 1 activated detector in each branch.
+Medium: 3 activated detectors in each branch.
+Long: 6 activated detectors in each branch.
+``State'' responses correspond to the networks where one branch was already activated before participants run the animation.
+``Event'' responses correspond to the networks where no detectors were activated before and include results for 2s, 4s and 6s delays between I1 and R2.
+BUT SEE WARNING ABOVE.}
 \label{fig:2}
@@
 
-\textbf{Discussion}. Paragraph...
-
 \subsection{Experiment 2}
 
-The objective of the experiment was this time to see if people's causal judgements are
+In the previous experiment we have seen that people generally chose R2 (identical to I2 in the short sequences) as the main the cause of the activation of the GCM.
+These results make sense since R2 and I2 are linked by a continuous sequence of changes.
+But this set up, where the first element of a chain ``waits'' for the occurrence of the last event of another chain before it can initiate a second sequence, is probably not the most common scenario in real life.
+For instance it might take more time for one causal process to end so that the causal process that has occurred first is not necessarily the one that ends first.
+In Experiment 2 we therefore wanted to break the temporal continuity between the root changes and the immediate changes of Experiment 1 such that the continuous sequence of changes be sometimes between R1 and I2 or R2 and I2.
+We predicted that when there is a continuous sequence of changes between R1 and I2, people would preferably choose R1 as the main cause.
 
 \textbf{Participants}.
-The experiment was hosted by Google App Engine and 89 participants were recruited via Amazon Mechanical Turk.
-There were 29 females (average age: 39.7), 59 males (average age: 37.6) and 1 declared ``other''.
-All the participants were english speakers.
+The experiment was hosted by Google App Engine and 98 participants were recruited via Amazon Mechanical Turk.
+There were 40 females (average age: 40.4), 57 males (average age: 37.6) and 1 declared ``other''.
+All the participants were english speakers except for 2 participants.
 
 \textbf{Method}.
-Like in Experiment 1 each participant was presented 15 different networks.
-However this time we introduced round detectors that, contrarily to square detectors, don't maintain their activation through time.
-By adding this feature we were able to include networks that have a loop of detectors (see Fig.\ref{fig:3b}).
-1 stimulus was a chain of square detectors and 1 stimulus was a chain of round detectors were included (for the latter one, only ``G'' remained a square).
-The other 13 stimuli were AND-Gates and were sorted into three main categories: 4 stimuli were Rolled networks, i.e.
-with a loop, as depicted in Fig.\ref{fig:3b}, 3 were Unrolled networks with square and round detectors as in Fig.\ref{fig:3a}, labeled ``Unrolled circles'' and 6 stimuli were Unrolled with square detectors in place of the round one (not represented here), labeled ``Unrolled squares''.
-For each category of AND-Gates all the stimuli were sorted according to two crossed dimensions: ``Event'' \textit{vs} ``State''and ``R1 cont'' \textit{vs} ``R2 cont''.
-In condition ``Event'' none of the detectors is initially active or -- if the branch is a loop -- is continuously (re-)activated.
-In condition ``State'' one of the two branches contains initially detectors that are active or is continuously (re-)activated.
-In condition ``R1 cont'' (represented in Fig.\ref{fig:3a}) the activation is such that there is a continuous sequence of changes going from R1 to I2 that leads to the activation of ``G''.
-In condition ``R2 cont'' (not represented here) this is R2 that is continously connected to I2 that leads to the activation of ``G''.
-The design of the experiment for the AND-Gates is given in Tab.\ref{tab:1}.
-For the AND-Gates (Rolled and Unrolled) the delay between I1 and I2 in condition ``Event'' was maintained at 11 nodes (i.e.1100 ms) for almost all the networks.
-There was only one Unrolled squares network for which the delay was set at 8 nodes (800ms).
-
-\begin{table}[ht]
-\begin{center} 
-\vskip 0.12in
-\begin{tabular}{c|c|c|c|c} 
-\hline
-\multicolumn{2}{c}{}&\multicolumn{2}{|c|}{Unrolled} & Rolled\\
-\cline{3-5}
-\multicolumn{2}{c|}{}& Squares    &  Circles & \\
-\cline{1-4}
-\multirow{2}{*}{Event} & R1 cont & 2 & 1 & 1\\
- & R2 cont & 2 & 1 & 1\\
-\hline
-\multirow{2}{*}{State} & R1 cont & N/A & N/A & 1\\
- & R2 cont & 2 & 1 & 1\\
-\hline
-\end{tabular} 
-\end{center}
-\caption{}
-\label{tab:1}
-\end{table}
+In Experiment 2 each participant was presented 7 different networks.
+One stimulus was a chain and all the other six stimuli were AND-Gates.
+Like in Experiment 1 we tested for the dimension ``Event'' \textit{vs} ``State''.
+Same as before in condition ``Event'' none of the detectors was initially active and in condition ``State'' one of the two branches was containing detectors that were already active before the animation started.
+However this time, within the condition ``Event'',  we also tested for the extra dimension ``R1 cont'' \textit{vs} ``R2 cont''.
+In condition ``R1 cont'' (represented in Fig.\ref{fig:3}) the activation is such that there is a continuous sequence of changes going from R1 to I2 leading to the activation of G.
+In condition ``R2 cont'' (not represented here) this is R2 that was continously connected to I2 leading to the activation of G.
+For the AND-Gates the delay between I1 and I2 in condition ``Event'' was maintained at 11 nodes (i.e.1100 ms) for almost all the networks.
+There was only one network for which the delay was set at 8 nodes (800ms).
 
 As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network.
-Participants waited 5000ms to see a first change of state occurring in a detector, except for the Rolled network case with activation already going around the loop (state condition) -- in this case they had to wait between 3800ms and 5700ms before activation starts in the other branch.
-Activation delay between any two successive squares was set on 100ms as well.
-
-\begin{figure}[ht]
-  \centering
-  \subfloat[Unrolled circles.]{\includegraphics[width=4.2cm]{stim_E2a}\label{fig:3a}}
-  \hfill
-  \subfloat[Rolled.]{\includegraphics[width=4.2cm]{stim_E2b}\label{fig:3b}}
-  \caption{Networks.}
-\end{figure}
-
-\textbf{Results}. Paragraph...
+Participants waited 5000ms to see a first change of state occurring in a detector.
+Activation delay between any two successive square activations was set on 100ms as well.
+All the stimuli were presented in a random order and participants were assigned to either group "Left" or "Right" as in Experiment 1, and each AND-Gate was randomly presented with the curved chain (see Fig.\ref{fig:3})) at the top or at the bottom.
 
@@
 \begin{center}
+\includegraphics[width=8cm]{stim_E2a}
+\end{center}
+\caption{An example of an activation spreading over a network of Experiment 2.
+Here a continuous sequence of changes links R1 and I2 -- in other words this is the sequence of activations initiated by the chronologically first root change (curved branch) that eventually leads to the activation of the last immediate change and therefore G.}
+\label{fig:3}
+\end{figure}
+
+\textbf{Results}.
+First we observed no significant effect (statistics?) of neither the orientation of the networks (left \textit{vs} right) nor the location of the curved chain (top \textit{vs} bottom).
+On average participants took 8 min to complete the experiment and overall 91\% of the time participants run an animation only once.
+Second the reversed response pattern for R1 and R2 between ``R1 cont'', ``R2 cont'' and ``State'' stimuli type shown in (Fig.\ref{fig:4}) confirms our hypothesis: when this is the first root change R1 that initiates a continuous sequence of changes until the activation of G, R1 is considered the cause of the activation of G about two times more than R2 (31\% against 15\%) while this is the opposite when this is the second root change that initiates a continuous sequence of changes (18\% against 29\%).
+Coherently, for ``State'' condition (when this is R2 that initiates the continuous sequence of changes) participants prefer R2 over R1 in a very large extent (47\% against 5\%).
+As expected the type of network had however no effect on the responses for I1 nor I2 and the percentages of responses for I2 and the root change that initiates the continuous sequence for the Event network (31\% and 34\% for respectively the ``R1 cont'' and ``R2 cont'' networks) were approximately the same.
+
+\begin{figure}[H]
+\begin{center}
 \includegraphics[width=8cm]{results_E2}
 \end{center}
-\caption{This is a figure.} 
+\caption{Results for AND-Gates where the continuous sequence of changes is either between R1 and I2 or between R2 and I2.
+``Event (R1 cont)'': responses for the networks where no detectors were already active before the participant runs the animation and where the continuous sequence of changes is between R1 and I2.
+``Event (R2 cont)'': responses for the networks where no detectors were already active and where the continuous sequence is between R2 and I2.
+``State (R2cont)'': responses for networks where one branch was already activated before the participant runs the animation and where the continuous sequence of activation is necessarily initiated by R2.}
 \label{fig:4}
@@
 
-\textbf{Discussion}. Paragraph...
+\newpage
+\section{General discussion}
 
-\section{General discussion}
+Outline:
+
+\textbf{I}. Summary and implications of the results
+
+\textbf{II}.
+To say why these results cannot really be explained by the current theories of causation and why our theory can better account for them.
+I don't know if I should talk more about CF and PP accounts than I do in introduction, like saying something like that is probably not useful:
+
+[\textit{According to the CF account, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B.
+More precisely A is a cause of B if and only if A and B are true, and A hadn't occurred B wouldn't have occurred.
+This interpretation calls upon counterfactual scenario or possible worlds where the presumed cause of the effect is removed from the system while all other relevant factors are kept as unchanged as possible compared to the actual world.
+According to us the main problem of this interpretation is that its best models explicitely rejects the idea of grounding causation on temporal information.
+According to the PP account, an event A causes B when there is a physical connection between them.
+In one of its latest and probably most convincing formulation, this theory distinguishes two things : causal process and causal interaction (i.e.
+causation).
+A causal process is a physical process involving an object which conserves a certain quantity, like mass-energy, accross space and time.
+A causal interaction or causation is an exchange of that conserved quantity.
+However the theory goes sometimes against some of our best causal intuitions and to makes implicitely use of the CF analysis to explain our judgements: if I hold the head of my ennemy under water and make him die, I'm not genuinely the cause of his death; rather I'm actually preventing the possibility of a genuine causation which is breathing oxygen in order to live.}]
+
+Or if I do have to explicit both theories I should insist on what the predictions of both theories are for our experiments and why they cannot explain people's intuition (the role of time).
+
+\textbf{III}.
+To insist, maybe, on the work of Glymour where our project in part originated: explaining his caveat about treating events as pure propositions.
+To explain how our experiments extend this idea and insist more on the role of changes in continuous time.
+Maybe we should present some insights about how we could formalize further the model?
+How we could use formal tools like Graphical models?
+
+\textbf{IV}. To present further experiments that would be interesting to make (with loops, cases of prevention, etc.)
 

### 2021-01-18 docs/resources/CogSci_paper/paper.tex
reworked figures

@@
 \usepackage{cogsci}
-\usepackage{dsfont}
+%\usepackage{dsfont}
 \usepackage{bm}
@@
 \usepackage{multirow}
+%\usepackage{epsfig}
+%\usepackage{epstopdf}
+%\epstopdfsetup{update=false}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 \begin{center}
-\includegraphics[width=7cm]{intro}
+%\includegraphics[width=7cm]{figures/demo.eps}
+\includegraphics{figures/introfig.eps}
 \end{center}
-\caption{An example of an AND-Gate causal system.
+\caption{(a) An example of an AND-Gate causal system.
 In an AND-Gate for an effect to occur it needs the occurence of all the causes it is linked to.
 Here for the smoke alarm to go off, both the ionization and photoelectric detectors it is connected to have to be activated.
-In this scenario the detectors activation spreads over the top chain first (b) and then over the bottom chain leading to the alarm activation (c).}
+In this scenario the detectors activation spreads over the top chain first (ii) and then over the bottom chain leading to the alarm activation (iii).}
 \label{fig:1}
-\end{figure}
+\end{figure*}
 
@@
 Relying on this principle we can split the system's time frame in units of time that are small enough to have no more than one event\footnote{Again we insist on the definition of \textit{event} as \textit{change of state}.} per unit.
-Lets represent by $\mu^{\emptyset}_{z_{i\rightarrow j}}(t)$ a change of state at time $t$, from variable $Z=z_i$ to $Z=z_j$, $\forall i,j\in \mathds{R}_+$.
+Lets represent by $\mu^{\emptyset}_{z_{i\rightarrow j}}(t)$ a change of state at time $t$, from variable $Z=z_i$ to $Z=z_j$, $\forall i,j\in {R}_+$.
 Here $Z$ is the variable we want to reason about and has no children (represented by the empty set $\{\emptyset\}$).
@@
 
+\nocite{halpernp01}
+
 \bibliographystyle{apacite}
@@
 
-\bibliography{CogSci_Template}
+\bibliography{uber}
 

### 2021-01-19 docs/resources/CogSci_paper/paper.tex
exp section

@@
 \usepackage{multirow}
-%\usepackage{epsfig}
-%\usepackage{epstopdf}
-%\epstopdfsetup{update=false}
 
@@
                    % still generally recommended!
+
+\newcommand{\ev}[2]{$#1_#2$}
 
@@
 In this scenario the detectors activation spreads over the top chain first (ii) and then over the bottom chain leading to the alarm activation (iii).}
-\label{fig:1}
+\label{fig:intro}
 \end{figure*}
@@
 
-\section{Theoretical proposition}
+% Intro:
+% smoke detector example
+%
 
-So far it seems that none of the current main theories of actual causation has focused on the role of temporal information for inferring causal relationship between events qua changes of states over time.
+\section{The continuity account of actual causation}
+
+This section uses the smoke detector example in Figure~\ref{fig:intro}a to introduce our theoretical approach to actual causation.
+Like physical process theory, we focus on processes unfolding in continuous time, but it will be convenient to divide up the dimension of time into a sequence of intervals brief enough that there is at most one event per interval.
+The effect to be explained is an event that happens within one of these intervals: for example, a smoke alarm's changeof state from inactive to active.
+Given this setup, our theory can be formulated using a procedure that starts with the effect and steps backwards through time until the actual cause of the effect is identified.
+
+The immediate cause of the smoke alarm's activation must be an event that took place in the interval preceding the activation.
+In Figure~\ref{fig:intro}a, the event immediately preceding the alarm's activation was the activation of \ev{I}{C}.
+Having identified this immediate cause, we then step backwards and identify the immediate cause of this event, and so on.
+The procedure terminates once we arrive at an event that has no immediate cause within the system of interest,
+and this cause is declared to be the main cause of the effect.  In Figure~\ref{fig:intro}a the main cause is the activation of \ev{R}{C}.
+
+In most cases, the backward-tracing procedure just described will identify a single main cause of the effect.
+If the effect has no immediate cause within the system, then no main cause will be identified.
+Because there is at most one event per interval, the procedure can never identify more than one main cause.
+
+The assumption that at most one event occurs per interval follows from the idea that there are no coincidences.
+In continuous time it is exceedingly improbable that two events would occur at exactly the same time --- in technical terms this kind of coincidence can be described as a ``measure zero'' possibility.
+If there are no coincidences, then slicing the time dimension sufficiently finely will ensure that there is at most one event per interval.
+
+The boundaries of the causal system under consideration will depend on context.
+For simplicity, our discussion of Figure~\ref{fig:intro}a focused only on events internal to the alarm system, and addressed the question of which detector is best viewed as the main cause of the alarm's activation.
+Questions about smoke alarms, however, will often be framed with respect to a causal system that includes both the alarm system and its surroundings.
+Expanding the causal system in this may can allow the alarm's activation to be traced back to an event such as an electrical fault that started a fire which in turn triggered the alarm.
+
+%So far it seems that none of the current main theories of actual causation has focused on the role of temporal information for inferring causal relationship between events qua changes of states over time.
 In contrast we hypothesised that people's causal judgements are influenced not only by the values of all relevant variables at the time the effect occurs, but also the temporal order in which these variables took their values.
@@
 
-To test our hypothesis that the main cause of an effect is the root change that initiates a continuous sequences of changes until the occurrence of the effect, participants were presented animations showing activation spreading over networks of nodes up to the final node.
-We run two experiments which shared the same plot that we wanted as intuitive as possible: participants were told that they were working in a nuclear control room and that their job was to monitor networks of particule detectors.
-The instructions said that when a detector, depicted by a square or circle (see below), absorbs a radioactive particule it becomes active and turns black, transmitting the activation across the links of the network so that an active detector activates the next one in the chain and so forth.
-All the networks include a special component, called 'Gauge of Critical Moment', that becomes active only if all of its input from the detectors it is connected to are active.
-The Gauge of Critical Moment is always represented by a square with ``GCM'' (Experiment 1) or ``G'' (Experiment 2) above.
-At the end of an activation sequence, it is asked to click on the detector(s) they considered as the main cause(s) of the activation of the Gauge of Critical Moment.
-Examples of network activations were shown in the instructions and participants had to answer to answer a survey at the end to see if they correctly understood the instructions.
-If they made a mistake they had to read again all the instructions and answer again the survey.
+We developed two experiments to test the continuity account of actual causation just presented.
+The first experiment explores some simple cases directly analogous to the smoke detector example.
+The second experiment explores cases that elaborate that distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
 
-In both experiments we included only two types of input-output logic, namely single input (chains) and AND-Gate (branches where the effect needs the activation of two inputs to occur).
-When people were presented chains, we first wanted to see if they identified the main cause of the effect with a change of state that occurred in the chain the furthest away from effect (\textit{root change}) or the closest to it (\textit{immediate change}).
-When they were presented AND-Gates, we wanted to see if they identified the main cause of the effect with the root change that occurred chronologically first (\textit{1\textsuperscript{st} root change}, labeled ``R1'' henceforth), the root change that occurred chronologically last (\textit{2\textsuperscript{nd} root change}, ``R2'' henceforth), the immediate change that occurred chronologically first (\textit{1\textsuperscript{st} immediate change}, ``I1'') or the immediate change that occurred chronologically last (\textit{2\textsuperscript{nd} immediate change}, ``I1'').
+The cover story for both experiments asked participants to imagine that they worked in a control room and that their job was to monitor networks of particle detectors.
+Detectors were shown as white squares (as in Figure~\ref{fig:intro} that become active and turn black when they absorb a radioactive particle.
+Activation is transmitted across links in the network, and a detector activates if all of its input detectors are active.
+Each network contains a special detector called the Gauge of Critical Moment (GCM, analogous to the smoke alarm in Figure~\ref{fig:intro}a) and participants were told that ``for each activation sequence that you see, your job is to decide what caused the activation of the GCM.'' The prompt after each sequence was ``In this sequence what caused the activation of the GCM?
+Respond by clicking on a detector,'' and participants were required to choose a single detector in response.
+
+Both experiments included chains in which the GCM received input from one detector, and dual branch networks such as Figure~\ref{fig:intro}a in which the GCM received input from two detectors.
+For chains, the continuity account predicts that participants will tend to choose the root cause of the effect (ie the detector that initiates the activation sequence) rather than the immediate cause (the detector that immediately precedes the GCM).
+For dual branch networks, the continuity account predicts that participants will tend to choose the root cause on the branch whose activation is temporally continuous with the activation of the GCM.
+We refer to this branch as the continuous branch, and refer to the root cause and the immediate cause on this branch as \ev{R}{C} and \ev{I}{C} respectively, where the subscript denotes ``continuous.'' We refer to the other branch as the ``delayed branch,'' and use \ev{R}{D} and \ev{I}{D} for the root and immediate cause on this branch.
+The ``delay'' in this naming scheme refers to the delay that occurs between the activation of \ev{I}{D} and the activation of the GCM.
 
 \subsection{Experiment 1}
+
+The purpose
 
@@
 
-\textbf{Participants}.
-The experiment was hosted by Google Cloud App Engine and 30 participants were recruited via Amazon Mahchanical Turk.
+\textbf{Participants}. 30 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment.
+
+%The experiment was hosted by Google Cloud App Engine and 30 participants were recruited via Amazon Mahchanical Turk.
 There were 10 females (average age: 37.3) and 20 males (average age: 37.6).
@@
 \textbf{Method}.
+
+% Examples of network activations were shown in the instructions and participants had to answer to answer a survey at the end to see if they correctly understood the instructions.
+If they made a mistake they had to read again all the instructions and answer again the survey.
+
 Each participant was presented 15 different networks of square detectors.
@@
 \begin{center}
-\includegraphics[width=8cm]{results_E1}
+%\includegraphics[width=8cm]{results_E1}
+\includegraphics{figures/expt1.eps}
 \end{center}
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=8cm]{stim_E2a}
-\end{center}
-\caption{An example of an activation spreading over a network of Experiment 2.
+%\begin{figure}[ht]
+%\begin{center}
+%\includegraphics[width=8cm]{stim_E2a}
+%\end{center}
+%\caption{An example of an activation spreading over a network of Experiment 2.
 Here a continuous sequence of changes links R1 and I2 -- in other words this is the sequence of activations initiated by the chronologically first root change (curved branch) that eventually leads to the activation of the last immediate change and therefore G.}
-\label{fig:3}
-\end{figure}
+%\label{fig:3}
+%\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=8cm]{results_E2}
+%\includegraphics[width=8cm]{results_E2}
+\includegraphics{figures/expt2.eps}
 \end{center}

### 2021-01-20 docs/resources/CogSci_paper/paper.tex
edited expt 1

@@
 We developed two experiments to test the continuity account of actual causation just presented.  
-The first experiment explores some simple cases directly analogous to the smoke detector example.
+%The first experiment explores some simple cases directly analogous to the smoke detector example.
 The second experiment explores cases that elaborate that distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
-
 The cover story for both experiments asked participants to imagine that they worked in a control room and that their job was to monitor networks of particle detectors.
@@
 Activation is transmitted across links in the network, and a detector activates if all of its input detectors are active.
-Each network contains a special detector called the Gauge of Critical Moment (GCM, analogous to the smoke alarm in Figure~\ref{fig:intro}a) and participants were told that ``for each activation sequence that you see, your job is to decide what caused the activation of the GCM.'' The prompt after each sequence was ``In this sequence what caused the activation of the GCM?
-Respond by clicking on a detector,'' and participants were required to choose a single detector in response.
+Each network contains a special detector called the Gauge of Critical Moment (GCM, analogous to the smoke alarm in Figure~\ref{fig:intro}a) and participants were told that ``for each activation sequence that you see, your job is to decide what caused the activation of the GCM.''
 
@@
 For dual branch networks, the continuity account predicts that participants will tend to choose the root cause on the branch whose activation is temporally continuous with the activation of the GCM.
-We refer to this branch as the continuous branch, and refer to the root cause and the immediate cause on this branch as \ev{R}{C} and \ev{I}{C} respectively, where the subscript denotes ``continuous.'' We refer to the other branch as the ``delayed branch,'' and use \ev{R}{D} and \ev{I}{D} for the root and immediate cause on this branch.
-The ``delay'' in this naming scheme refers to the delay that occurs between the activation of \ev{I}{D} and the activation of the GCM.
+We refer to this branch as the continuous branch, and refer to the root cause and the immediate cause on this branch as \ev{R}{C} and \ev{I}{C} respectively, where the subscript denotes ``continuous.'' We refer to the other branch as the ``delayed branch,'' and use \ev{R}{D} and \ev{I}{D} for the root and immediate causes on this branch.
+The ``delay'' in this naming scheme refers to the delay that occurs between \ev{I}{D} and the activation of the GCM.
+It is convenient to use the same labels for both detectors and events: for example, \ev{R}{C} will be used to denote both a detector on the continuous branch and the activation of this detector.
+This flexible use of notation, however, should not be taken to suggest that actual causation involves a relationship between objects instead of a relationship between events.
+
+Our presentation of the experiments focuses on two theoretical accounts: the continuity account and the probability raising account.
+For completeness, though, we first discuss the counterfactual and physical process accounts.
+Both accounts can be formulated in different ways --- here we present what we take to be the most natural predictions of the two accounts, and the General Discussion considers ways in which the accounts can be adjusted to make different predictions.
+
+The counterfactual account makes no clear prediction about whether root causes or immediate causes should take priority.
+For a chain network, if the root cause had not occurred, the effect would not have occurred, but it is equally true that if the immediate cause had not occurred, the effect would not have occurred.
+If we consider dual branch networks and restrict attention to root causes only, then the account makes no clear preediction about whether \ev{R}{C} should be preferred to \ev{R}{D}.
+If \ev{R}{C} had not occurred, then the effect would not have occurred, and likewise for \ev{R}{D}.
+
+The physical process account is similarly inconclusive.
+Both root and immediate causes are connected by a physical process to the effect, and there seem to be no clear grounds for preferring one to the another.
+If we focus only on root causes, again there is no clear preference between \ev{R}{C} and \ev{R}{D}.
+After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM, but this unbroken stream of activation only ``unlocks`` the detector once \ev{I}{C} is also active.
+It follows that both \ev{R}{C} and \ev{R}{D} are connected to the effect by physical processes.
+
+Because both of these accounts make no clear predictions, it seemed possible that people's inferences about the networks in our experiments would be highly variable and would reveal no clear trends.
+Our first experiment therefore explored simple cases analogous to the smoke detector example in Figure~\ref{fig:intro} with the goal of establishing whether the basic experimental procedure was viable.
+The second experiment focused on more elaborate dual-branch cases that aimed to distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
 
@@
 
-The purpose
-
-The objective of this experiment was to see whether people's causal judgement were influenced by: the length of the sequences of changes (for chains and AND-Gates), and -- for AND-Gates only -- the delay between the last activation of the first branch (I1) and the first activation of the second branch (R2).
-In this experiment activation starts to spread over the second branch after the activation reached the end of the first branch (like in Fig.\ref{fig:1}).
+Experiment 1 included both causal chains and dual branch networks.
+For all of the dual branch stimuli, \ev{I}{D} occurred before \ev{R}{C} (Figure~\ref{fig:expt1}a) which means that the delayed branch completed its activation before the continuous branch began to activate.
+The probability raising and continuity accounts both predict that root causes should be preferred to immediate causes, and that for dual branch networks \ev{R}{C} should be preferred to \ev{R}{D}.
+The primary purpose of the experiment was to test both predictions.
 
@@
 
-\textbf{Method}.
+\textbf{Materials}.
+The experiment used a customized interface built using the jsPsych library~\cite{XXX}.
+For all networks presented, participants clicked a ``Run'' button to observe an activation sequence.
+The first event in the sequence (i.e.\ the first change of state) always took place 5 seconds after the Run button was clicked, and the delay between successive activations along a chain of detectors was set to 100 ms.
+After the final event in a sequence (i.e.\ the activation of the GCM), all detectors became clickable after a delay of 1 second.
+Clicking on a detector turned its border red, and at most one detector could be selected at any time.
+After a sequence completed, participants could view it again if they wished by clicking a ``Run again'' button.
 
-% Examples of network activations were shown in the instructions and participants had to answer to answer a survey at the end to see if they correctly understood the instructions.
-If they made a mistake they had to read again all the instructions and answer again the survey.
+\textbf{Design}.
+The experiment included activation sequences over 15 networks (3 chains and 12 dual branch networks).
+Excluding the GCM, each chain and each branch of each dual network had 7 detectors.
+We refer to the sequences as \emph{short}, \emph{medium} or \emph{long} based on the distance between the root causes and the GCM.
+Excluding the GCM, the short, medium and long chain sequences showed 1, 3 and 6 active detectors respectively at the end of the sequence.
+Within each dual branch sequence, the root causes on the two branches (\ev{R}{D} and \ev{R}{C}) were equidistant from the GCM, but these distances varied across sequences.
+The sequence in Figure~\ref{fig:intro}a is a medium dual branch sequence (3 active detectors in each branch excluding the GCM), and short and long versions had 1 and 6 active detectors respectively per branch.
 
-Each participant was presented 15 different networks of square detectors.
-Square detectors maintain their activation through time.
-Participants were first presented the network in its initial and static state and were asked to click the ``Run'' button to observe an activation sequence over the network.
-They waited 5000ms to see a first change of state occurring in a detector and the activation delay between any two successive detectors was set on 100ms.
+Each dual branch sequence (short, medium and long) came in four versions.
+The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} \ev{I}{D}) as active from the very beginning of the sequence.
+The activation of this branch was therefore presented as having occurred at some indefinite time in the past, resulting in a steady state of activation.
+The three \emph{event} versions all showed \ev{R}{D} activating 5 seconds into the sequence, and had delays of 2, 4 and 6 seconds between the activation of \ev{I}{D} and \ev{R}{C}.
 
-Three stimuli were chains that differed in length of activation sequence, that is including 2, 4 or 7 activated squares (with the GCM).
-These lengths were labeled ``Short'', ``Medium'' and ``Long''.
-The AND-Gates were similarily categorised ``Short'', ``Medium'' (the one in Fig.\ref{fig:1}) and ``Long'' with both branches being the same length.
-In each length category participants were shown three similar AND-Gates that differed uniquely in the delays between I1 and R2.
-The delays were 2s, 4s and 6s.
+\textbf{Procedure}.
 
-All the stimuli were presented in a random order.
-Two groups, ``Left'' and ``Right'', were designed and partcipants were randomly assigned to either group at the begining of the experiment.
-In the ``Left'' group all the networks were shifted towards left (the GCM being on the left) and in the ``Right'' group all the networks were shifted towards right.
-For each AND-Gate participants saw randomly either the top branch or the bottom branch activating first.
+Participants first read a set of instructions which introduced the task and included examples of a chain and a dual branch network.
+They then answered three questions about the task and the detectors, and were sent back to read the instructions again if they answered incorrectly.
+They continued cycling through the instructions and the test questions until they answered all questions correctly.
 
-After the activation of the GCM (the end of the animation) participants had to wait 1000ms before the squares became clickable and the following intruction appeared: ``In this sequence what caused the activation of the GCM?
-Respond by clicking on a detector''.
-They had the option to run again the animation (no more than 9 times) or to go to the next network assuming they clicked on one detector -- they couldn't select more than one detector.
-When selected the edges of a detector turned red.
+The 15 activation sequences in the experiment proper were presented in random order.
+For dual branch sequences, the position of the delayed branch was also randomized (Figure~\ref{fig:intro} shows the delayed branch on the top rather than the bottom).
+The orientation of the network (GCM on the left or the right) was randomized within participants.
+The prompt after each sequence was ``In this sequence what caused the activation of the GCM?
+Respond by clicking on a detector,'' and participants were required to choose a single detector in response.
 
 \textbf{Results}.
-First of all we observed no significant effect [STATISTICS?] of neither the orientation of the networks (left \textit{vs} right) nor the location where the activation starts first (top \textit{vs} bottom).
+
+Network orientation (left or right) and position of the delay branch (top or bottom) had no apparent effect, and we therefore collapse across these variables.
+Figure~\ref{fig:expt1} summarizes the results for short, medium and long dual-branch sequences.
+The delay between \ev{I}{D} and \ev{R}{C} had no significant effect, and Figure~\ref{fig:expt1} combines results for all three delays.
+
+Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the main cause.
+This result can be separated into two general conclusions.
+First,  participants were more likely to choose root causes than immediate causes.
+Root and immediate causes were identical for the short sequences and therefore cannot bet distinguished, but the results for medium and long sequences reveal a preference for root causes.
+STATISTICAL ANALYSIS.
+The data for chains (not shown in Figure~\ref{fig:expt1}) provide further evidence or a preference for root causes.
+NUMBERS HERE
+
+The second conclusion is that for dual branch sequences, the root cause on the continuous branch (\ev{R}{C}) is preferred to the root cause on the delayed branch (\ev{R}{D}).
+This result is consistent with the continuity account but also consistent with the probability raising account, because \ev{R}{C} guarantees the occurrence of the effect but \ev{R}{D} does not.
+
+Figure \ref{fig:intro} suggests that event and state sequences lead to similar patterns of responses.
+The continuity account makes the same prediction for both kinds of sequences, but we thought that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}.
+If this effect exists, our data suggest that it is relatively small.
+
+It is also notable that people's judgments are largely unaffected by the delay between \ev{I}{D} and \ev{R}{C} and the activation length of each sequence (short, medium or long).
+Both of these variables affect the time that elapses between the root causes and the effect, and our results suggest that people's judgments are not exquisitely sensitive to this kind of variation.
+
+%First of all we observed no significant effect [STATISTICS?] of neither the orientation of the networks (left \textit{vs} right) nor the location where the activation starts first (top \textit{vs} bottom).
 On average participants took 14 min to complete the experiment and overall 96\% of the time participants run an animation only once.
@@
 
-[WARNING: THE GRAPH BELOW SHOW THE RESULTS FOR SHORT, MEDIUM, LONG \textbf{LENGTHS} (ALL THE DELAYS MIXED) SO NOT THE DIFFERENT DELAYS BETWEEN I1 AND R2.
+%[WARNING: THE GRAPH BELOW SHOW THE RESULTS FOR SHORT, MEDIUM, LONG \textbf{LENGTHS} (ALL THE DELAYS MIXED) SO NOT THE DIFFERENT DELAYS BETWEEN I1 AND R2.
 BUT AS OUR FOCUS IS ON TEMPORAL INFORMATION, MAYBE WE SHOULD GRAPHICALLY SHOW THAT THE DELAY HAS NO EFFECT??
@@
 \end{center}
-\caption{Comparison of the results for different length of activation in AND-Gates.
+\caption{Experiment 1.
+Comparison of the results for different length of activation in AND-Gates.
 Short: 1 activated detector in each branch.

### 2021-01-20 docs/resources/CogSci_paper/paper.tex
pass through expt2

@@
 
-\begin{figure*}[h]
+\begin{figure*}
 \begin{center}
@@
 This section uses the smoke detector example in Figure~\ref{fig:intro}a to introduce our theoretical approach to actual causation.
-Like physical process theory, we focus on processes unfolding in continuous time, but it will be convenient to divide up the dimension of time into a sequence of intervals brief enough that there is at most one event per interval.
-The effect to be explained is an event that happens within one of these intervals: for example, a smoke alarm's changeof state from inactive to active.
+Like physical process theory, we focus on processes unfolding in continuous time, but it will be convenient to divide up the temporal dimension into intervals brief enough that there is at most one event per interval.
+The effect to be explained is an event that happens within one of these intervals: for example, a smoke alarm's change of state from inactive to active.
 Given this setup, our theory can be formulated using a procedure that starts with the effect and steps backwards through time until the actual cause of the effect is identified.
@@
 Questions about smoke alarms, however, will often be framed with respect to a causal system that includes both the alarm system and its surroundings.
-Expanding the causal system in this may can allow the alarm's activation to be traced back to an event such as an electrical fault that started a fire which in turn triggered the alarm.
+Expanding the causal system in this way may allow the alarm's activation to be traced back to an event such as an electrical fault that started a fire that generated smoke that triggered the alarm.
 
@@
 The second experiment explores cases that elaborate that distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
-The cover story for both experiments asked participants to imagine that they worked in a control room and that their job was to monitor networks of particle detectors.
-Detectors were shown as white squares (as in Figure~\ref{fig:intro} that become active and turn black when they absorb a radioactive particle.
+The cover story asked participants to imagine that they worked in a nuclear control room and that their job was to monitor networks of particle detectors.
+Detectors were shown as white squares (as in Figure~\ref{fig:intro}) that become active and turn black when they absorb a radioactive particle.
 Activation is transmitted across links in the network, and a detector activates if all of its input detectors are active.
@@
 It is convenient to use the same labels for both detectors and events: for example, \ev{R}{C} will be used to denote both a detector on the continuous branch and the activation of this detector.
-This flexible use of notation, however, should not be taken to suggest that actual causation involves a relationship between objects instead of a relationship between events.
+This flexible use of notation, however, does not imply that actual causation involves a relationship between objects rather than events.
 
@@
 For completeness, though, we first discuss the counterfactual and physical process accounts.
-Both accounts can be formulated in different ways --- here we present what we take to be the most natural predictions of the two accounts, and the General Discussion considers ways in which the accounts can be adjusted to make different predictions.
+Both accounts can be formulated in different ways.
+Here we present what we take to be the default predictions of the two accounts, and the General Discussion considers ways in which the accounts can be adjusted to make different predictions.
 
 The counterfactual account makes no clear prediction about whether root causes or immediate causes should take priority.
-For a chain network, if the root cause had not occurred, the effect would not have occurred, but it is equally true that if the immediate cause had not occurred, the effect would not have occurred.
-If we consider dual branch networks and restrict attention to root causes only, then the account makes no clear preediction about whether \ev{R}{C} should be preferred to \ev{R}{D}.
+For a chain network, if the root cause had not occurred, the effect would not have occurred, but if the immediate cause had not occurred, the effect would also not have occurred.
+If we consider dual branch networks and restrict attention to root causes only, then the account makes no clear prediction about whether \ev{R}{C} should be preferred to \ev{R}{D}.
 If \ev{R}{C} had not occurred, then the effect would not have occurred, and likewise for \ev{R}{D}.
@@
 The physical process account is similarly inconclusive.
-Both root and immediate causes are connected by a physical process to the effect, and there seem to be no clear grounds for preferring one to the another.
+Both root and immediate causes are connected by a physical process to the effect, and there seems to be no clear reason for preferring one to the another.
 If we focus only on root causes, again there is no clear preference between \ev{R}{C} and \ev{R}{D}.
-After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM, but this unbroken stream of activation only ``unlocks`` the detector once \ev{I}{C} is also active.
+After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM which only ``unlocks`` the detector once \ev{I}{C} is also active.
 It follows that both \ev{R}{C} and \ev{R}{D} are connected to the effect by physical processes.
@@
 Each dual branch sequence (short, medium and long) came in four versions.
-The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} \ev{I}{D}) as active from the very beginning of the sequence.
+The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} and \ev{I}{D}) as active from the very beginning of the sequence.
 The activation of this branch was therefore presented as having occurred at some indefinite time in the past, resulting in a steady state of activation.
@@
 \textbf{Procedure}. 
-
-Participants first read a set of instructions which introduced the task and included examples of a chain and a dual branch network.
+Participants first read instructions which introduced the task and included examples of a chain and a dual branch network.
 They then answered three questions about the task and the detectors, and were sent back to read the instructions again if they answered incorrectly.
@@
 The 15 activation sequences in the experiment proper were presented in random order.
-For dual branch sequences, the position of the delayed branch was also randomized (Figure~\ref{fig:intro} shows the delayed branch on the top rather than the bottom).
+For dual branch sequences, the vertical position of the delayed branch was also randomized (Figure~\ref{fig:intro} shows the delayed branch on the top rather than the bottom).
 The orientation of the network (GCM on the left or the right) was randomized within participants.
@@
 \textbf{Results}. 
-
 Network orientation (left or right) and position of the delay branch (top or bottom) had no apparent effect, and we therefore collapse across these variables.
@@
 STATISTICAL ANALYSIS.
-The data for chains (not shown in Figure~\ref{fig:expt1}) provide further evidence or a preference for root causes.
-NUMBERS HERE
 
@@
 Figure \ref{fig:intro} suggests that event and state sequences lead to similar patterns of responses.
-The continuity account makes the same prediction for both kinds of sequences, but we thought that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}.
+The continuity account makes the same prediction for both kinds of sequences, but we anticipated that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}.
 If this effect exists, our data suggest that it is relatively small.
 
-It is also notable that people's judgments are largely unaffected by the delay between \ev{I}{D} and \ev{R}{C} and the activation length of each sequence (short, medium or long).
-Both of these variables affect the time that elapses between the root causes and the effect, and our results suggest that people's judgments are not exquisitely sensitive to this kind of variation.
+It is notable that people's judgments are largely unaffected by the delay between \ev{I}{D} and \ev{R}{C} and the activation length of each sequence (short, medium or long).
+Both of these variables affect the time that elapses between the root causes and the effect, and our results suggest that people's judgments are not exquisitely sensitive to this sort of variation.
+
+Thus far we have focused on dual branch sequences, and the chain sequences are of interest primarily because they provide additional evidence of a preference for root causes.  NUMBERS HERE
 
@@
 
-\begin{figure}[ht]
+\begin{figure}
 \begin{center}
@@
 \caption{Experiment 1.
-Comparison of the results for different length of activation in AND-Gates.
-Short: 1 activated detector in each branch.
-Medium: 3 activated detectors in each branch.
-Long: 6 activated detectors in each branch.
-``State'' responses correspond to the networks where one branch was already activated before participants run the animation.
-``Event'' responses correspond to the networks where no detectors were activated before and include results for 2s, 4s and 6s delays between I1 and R2.
-BUT SEE WARNING ABOVE.}
-\label{fig:2}
+(a) Order of the four key events for dual branch sequences.
+The probability raising and continuity accounts both single out \ev{R}{C} as the main cause.
+(b)  Responses for dual branch sequences.
+The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence.}
+\label{fig:expt1}
 \end{figure}
@@
 
-In the previous experiment we have seen that people generally chose R2 (identical to I2 in the short sequences) as the main the cause of the activation of the GCM.
-These results make sense since R2 and I2 are linked by a continuous sequence of changes.
-But this set up, where the first element of a chain ``waits'' for the occurrence of the last event of another chain before it can initiate a second sequence, is probably not the most common scenario in real life.
-For instance it might take more time for one causal process to end so that the causal process that has occurred first is not necessarily the one that ends first.
-In Experiment 2 we therefore wanted to break the temporal continuity between the root changes and the immediate changes of Experiment 1 such that the continuous sequence of changes be sometimes between R1 and I2 or R2 and I2.
-We predicted that when there is a continuous sequence of changes between R1 and I2, people would preferably choose R1 as the main cause.
+The results of Experiment 1 are broadly consistent with both the continuity and probability raising accounts, and the goal of Experiment 2 was to distinguish between these accounts.
+The two make different predictions for dual branch sequences in which the delayed branch activates after activation has already begun along the continuous branch.
+One such sequence is shown in Figure~\ref{fig:intro}b.
+In cases like this, \ev{R}{D} guarantees the occurrence of the effect, and therefore qualifies as the primary cause according to the probability raising account.
+The continuity account, however, still treats \ev{R}{C} as the primary cause.
+
+The materials and procedure for Experiment 2 are similar to those for Experiment 1, and we highlight only the few points of difference.
 
 \textbf{Participants}.
-The experiment was hosted by Google App Engine and 98 participants were recruited via Amazon Mechanical Turk.
+100 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment.
+Two did not submit a valid completion code, leaving 98 sets of responses for analysis.
+
+%The experiment was hosted by Google App Engine and 98 participants were recruited via Amazon Mechanical Turk.
 There were 40 females (average age: 40.4), 57 males (average age: 37.6) and 1 declared ``other''.
@@
 
-\textbf{Method}.
-In Experiment 2 each participant was presented 7 different networks.
+\textbf{Design}. 
+The experiment included one chain sequence and 6 dual branch sequences.
+Each dual branch included one straight branch and a longer curved branch, as shown in Figure~\ref{fig:intro}b.
+The dual branch sequences included the events \ev{R}{C}, \ev{R}{D}, \ev{I}{C}, and \ev{I}{D} in three different orders shown in Figures~\ref{fig:expt2}a and \ref{fig:expt2}b.
+The sequence in Figure~\ref{fig:intro}b is an instance of the order in Figure~\ref{fig:expt2}a, because the activation on the delayed branch starts (\ev{R}{D}) and finishes (\ev{I}{D}) while activation is propagating along the continuous branch.
+The 6 dual branch sequences included 2 variants of each of the three orders.
+In one variant \ev{R}{C} belonged to the curved branch (as in Figure~\ref{fig:intro}b) and in the other \ev{R}{C} belonged to the straight branch.
+
+% XXX: event vs state?
+
+\textbf{Procedure}. 
+The position of the curved branch (top or bottom) was randomized. The presentation order of the activation sequences and the orientation (GCM on the left or right) were randomized as for Experiment 1.
+%In Experiment 2 each participant was presented 7 different networks.
 One stimulus was a chain and all the other six stimuli were AND-Gates.
@@
 
-As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network.
+%As in Experiment 1 participants were first presented the network in its initial state and were asked to click the ``Run'' button to observe an activation sequence over the network.
 Participants waited 5000ms to see a first change of state occurring in a detector.
@@
 \textbf{Results}.
-First we observed no significant effect (statistics?) of neither the orientation of the networks (left \textit{vs} right) nor the location of the curved chain (top \textit{vs} bottom).
+Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.
+Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b.
+Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C},\ev{I}{C}\}, and a similar (but weaker) effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\} sequences.
+The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D},\ev{I}{C}\} sequences, even though the probability raising account makes the opposite prediction.
+STATISTICS HERE.
+
+Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2.
+This difference may reflect the increased difficulty of Experiment 2.
+When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy fo choosing the immediate cause that directly precedes the effect.
+
+%First we observed no significant effect (statistics?) of neither the orientation of the networks (left \textit{vs} right) nor the location of the curved chain (top \textit{vs} bottom).
 On average participants took 8 min to complete the experiment and overall 91\% of the time participants run an animation only once.
@@
 
-\begin{figure}[H]
+\begin{figure}
 \begin{center}
@@
 \end{center}
-\caption{Results for AND-Gates where the continuous sequence of changes is either between R1 and I2 or between R2 and I2.
-``Event (R1 cont)'': responses for the networks where no detectors were already active before the participant runs the animation and where the continuous sequence of changes is between R1 and I2.
-``Event (R2 cont)'': responses for the networks where no detectors were already active and where the continuous sequence is between R2 and I2.
-``State (R2cont)'': responses for networks where one branch was already activated before the participant runs the animation and where the continuous sequence of activation is necessarily initiated by R2.}
-\label{fig:4}
+\caption{Experiment 2.
+(a) The probability raising and continuity accounts make different predictions for dual branch sequences in which \ev{R}{C} occurs before \ev{R}{D}.
+(b) The two accounts agree when \ev{R}{C} occurs after \ev{R}{D}.
+(c) Responses for the three possible orders of the key events.} 
+\label{fig:expt2}
 \end{figure}
 
-\newpage
 \section{General discussion}

### 2021-01-20 docs/resources/CogSci_paper/paper.tex
fig 1 caption

@@
 
+%\title{A continuity-based account of actual causation}
+%\title{Identifying actual causes by tracking Changes of State in Continuous Time}
 \title{Actual Causes as Changes of State in Continuous Time}
@@
 \end{center}
-\caption{(a) An example of an AND-Gate causal system.
-In an AND-Gate for an effect to occur it needs the occurence of all the causes it is linked to.
-Here for the smoke alarm to go off, both the ionization and photoelectric detectors it is connected to have to be activated.
-In this scenario the detectors activation spreads over the top chain first (ii) and then over the bottom chain leading to the alarm activation (iii).}
+\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch).
+All components are initially inactive (ii) Detector \ev{R}{D} activates and activity propagates along the top branch.
+(iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the alarm, which activates only when both  \ev{I}{D} and \ev{I}{C} are active.
+(b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the smoke alarm.}
 \label{fig:intro}

### 2021-01-21 docs/resources/CogSci_paper/paper.tex
started working through intro

@@
 \section{Introduction}
+
+Around 5:27 pm on November 9, 1965, Martin Saltzman found himself trapped in a dark elevator roughly a quarter of a way up the Empire State building~\cite{rosenthalg65}.
+The stopping of Saltzman's elevator was one of thousands of unexpected events that occurred that evening, and virtually all of the people involved must have wondered about the causes of these events.
+Subsequent accounts of the Great Northeastern Blackout often trace these events back to the tripping of a relay on line Q29BD in Ontario that triggered a cascade of failures.
+
+Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}.
+Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{}, force dynamics \cite{wolff} and physical simulation~\cite{gerstenberg}.
+Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.
+On this account, the tripping of relay Q29BD is the main cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.
+
+Most accounts of actual causation are consistent with one of two broad views of causation: the counterfactual approach and the physical process approach.
+The counterfactual approach suggests that the tripping of relay Q29BD is a cause of the elevator stopping because if the relay had not tripped then the elevator would not have stopped.
+The physical process view suggests that the relay tripping is a cause of the elevator stopping because the two are linked by a physical process involving the transmission of force or energy.
+Both accounts need additional machinery in order to specify which among the many causes of an event is singled out as the main cause.
+Our continuity account fits most naturally with the physical process approach, and can be viewed as an attempt to bring out some implications of this approach for actual causation.
+
+Two key principles.
+
+Limited in scope.
+
+Like most other formulations of the physical process view, our
 

### 2021-01-21 docs/resources/CogSci_paper/paper.tex
pass through introduction

@@
-% 
-% Annual Cognitive Science Conference
-% Sample LaTeX Paper -- Proceedings Format
-%
-
-% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
-% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
-% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
-% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
-% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
-% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
-% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
-% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
-% Modified : Ken Forbus                              01/23/2004
-% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
-% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
-% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
-% Modified : Roger Levy (rplevy@mit.edu)     12/31/2018
-
 %% Change "letterpaper" in the following line to "a4paper" if you must.
@@
 
-Around 5:27 pm on November 9, 1965, Martin Saltzman found himself trapped in a dark elevator roughly a quarter of a way up the Empire State building~\cite{rosenthalg65}.
+Around 5:27 pm on November 9, 1965, Martin Saltzman found himself trapped in a dark elevator roughly a quarter of a way up the Empire State building~\cite{gelbr1965night}.
 The stopping of Saltzman's elevator was one of thousands of unexpected events that occurred that evening, and virtually all of the people involved must have wondered about the causes of these events.
@@
 
-Two key principles.
+The continuity account relies on two core principles.
+First, it assumes that the effect to be explained and its cause are both changes of state.
+The tripping of a relay qualifies as a candidate cause, but the steady-state setting of a relay does not.
+Our emphasis on state changes is consistent with the common view that causal relationships are relationships between events rather than facts or states of affairs, but  \citeA{glymourstone} point out that most Bayes net accounts of actual causation ignore changes of state.
+State changes are embedded in continuous time, and a change that begins at a specific moment is typically the direct result of an event that occurred an instant before.
+Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause.
+Some previous work on actual causation highlights the importance of time~\cite{youngs09,stephanmw20e}, but to our knowledge previous work has not explored the implications of temporal continuity in the way we do here.
 
-Limited in scope.
+The continuity account does not aspire to capture all of people's intuitions about actual causation.
+Like the physical process approach more broadly, it is most applicable to judgments about physical rather than social systems, and does not capture cases of causation by omission~\cite{wolffbh10}.
+To us it seems likely that judgments of actual causation rely on multiple principles that resist unification under a single heading~\cite{hall04,danks17}.
+In focusing on physical systems we aim to characterize one paradigmatic class of judgments but acknowledge that additional approaches are needed to understand actual causation in other settings.
 
-Like most other formulations of the physical process view, our
+The following sections introduce the continuity account of actual causation and present two experiments designed to test core predictions of the account.
+Among previous theories of actual causation the most natural comparison is Spellman's probability raising account, which proposes that the actual cause of an effect is the cause that increased its probability to the greatest extent~\cite{spellman97}.
+Our second experiment directly compares the continuity account with the probability raising account and we find that the continuity account provides the better account of our data.
 
-Identifying the causes of an event involves not only logical rules but also different senses of ``causation'' that people might have.
+%Identifying the causes of an event involves not only logical rules but also different senses of ``causation'' that people might have.
 Let's imagine that, to mitigate false alarms rate and improve the detection of fires, Lilly decides to install at home a dual alarm system that goes off if and only if two chains of ionization and photoelectric detectors are triggered (Fig.\ref{fig:1}a).
@@
 
-What would Lilly think caused the alarm to go off?
+%What would Lilly think caused the alarm to go off?
 Several answers can be considered.
@@
 
-The alarm case can be seen as an example of ``singular causation''\footnote{Also called ``token'' or ``actual causation'' by opposition with ``general causation'' -- following a distinction that has been frequently made in the litterature.} where we want to understand, in a singular context, what caused a particular event to occurr at the time it did.
+%The alarm case can be seen as an example of ``singular causation''\footnote{Also called ``token'' or ``actual causation'' by opposition with ``general causation'' -- following a distinction that has been frequently made in the litterature.} where we want to understand, in a singular context, what caused a particular event to occurr at the time it did.
 Here specifically it instantiates a class of causal systems where an event is a common effect of distinct sequences of necessary causes.
@@
 
-In this paper we present two experiments that show the importance of taking into account temporal information to explain the causal judgements people make.
+%In this paper we present two experiments that show the importance of taking into account temporal information to explain the causal judgements people make.
 In particular we relied on the intuition behind the example above that the chronological order and the temporal relationship between distinct causes can bias people toward an event rather than another in considering which one is the main cause of the effect.
@@
 
-This section uses the smoke detector example in Figure~\ref{fig:intro}a to introduce our theoretical approach to actual causation.
-Like physical process theory, we focus on processes unfolding in continuous time, but it will be convenient to divide up the temporal dimension into intervals brief enough that there is at most one event per interval.
-The effect to be explained is an event that happens within one of these intervals: for example, a smoke alarm's change of state from inactive to active.
+We introduce the continuity account using a scenario similar to the cover story used in the experiments.
+Figure \ref{fig:intro}a shows a network of particle detectors (white squares), including a special detector called the Gauge of Critical Moment (GCM).
+The detectors activate and turn black when they absorb a radioactive particle.
+Activation is transmitted across links in the network, and a detector activates if all of its input detectors are active.
+
+Although we focus on processes unfolding in continuous time, it will be convenient to divide up the temporal dimension into intervals brief enough that there is at most one event per interval.
+The effect to be explained is an event that happens within one of these intervals: for example, the GCM's change of state from inactive to active.
 Given this setup, our theory can be formulated using a procedure that starts with the effect and steps backwards through time until the actual cause of the effect is identified.
 
-The immediate cause of the smoke alarm's activation must be an event that took place in the interval preceding the activation.
-In Figure~\ref{fig:intro}a, the event immediately preceding the alarm's activation was the activation of \ev{I}{C}.
+The immediate cause of the GCM's activation must be an event that took place in the interval preceding the activation.
+In Figure~\ref{fig:intro}a, the event immediately preceding the GCM's activation was the activation of \ev{I}{C}.
 Having identified this immediate cause, we then step backwards and identify the immediate cause of this event, and so on.
@@
 
-The assumption that at most one event occurs per interval follows from the idea that there are no coincidences.
-In continuous time it is exceedingly improbable that two events would occur at exactly the same time --- in technical terms this kind of coincidence can be described as a ``measure zero'' possibility.
+The assumption that at most one event occurs per interval follows from the idea that there can be no coincidences in continuous time.
+If the temporal dimension is continuous it is exceedingly improbable that two events would occur at precisely the same time --- in technical terms this kind of coincidence can be described as a ``measure zero'' possibility.
 If there are no coincidences, then slicing the time dimension sufficiently finely will ensure that there is at most one event per interval.
@@
 The boundaries of the causal system under consideration will depend on context.
-For simplicity, our discussion of Figure~\ref{fig:intro}a focused only on events internal to the alarm system, and addressed the question of which detector is best viewed as the main cause of the alarm's activation.
+For simplicity, our discussion of Figure~\ref{fig:intro}a has focused only on events internal to the particle detection network, and has attemped to characterize which of these events is best viewed as the main cause of the GCM's activation.
+The particle detection network, however, could also be considered part of a broader causal system that includes both the network and the network's surroundings.
+For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom will probably be identified as the immediate cause of \ev{R}{C} and the main cause of the GCM's activation.
+
+%addressed the question of which detector is best viewed as the main cause of the alarm's activation.
 Questions about smoke alarms, however, will often be framed with respect to a causal system that includes both the alarm system and its surroundings.
@@
 We developed two experiments to test the continuity account of actual causation just presented.
+Both experiments asked participants to imagine that they worked in a nuclear control room, and that their job was to monitor networks of particle detectors.
+Participants were told that ``for each activation sequence that you see, your job is to decide what caused the activation of the GCM.''
+
 %The first experiment explores some simple cases directly analogous to the smoke detector example.
 The second experiment explores cases that elaborate that distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
-The cover story asked participants to imagine that they worked in a nuclear control room and that their job was to monitor networks of particle detectors.
+%The cover story asked participants to imagine that they worked in a nuclear control room and that their job was to monitor networks of particle detectors.
 Detectors were shown as white squares (as in Figure~\ref{fig:intro}) that become active and turn black when they absorb a radioactive particle.
@@
 Both experiments included chains in which the GCM received input from one detector, and dual branch networks such as Figure~\ref{fig:intro}a in which the GCM received input from two detectors.
-For chains, the continuity account predicts that participants will tend to choose the root cause of the effect (ie the detector that initiates the activation sequence) rather than the immediate cause (the detector that immediately precedes the GCM).
+For chains, the continuity account predicts that participants will tend to choose the root cause of the effect (i.e.\ the detector that initiates the activation sequence) rather than the immediate cause (the detector that immediately precedes the GCM).
 For dual branch networks, the continuity account predicts that participants will tend to choose the root cause on the branch whose activation is temporally continuous with the activation of the GCM.
@@
 Our presentation of the experiments focuses on two theoretical accounts: the continuity account and the probability raising account.
-For completeness, though, we first discuss the counterfactual and physical process accounts.
+For completeness, though, we first discuss the counterfactual and physical process accounts of causation.
 Both accounts can be formulated in different ways.
-Here we present what we take to be the default predictions of the two accounts, and the General Discussion considers ways in which the accounts can be adjusted to make different predictions.
+Here we present what we take to be the default version of each account, and argue that neither makes clear predictions about our task.
+The General Discussion considers ways in which these default formulations can be adjusted to better account for our data.
 
@@
 
-Because both of these accounts make no clear predictions, it seemed possible that people's inferences about the networks in our experiments would be highly variable and would reveal no clear trends.
+Because both of these general accounts of causation make no clear predictions, it seemed possible that people's inferences about the networks in our experiments would be highly variable and would reveal no clear trends.
 Our first experiment therefore explored simple cases analogous to the smoke detector example in Figure~\ref{fig:intro} with the goal of establishing whether the basic experimental procedure was viable.

### 2021-01-21 docs/resources/CogSci_paper/paper.tex

@@
 Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}.
-Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{}, force dynamics \cite{wolff} and physical simulation~\cite{gerstenberg}.
+Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.
 Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.
@@
 
-Most accounts of actual causation are consistent with one of two broad views of causation: the counterfactual approach and the physical process approach.
+Most accounts of actual causation are consistent with one of two broad views of causation: the counterfactual approach and the physical process approach~\cite{hall04}.
 The counterfactual approach suggests that the tripping of relay Q29BD is a cause of the elevator stopping because if the relay had not tripped then the elevator would not have stopped.
@@
 
-\nocite{halpernp01}
-
 \bibliographystyle{apacite}

### 2021-01-22 docs/resources/CogSci_paper/paper.tex
added to discussion

@@
 %\title{Identifying actual causes by tracking Changes of State in Continuous Time}
-\title{Actual Causes as Changes of State in Continuous Time}
+%\title{Actual Causes as Changes of State in Continuous Time}
+\title{Temporal Continuity and the Judgment of Actual Causation}
 
@@
 
-\textbf{Keywords:} 
+abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract
+abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract
+
+\textbf{Keywords: causal inference; causal selection; causal explanation} 
 \end{abstract}
@@
 
-Outline:
-
-\textbf{I}. Summary and implications of the results
-
-\textbf{II}.
+%Outline:
+%
+%\textbf{I}. Summary and implications of the results
+%
+%\textbf{II}.
 To say why these results cannot really be explained by the current theories of causation and why our theory can better account for them.
 I don't know if I should talk more about CF and PP accounts than I do in introduction, like saying something like that is probably not useful:
-
-[\textit{According to the CF account, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B.
+%
+%[\textit{According to the CF account, the general idea is that an event A is said to be a cause of a distinct event B if the occurrence of A makes a difference in the occurrence of B.
 More precisely A is a cause of B if and only if A and B are true, and A hadn't occurred B wouldn't have occurred.
@@
 According to us the main problem of this interpretation is that its best models explicitely rejects the idea of grounding causation on temporal information.
-According to the PP account, an event A causes B when there is a physical connection between them.
+%According to the PP account, an event A causes B when there is a physical connection between them.
 In one of its latest and probably most convincing formulation, this theory distinguishes two things : causal process and causal interaction (i.e.
@@
 However the theory goes sometimes against some of our best causal intuitions and to makes implicitely use of the CF analysis to explain our judgements: if I hold the head of my ennemy under water and make him die, I'm not genuinely the cause of his death; rather I'm actually preventing the possibility of a genuine causation which is breathing oxygen in order to live.}]
-
-Or if I do have to explicit both theories I should insist on what the predictions of both theories are for our experiments and why they cannot explain people's intuition (the role of time).
-
-\textbf{III}.
+%
+%Or if I do have to explicit both theories I should insist on what the predictions of both theories are for our experiments and why they cannot explain people's intuition (the role of time).
+%
+%\textbf{III}.
 To insist, maybe, on the work of Glymour where our project in part originated: explaining his caveat about treating events as pure propositions.
@@
 How we could use formal tools like Graphical models?
+%
+%\textbf{IV}. To present further experiments that would be interesting to make (with loops, cases of prevention, etc.)
 
-\textbf{IV}. To present further experiments that would be interesting to make (with loops, cases of prevention, etc.)
+We presented an account of actual causation that highlights the role of temporal continuity and described experiments that support two of its predictions.
+First, people tend to identify root causes rather than immediate causes as the main cause of an effect.
+Second, when an effect is produced by the convergence of multiple causal pathways, participants tend to identify the pathway that has temporal continuity with the effect as the primary cause.
+
+Because the continuity account considers changes of state in continuous time, it can exploit a ``no coincidences'' principle to identify a single main cause of an effect.
+This principle applies broadly to causation in physical systems, but is less applicable to social scenarios such as voting scenarios~\cite{livengood13} Exceptions to the ``no coincidences'' principle are possible even for physical systems, and the continuity account can explain the illusory perception of actual causation that occurs in some such cases~\cite{thorstadw16}.
+For example, an 11 year old boy whacked a stick against a telephone pole at the instant at which the Great Northeastern Blackout hit his town, and ran home terrified that he had caused the power outage~\cite{gelbr1965night}.
+
+The continuity account highlights ideas such as the ``no coincidences'' principle that go beyond default formulations of the counterfactual and physical process accounts of causation.
+Key predictions of the account, however, can be reconstructed within both of these general frameworks.
+The counterfactual account can exploit fine-grained temporal information if each event is supplemented with a timestamp.
+If the effect to be explained in Figure~\ref{fig:intro} ``the activation of the GCM at exactly 10 sec after the start of the trial,'' then the counterfactual account suggests that \ev{R}{C} is a cause of the effect but \ev{R}{D} is not, because if \ev{R}{D} had been different (e.g.\ by occurring slightly earlier or later) then the effect would still have occurred at the same instant.
+
+As suggested earlier, standard formulations of the physical process view treat both \ev{R}{C} and \ev{R}{D} as causes of the effect in Figure~\ref{fig:intro}.
+Our data are consistent with the possibility that people view \ev{R}{C} and \ev{R}{D} as equally valid causes, but pick \ev{R}{D} when forced to break the tie for reasons that may be relatively superficial.
+For example, perhaps \ev{R}{D} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch.
+We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a paradigm in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.
+
+% Other future studies
 

### 2021-01-27 docs/resources/CogSci_paper/paper.tex
paper modified
added stats

@@
 
-abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract
-abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract
+Psychological theories of actual causation aim to characterize which of multiple causes of an event is singled out as the primary cause.
+We present one such theory called the continuity account of actual causation.
+The continuity account treats events as changes of state in continuous time and traces a sequence of stage changes backwards through time from an event to its primary cause.
+The account is broadly compatible with the physical process view of causation, and we test it by asking people to identify the primary cause of events occurring in simple physical systems.
+An initial experiment confirms that root causes are more likely to be chosen as primary causes than are immediate causes.
+A second experiment demonstrates that root causes that have temporal continuity with the effect are preferred even when probability raising accounts would predict otherwise.
+The results of both experiments are consistent with the continuity account, and suggest that inferences about changes of state in continuous time may underpin an important class of actual causation judgments.
 
-\textbf{Keywords: causal inference; causal selection; causal explanation} 
+\textbf{Keywords: causal inference; causal selection; singular causation; token causation; causal explanation} 
 \end{abstract}
 
-\section{Introduction}
+%\section{Introduction}
 
 Around 5:27 pm on November 9, 1965, Martin Saltzman found himself trapped in a dark elevator roughly a quarter of a way up the Empire State building~\cite{gelbr1965night}.
-The stopping of Saltzman's elevator was one of thousands of unexpected events that occurred that evening, and virtually all of the people involved must have wondered about the causes of these events.
+The stopping of Saltzman's elevator was one of thousands of unexpected events that occurred that evening, and all of the people involved must have wondered about the causes of these events.
 Subsequent accounts of the Great Northeastern Blackout often trace these events back to the tripping of a relay on line Q29BD in Ontario that triggered a cascade of failures.
@@
 Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}.
-Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.
+Psychologists and philosophers have explored these judgments in detail~and have developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.
 Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.
@@
 The continuity account relies on two core principles.
-First, it assumes that the effect to be explained and its cause are both changes of state.
+First, the effect to be explained and its cause are both changes of state.
 The tripping of a relay qualifies as a candidate cause, but the steady-state setting of a relay does not.
@@
 Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause.
-Some previous work on actual causation highlights the importance of time~\cite{youngs09,stephanmw20e}, but to our knowledge previous work has not explored the implications of temporal continuity in the way we do here.
+Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09}.
+To our knowledge, however,  previous work has not explored the implications of temporal continuity in the way that we do here.
 
@@
 
+%\begin{figure*}
+%\begin{center}
+%\includegraphics[width=7cm]{figures/demo.eps}
+%\includegraphics{figures/introfig.eps}
+%\end{center}
+%\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch).
+All components are initially inactive (ii) Detector \ev{R}{D} activates and activity propagates along the top branch.
+(iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the alarm, which activates only when both  \ev{I}{D} and \ev{I}{C} are active.
+(b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the smoke alarm.}
+%\label{fig:intro}
+%\end{figure*}
+
 \begin{figure*}
 \begin{center}
-%\includegraphics[width=7cm]{figures/demo.eps}
-\includegraphics{figures/introfig.eps}
+\null\hfill
+\subfloat[\label{subfig:1a}]{%
+\includegraphics{figures/fig1a.eps}
+}
+\hfill
+\subfloat[\label{subfig:1b}]{%
+\includegraphics{figures/fig1b.eps}
+}
+\hfill\null
 \end{center}
-\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch).
+\caption{(a)(i) A network of particle detectors.
+All components are initially inactive.
+(ii) Detector \ev{R}{D} activates and activity propagates along the top branch.
+(iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the GCM, which activates only when both  \ev{I}{D} and \ev{I}{C} are active.
+(b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the GCM.}
+%\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch).
 All components are initially inactive (ii) Detector \ev{R}{D} activates and activity propagates along the top branch.
@@
 The particle detection network, however, could also be considered part of a broader causal system that includes both the network and the network's surroundings.
-For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom will probably be identified as the immediate cause of \ev{R}{C} and the main cause of the GCM's activation.
+For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom could be identified as the immediate cause of \ev{R}{C} and the main cause of the GCM's activation.
 
@@
 
-\textbf{Participants}. 30 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment.
+\textbf{Participants}. 30 participants were recruited via Amazon Mechanical Turk and paid \$3 for an 18 minute experiment.
 
@@
 \textbf{Materials}.
-The experiment used a customized interface built using the jsPsych library~\cite{XXX}.
+The experiment used a customized interface built using the jsPsych library~\cite{jspsych}.
 For all networks presented, participants clicked a ``Run'' button to observe an activation sequence.
@@
 The three \emph{event} versions all showed \ev{R}{D} activating 5 seconds into the sequence, and had delays of 2, 4 and 6 seconds between the activation of \ev{I}{D} and \ev{R}{C}.
+The predictions of the continuity account are unaffected by the delay, but we tested different delays just in case this variable affected people's responses.
+The continuity account also makes the same prediction about both state and event sequences, but we anticipated that the state sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}.
 
@@
 First,  participants were more likely to choose root causes than immediate causes.
-Root and immediate causes were identical for the short sequences and therefore cannot bet distinguished, but the results for medium and long sequences reveal a preference for root causes.
-STATISTICAL ANALYSIS.
-
+Root and immediate causes were identical for the short sequences and therefore cannot be distinguished, but the results for medium and long sequences reveal a preference for root causes.
 The second conclusion is that for dual branch sequences, the root cause on the continuous branch (\ev{R}{C}) is preferred to the root cause on the delayed branch (\ev{R}{D}).
@@
 
-Figure \ref{fig:intro} suggests that event and state sequences lead to similar patterns of responses.
-The continuity account makes the same prediction for both kinds of sequences, but we anticipated that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}.
-If this effect exists, our data suggest that it is relatively small.
+To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms}.
+For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded.
+We then ran a logistic regression in which stativity (i.e.\ state vs event) and network length (long vs medium) were included as predictors of a binary dependent variable (root vs immediate).
+We included a random intercept for participant to allow for statistical dependencies between multiple responses from the same participant.
+The posterior mean of the intercept variable was 10.54 and the 95\% credible interval was $[2.30, 26.54]$, which supports the conclusion that root causes were chosen more often than immediate causes.
+The credible intervals for the other coefficients both included 0 (state: -2.75, $[-10.03, 1.14]$ ; long: 1.85, $[-1.78, 7.33]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.
+
+The second analysis was very similar except that the binary dependent variable now indicated whether participants chose a cause on the continuous branch or the delayed branch.
+The posterior mean of the intercept variable was 2.42 (95\% credible interval $[0.33, 4.83]$), indicating that the continuous branch was chosen more often than the delayed branch.
+The credible intervals for the state variable also excluded 0 (1.39, $[0.32, 2.51]$), suggesting that state sequences led to a stronger preference for the continuous branch than did event sequences.
+Credible intervals for both length variables included 0 (medium: 0.17, $[-0.91,1.31]$ ; long: 0.11, $[-1.01,1.21]$, where the reference level is small), suggesting that the preference for the continuous branch was not affected by network length.
 
@@
 
-Thus far we have focused on dual branch sequences, and the chain sequences are of interest primarily because they provide additional evidence of a preference for root causes.  NUMBERS HERE
+%Thus far we have focused on dual branch sequences, and the chain sequences are of interest primarily because they provide additional evidence of a preference for root causes.  NUMBERS HERE
 
@@
 %\includegraphics[width=8cm]{results_E1}
-\includegraphics{figures/expt1.eps}
+%\includegraphics{figures/expt1.eps}
+\includegraphics{figures/expt1test.eps}
 \end{center}
@@
 (b)  Responses for dual branch sequences.
-The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence.}
+The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence.
+Error bars show the standard error of the mean estimated using XXX bootstrap samples.}
 \label{fig:expt1}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+%\includegraphics[width=8cm]{results_E2}
+\includegraphics{figures/expt2.eps}
+%\includegraphics{figures/expt2test.eps}
+%\includegraphics{figures/results_E2.eps}
+\end{center}
+\caption{Experiment 2.
+(a) The probability raising and continuity accounts make different predictions for dual branch sequences in which \ev{R}{C} occurs before \ev{R}{D}.
+(b) The two accounts agree when \ev{R}{C} occurs after \ev{R}{D}.
+(c) Responses for the three possible orders of the key events.} 
+\label{fig:expt2}
 \end{figure}
@@
 \textbf{Participants}.
-100 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment.
+100 participants were recruited via Amazon Mechanical Turk and paid \$1 for a 6 minute experiment.
 Two did not submit a valid completion code, leaving 98 sets of responses for analysis.
@@
 Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b.
-Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C},\ev{I}{C}\}, and a similar (but weaker) effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\} sequences.
-The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D},\ev{I}{C}\} sequences, even though the probability raising account makes the opposite prediction.
-STATISTICS HERE.
+Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}, and a similar effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\} sequences.
+The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\} sequences, even though the probability raising account makes the opposite prediction.
+When referring to the orders we have dropped the fourth event because this event is always \ev{I}{C}.
 
+To further analyze the data we used mixed effects models analogous to the two described for Experiment 1.
+The first analysis used a logistic regression in which order (i.e.\ the temporal order of the key events) was included as a predictor of a binary dependent variable (root or immediate).
+The posterior mean of the intercept variable was 5.75 and the 95\% credible interval $[2.44, 10.52]$, which supports the conclusion that root causes were chosen more often than immediate causes.
+The credible intervals for both order variables included 0 (\{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: 2.11, $[-1.53,1.21]$; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.38 $[-1.78,0.90]$, where the reference level was \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\}), suggesting that the preference for root causes was fairly consistent across the three orders.
+
+The second and more critical analysis used a binary dependent variable that indicated whether participants chose a cause on the continuous branch or the delayed branch.
+The posterior mean of the intercept variable was 2.0 (95\% credible interval $[1.4, 2.65]$), indicating that the continuous branch was chosen more often than the delayed branch.
+The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
 Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2.
 This difference may reflect the increased difficulty of Experiment 2.
-When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy fo choosing the immediate cause that directly precedes the effect.
+When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy of choosing the immediate cause that directly precedes the effect.
 
@@
 As expected the type of network had however no effect on the responses for I1 nor I2 and the percentages of responses for I2 and the root change that initiates the continuous sequence for the Event network (31\% and 34\% for respectively the ``R1 cont'' and ``R2 cont'' networks) were approximately the same.
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=8cm]{results_E2}
-\includegraphics{figures/expt2.eps}
-\end{center}
-\caption{Experiment 2.
-(a) The probability raising and continuity accounts make different predictions for dual branch sequences in which \ev{R}{C} occurs before \ev{R}{D}.
-(b) The two accounts agree when \ev{R}{C} occurs after \ev{R}{D}.
-(c) Responses for the three possible orders of the key events.} 
-\label{fig:expt2}
-\end{figure}
 
@@
 Because the continuity account considers changes of state in continuous time, it can exploit a ``no coincidences'' principle to identify a single main cause of an effect.
-This principle applies broadly to causation in physical systems, but is less applicable to social scenarios such as voting scenarios~\cite{livengood13} Exceptions to the ``no coincidences'' principle are possible even for physical systems, and the continuity account can explain the illusory perception of actual causation that occurs in some such cases~\cite{thorstadw16}.
+This principle applies broadly to causation in physical systems, but is less applicable to social scenarios such as voting scenarios~\cite{livengood13}.
+Exceptions to the ``no coincidences'' principle are possible even for physical systems, and the continuity account can capture the illusory perceptions of actual causation that arise in some such cases~\cite{thorstadw16}.
 For example, an 11 year old boy whacked a stick against a telephone pole at the instant at which the Great Northeastern Blackout hit his town, and ran home terrified that he had caused the power outage~\cite{gelbr1965night}.
@@
 For example, perhaps \ev{R}{D} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch.
-We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a paradigm in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.
+We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a task in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.
 
 % Other future studies
+
+The experimental paradigm we have developed can easily be extended to a more diverse set of causal systems.
+For example, by introducing preventive causal links, we can explore case of double prevention (e.g.\ cases where an active detector that would have prevented the activation of the GCM is itself inactivated by another detector).
+\citeA{glymourstone} point out that current work on actual causation is based largely on discussions of an ``infinitesimal fraction'' of the set of possible cases, but our particle detector task can support a comprehensive survey of people's judgments about actual causation.
 

### 2021-01-27 docs/resources/CogSci_paper/paper.tex
bramley citation

@@
 Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause.
-Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09}.
+Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09,davisbr20}.
 To our knowledge, however,  previous work has not explored the implications of temporal continuity in the way that we do here.
@@
 \textbf{Results}.
+
 Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.

### 2021-01-27 analysis/statistical_analyses.Rmd
stats analyses

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)
library(here)
library(ggthemes)
library(lme4)
library(brms)
```

```{r data, include=TRUE}
dpath_exp1 <- here("data", "responsesOnly_E1_100def.csv")
dpath_exp2 <- here("data", "responsesOnly_E2_100def.csv")

mytheme <-  theme_classic(base_size = 10)  + 
            theme(strip.background = element_blank())

theme_set(mytheme)
```

```{r loadds, include=TRUE}

# Experiment 1 data frame 
d1 <- read_csv(dpath_exp1) #%>%

```

### Plot data

```{r exp1plot, include=TRUE, fig.height=4, fig.width=6}

d1plot <- d1 %>% 
  filter(stimulustype != "chain") %>% 
  group_by(state_or_event, length, response) %>% 
  summarize(count = n()) %>% 
  mutate(response = factor(response, levels = c("RC_IC", "RC", "RD_ID", "RD", "IC", "other")))

plot1 <- d1plot %>% 
  ggplot(aes(x = response, y=count)) +
  geom_bar(stat="identity") +
  facet_grid(state_or_event~length)
plot1

```

### Stats for Expt 1

```{r exp1stats, include=TRUE, fig.height=4, fig.width=6}

sdif_coding <- function(x) {
  x <- as.factor(x)
  contrasts(x) <- MASS::contr.sdif(levels(x))
  x
}

d1_stats1 <- d1 %>% 
  mutate(ri = factor(case_when(response=="RC_IC" ~  NA_character_,
                        response=="RC" ~     "root",
                        response=="RD_ID" ~  NA_character_,
                        response=="RD" ~     "root",
                        response=="IC" ~     "immediate",
                        response=="ID" ~     "immediate",
                        TRUE ~ NA_character_ 
                        )),
         cd = factor(case_when(response=="RC_IC" ~  "continuous",
                        response=="RC" ~     "continuous",
                        response=="RD_ID" ~  "delayed",
                        response=="RD" ~     "delayed",
                        response=="IC" ~     "continuous",
                        response=="ID" ~     "delayed",
                        TRUE ~ NA_character_ 
                        )),
         #lengthc = sdif_coding(factor(length))
         lengthc = contr.sum(factor(length))  # use sum coding
        )

m0 <- glmer(ri ~ state_or_event + length + (1|participantnum), data = d1_stats1, 
            family = binomial(link = "logit"))

nchain <- 1 # for testing

fit1_ri <- brm( ri ~ state_or_event + lengthc + (1|participantnum), family=bernoulli(), data = d1_stats1, chains = nchain)

fit1_cd <- brm( cd ~ state_or_event + lengthc + (1|participantnum), family=bernoulli(), data = d1_stats1, chains = nchain)

```

### Fig 2a

```{r fig2a, include=TRUE, fig.height=4, fig.width=6}

#dsumm <- d %>% 
#  select(character, Oracle, Bronze, Seal, simplified, traditional, `first period`) %>% 
#  pivot_longer(cols=all_of(periods), names_to="period", values_to = "complexity" ) %>% 
#  mutate(period = factor(period, levels=periods, order=TRUE)) %>% 
#  rename(first = `first period`)

fig2a <- dsumm %>% 
  ggplot(aes(x=period, y=complexity, group=character))+
  geom_line(alpha = 0.05) 
fig2a
```

### Fig 2b

```{r fig2b, include=TRUE, fig.height=4, fig.width=6}

dmeans <- dsumm %>%
  group_by(period, first) %>%
  dplyr::summarize( n=sum(!is.na(complexity)), sd = sd(complexity, na.rm=TRUE), se=sd/sqrt(n), complexity=mean(complexity, na.rm=TRUE), .groups="drop")  %>%
  group_by(first) %>%
  mutate(n = mean(n))

fig2b <- dmeans %>%
  #filter(first!="traditional") %>%
  ggplot(aes(x=period, y=complexity, group=first, color=first,size=n))+
  geom_line(alpha = 0.7) +
  geom_point(alpha = 1.0, size=1) +
  geom_errorbar(aes(ymin=complexity-se, ymax=complexity+se), size = 0.5 , width=0.1) +
  scale_color_manual(values = period_col, guide='none') +
  scale_size_continuous(name="count") +
  theme(legend.position = c(0.15, 0.82))

fig2b
ggsave(here("analysis","output", "avcomplexityperfirstperiod.pdf"), plot = fig2b, width = 3.4252,height = 2.8)

```
## Fig 3

```{r fig3, include=TRUE, fig.height=4, fig.width=6}
d_trad_vs_ob <- dsumm %>%
  select(-distinctiveness) %>% 
  pivot_wider(names_from=period, values_from=complexity) %>%
  mutate(complexification= Traditional - Oracle) %>%
  select(character, C1Type, complexification, numcomp)

fig3a <- d_trad_vs_ob %>%
  filter(numcomp <= 5) %>%
  ggplot(aes(x=numcomp, y=complexification, group=numcomp))+
  geom_violin(fill="gray80", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_hline(yintercept=0,linetype="dashed") +
  xlab("Number of components")

fig3a
fig3b <- d_trad_vs_ob %>%
  ggplot(aes(x=C1Type, y=complexification, group=C1Type))+
  geom_violin(fill="gray80", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_hline(yintercept=0,linetype="dashed") +
  xlab("Character type")
fig3b

patchr <- fig3b / fig3a

patchrprint <- patchr + 
  plot_annotation(tag_levels = c('a'), tag_suffix = ')')

ggsave(here("analysis","output", "changebyclasses.pdf"), plot = patchrprint, width = 3.5, height = 2.8)

patchall <- fig2b + 
  patchr + 
  plot_layout(widths=c(1.7,2)) + 
  plot_annotation(tag_levels = c('a'), tag_suffix = ')')

patchall
ggsave(here("analysis","output", "complexitychange.pdf"), plot = patchall, width = 7.08661,height = 2.8)

```

##  Frequency
```{r frequency, include=TRUE}

dsumm_bin <- dsumm %>% 
  filter(first == 'Oracle') %>% 
  mutate(freqbin = -floor(log10(frequency))) %>% 
  group_by(period, freqbin) %>% 
  dplyr::summarize(complexity = mean(complexity)) %>% 
  drop_na()

wireframe(complexity ~ freqbin * period, 
          data = dsumm_bin,
          screen = list(z = 45, x=-60),
          )

freqplot <- dsumm_bin %>% 
  mutate(freqbin = -freqbin) %>% 
  ggplot(aes(x=period, y=complexity, group=freqbin, color=freqbin)) +
  geom_line()
freqplot

```  
##  Distinctiveness

```{r distinct, include=TRUE}
dmeans_distinct <- dsumm %>%
  group_by(period, first) %>%
  dplyr::summarize( n=sum(!is.na(distinctiveness)), sd = sd(distinctiveness, na.rm=TRUE), se=sd/sqrt(n), complexity=mean(distinctiveness, na.rm=TRUE), .groups="drop")  %>%
  group_by(first) %>%
  mutate(n = mean(n))

fig2b_distinct <- fig2b %+% dmeans_distinct +
  ylab('distinctiveness') +
  theme(legend.position = c(0.15, 0.76))

fig2b_distinct
ggsave(here("analysis","output", "avdistinctivenessperfirstperiod.pdf"), plot = fig2b_distinct, width = 3.4252,height = 2.8)

fig_complexvsdistinct <- dsumm %>% 
  ggscatter(x = "complexity", y = "distinctiveness",
    size = 0.5,
    add = "reg.line",  
    add.params = list(color = "blue", fill = "lightgray"), 
    ggtheme = mytheme
  ) +
  stat_cor(
    aes(label = ..r.label..),
    label.x = 3,
    cor.coef.name='r'
  ) +
  coord_fixed(ratio=2)

fig_complexvsdistinct

patchall <- fig_complexvsdistinct +
  fig2b_distinct + 
  #plot_layout(widths=c(1.7,2)) + 
  plot_annotation(tag_levels = c('a'), tag_suffix = ')')

ggsave(here("analysis","output", "distinctivenesschange.pdf"), plot = patchall, width = 7.08661,height = 2.8)

#  ggplot(aes(x=complexity, y=discriminability)) +
#  geom_point() +
#  geom_smooth(method=lm) +
#  stat_cor(method="pearson")

```

##  Mixed effects regression.

For simplicity we treat period as a numeric variable. We run the same approach with complexity and distinctiveness as dependent variables.

```{r frequentist, include=TRUE}

# Model comparison:

# M0: Complexity ~ 1 + (1+Period | Character)
# M1: Complexity ~ Period + (1+Period | Character)

dsumm_numeric <- dsumm %>% 
  mutate(period = as.numeric(period))

dc_numeric <- d_c %>% 
  mutate(period = as.numeric(period))

dvs <- c("complexity")

for (dv in dvs) {
  dforstats <- dsumm_numeric #%>% 
    #filter(!is.na(!!as.symbol(dv)))

  m0 <- lmer(as.formula(paste0(dv, "~ 1 + (1+period|character)")), data = dforstats, REML = FALSE)  
  m1 <- lmer(as.formula(paste0(dv, "~ period + (1+period|character)")), data = dforstats, REML = FALSE)

  beta_period <- round(coef(summary(m1))[2,"Estimate"], 2)
  ci_period <- confint(m1)

  a <- anova(m1, m0)
  # pull out p value and chi-squared statistic
  p_value <-  signif(a$"Pr(>Chisq)"[2] , digits=3)
  #chi_sq <- signif(a$Chisq[2], digits=3)
  chi_sq <- a$Chisq[2]
  df <- a$Df[2]

  summ <- sprintf( "The likelihood ratio test reveals a statistically significant relationship between period and %s ($\\beta$ = %.2f, 95%% CI =  [ %.2f, %.2f], $\\chi^2$(%d) = %.2f, p =  %s). The coefficient for period indicates that moving from one period to the next tends to increase %s by around %.2f units", dv, beta_period, round(ci_period[6,1], 2), round(ci_period[6,2], 2), df, chi_sq, as.character(p_value), dv, beta_period)

  print(summ)

  aprint <- a %>% 
    select(AIC, BIC, Chisq, Df, `Pr(>Chisq)`) %>% 
    rownames_to_column() %>% 
    rename(model=rowname) %>% 
    mutate(model = recode(model, "m0"="1 + (1+period|character)", 
                                 "m1"= "period + (1+period|character)")) %>% 
    mutate( `Pr(>Chisq)` = as.character(signif(`Pr(>Chisq)`, digits=3)))
  mstable <- xtable(aprint, caption = paste0("Comparison between models with and without period as a predictor. The dependent variable is ", dv, "."), label = paste0("tab:regtable_linear_", dv))
  print(mstable, file = here("analysis", "output", paste0("regtable_linear_", dv, ".tex")), include.rownames = FALSE)
}

```

The likelihood ratio test reveals a statistically significant relationship between period and complexity ($\beta$ = `r beta_period`, 95% CI =  [ `r round(ci_period[6,1], 2)`, `r round(ci_period[6,2], 2)`], $\chi^2$(`r df`) = `r chi_sq `, p =  `r p_value `). The coefficient for period indicates that moving from one period to the next tends to increase complexity by around `r beta_period` units.

```{r frequentistfreqdata, include=TRUE}

dsumm_numeric_freq <- dsumm_numeric %>% 
  filter(first=="Oracle") %>%   # only characters in Oracle stream
  filter(period<5) %>%   # ignore simplified b/c singular fit when we include it
  filter(!is.na(logfrequency)) %>% 
  mutate(periodscale = scale(period)[,1]) %>% 
  #mutate(periodscale = period) %>% 
  mutate(rarity= -logfrequencyscale) %>% 
  mutate(rarityraw= -logfrequencyrawscale)

suffix<- "raw"

m0 <- lmer(complexity ~ period + (1+period|character), data = dsumm_numeric_freq, REML = FALSE)  
m1 <- lmer(formula(paste0("complexity ~ rarity", suffix, " + period + (1+period|character)")), data = dsumm_numeric_freq, REML = FALSE)  
m2 <- lmer(formula(paste0("complexity ~ rarity", suffix, " + period + rarity", suffix, "*period + (1+period|character)")), data = dsumm_numeric_freq, REML = FALSE)

beta_freq <- round(coef(summary(m2))[2,"Estimate"], 2)
beta_period <- round(coef(summary(m2))[3,"Estimate"], 2)
beta_product<- round(coef(summary(m2))[4,"Estimate"], 2)

ci_period <- confint(m2)

ci_periodtab <- ci_period %>%
  as_tibble() %>% 
  tail(4) %>% 
  rename(lint = "2.5 %", rint = "97.5 %") %>% 
  mutate(lint = paste0("[", round(lint, digits=1), ","), rint = paste0(round(rint, digits=1), "]")) %>% 
  unite(`95 % CI`, lint, rint, sep=" ")

m2table <- tidy(m2) %>% 
  select(term, estimate) %>% 
  rename(coefficient = "term") %>% 
  head(4) %>% 
  cbind(ci_periodtab)

m2tableout <- xtable(m2table, caption = "Coefficients for a mixed-effects regression model with with formula complexity $\\sim$ rarity + frequency + rarity*frequency + (1 + period$|$ character). Both rarity and frequency were standardized so the coefficients for these variables indicate the increase in complexity expected when the predictor increases by one standard deviation.", label = paste0("tab:coeftable_freq", suffix))
print(m2tableout, file = here("analysis", "output", paste0("coeftable_freq", suffix, ".tex")), include.rownames = FALSE)

a <- anova(m2, m1, m0)

aprint <- a %>% 
  select(AIC, BIC, Chisq, Df, `Pr(>Chisq)`) %>% 
  rownames_to_column() %>% 
  rename(model=rowname) %>% 
  mutate(model = recode(model, "m0"="period + (1+period|character)", 
                               "m1"= "rarity + period + (1+period|character)",
                               "m2"= "rarity + period + rarity * period + (1+period|character)")) %>% 
  mutate( `Pr(>Chisq)` = as.character(signif(`Pr(>Chisq)`, digits=3)))
mstable <- xtable(aprint, caption = "Comparison between models that add rarity and an interaction between rarity and period  to a null model.", label = paste0("tab:regtable_freq", suffix))
print(mstable, file = here("analysis", "output", paste0("regtable_freq", suffix, ".tex")), include.rownames = FALSE)

# pull out p value and chi-squared statistic
p_value <-  signif(a$"Pr(>Chisq)"[2] , digits=3)
chi_sq <- signif(a$Chisq[2], digits=3)
df <- a$Df[2]
```

Here we've scaled both `period` and `logfrequency`, and for ease of interpretation we've inverted `logfrequency` to create a variable called `rarity` (ie bigger values denote rarity). The likelihood ratio tests indicate that the best model includes both a fixed effect of rarity and an interaction between rarity and period. All coefficients are positive, suggesting that complexity increases as period increases, complexity increases as rarity increases, and that the complexity increases faster with period as rarity increases.

## Bayesian analysis with brms

```{r bayesian, include=FALSE}

# Model comparison:

# M0: Complexity ~   1 + (1 + mo(Period) | Character)
# M1: Complexity ~   mo(Period) + (1 + mo(Period) | Character)

# Based in part on
#https://github.com/paul-buerkner/monotonic-effects-paper/blob/master/monotonic_effects.Rmd

sdif_coding <- function(x) {
  x <- as.factor(x)
  contrasts(x) <- MASS::contr.sdif(levels(x))
  x
}

dsumm_brm <- dsumm %>% 
  mutate(periodl = as.numeric(period)) %>%  
  mutate(periodc = sdif_coding(factor(period))) %>% 
  select(character, period, periodl, periodc, complexity, distinctiveness, first, logfrequency, logfrequencyscale, logfrequencyraw, logfrequencyrawscale)

nchain <- 1 # for testing

# Takes a while to run
#fit1 <- brm( complexity ~ mo(period) + (1 + mo(period) | character), data = dsumm_brm, chains = nchain)
#plot(conditional_effects(fit1))

#fit2 <- brm( complexity ~ periodl + (1 + period| character), data = dsumm_brm, chains = nchain)

fit3 <- brm( complexity ~ periodc + (1 + period| character), data = dsumm_brm, chains = nchain)

#pme1 <- plot(conditional_effects(fit1), plot = FALSE, ask = FALSE)
#pme2 <- plot(conditional_effects(fit2), plot = FALSE, ask = FALSE)
pme3 <- plot(conditional_effects(fit3), plot = FALSE, ask = FALSE)

#pmeall <- pme1[[1]] + pme2[[1]] + pme3[[1]]
#loo(fit1, fit2, fit3)
#model_weights(fit1, fit2, fit3, weights = "loo")

coefnames <- fixef(fit3) %>% rownames()
coefsamples <- fixef(fit3, summary=FALSE, pars=coefnames[-1]) %>% 
  as_tibble() %>% 
  rename("Bronze-Oracle"=periodcBronzeMOracle, "Seal-Bronze"=periodcSealMBronze, "Traditional-Seal"=periodcTraditionalMSeal, "Simplified-Traditional"=periodcSimplifiedMTraditional)
postplot <- mcmc_areas(coefsamples)

ggsave(here("analysis","output", "bayesreg_period.pdf"), plot = postplot,  width=4,height = 2.8)

#comp2 <- loo_compare(loo(fit1), loo(fit2), loo(fit3)) %>%
#  as.data.frame() %>%
#  select(elpd_loo, elpd_diff, se_diff, p_loo, looic) %>%
#  rename(
#    "ELPD-Diff" = elpd_diff, "SE-Diff" = se_diff, 
#    "ELPD-LOO" = elpd_loo, "P-LOO" = p_loo, "LOOIC" = looic
#  )
#mw2 <- model_weights(fit1, fit2, fit3, weights = "loo")[rownames(comp2)]
#comp2 %>%
#  cbind("Akaike-Weight" = mw2) %>%
#  apa_table(
#    format = "latex", booktabs = TRUE, digits = 2,
#    caption = "Comparison of models fit4 to fit6 based on approximate leave-one-out cross-validation.",
#    note =  "ELPD-LOO = expected log posterior predictive density (higher is better); ELPD-DIFF = difference in ELPD values compared to the best model. SE-DIFF = standard error of the ELPD difference. P-LOO = effective number of model parameters (lower is better); LOOIC: leave-one-out information criterion (lower is better); Akaike-Weight = Model weight based on the LOOIC values (higher is better).",
#    align = c("l", rep("r", 6))
#  )

```

```{r bayesianfreqdata, include=FALSE}
dsumm_brm_freq <- dsumm_brm %>% 
  filter(first=="Oracle") %>%   # only characters in Oracle stream
  filter(period<"Simplified") %>%   # ignore simplified b/c singular fit when we include it
  filter(!is.na(logfrequency)) %>% 
  mutate(rarity= -logfrequencyscale) %>% 
  mutate(rarityraw= -logfrequencyrawscale) %>% 
  mutate(periodc = sdif_coding(factor(period)))

suffix <- "raw"

dsumm_forstats <- dsumm_brm_freq %>% 
  mutate(rarity = !!as.symbol(paste0("rarity", suffix)))

fit3_freq <- brm( complexity ~ periodc + rarity + periodc*rarity + (1 + period| character), data = dsumm_forstats, chains = nchain)

coefnames <- fixef(fit3_freq) %>% rownames()
coefsamples <- fixef(fit3_freq, summary=FALSE, pars=coefnames[-1]) %>% 
  as_tibble() %>% 
  rename("Bronze-Oracle"=periodcBronzeMOracle, 
         "Seal-Bronze"=periodcSealMBronze, 
         "Traditional-Seal"=periodcTraditionalMSeal, 
         "(Bronze-Oracle):rarity"="periodcBronzeMOracle:rarity", 
         "(Seal-Bronze):rarity"="periodcSealMBronze:rarity", 
         "(Traditional-Seal):rarity"="periodcTraditionalMSeal:rarity")

postplot <- mcmc_areas(coefsamples)

ggsave(here("analysis","output", paste0("bayesreg_freq", suffix, ".pdf")), plot = postplot,  width=5,height = 5)

```

```{r bayesianpictographic, include=FALSE}

fit3_ctype <- brm( complexification ~ 0+C1Type, data = d_trad_vs_ob, chains = nchain)

d_trad_vs_ob_smallcomp <- d_trad_vs_ob %>%
  filter(numcomp <= 5) %>% 
  mutate(numcomp = factor(numcomp))

fit3_numcomp <- brm( complexification ~ 0+numcomp, data = d_trad_vs_ob_smallcomp, chains = nchain)

coefnames <- fixef(fit3_ctype) %>% rownames()
coefsamples <- fixef(fit3_ctype, summary=FALSE, pars=coefnames) %>% 
  as_tibble() %>% 
  rename("pictographic"="C1Typepictographic",
         "pictologic"="C1Typepictologic", 
         "pictosyn"="C1Typepictosyn.", 
         "pictophon"="C1Typepictophon.", 
         "other"="C1Typeother")

postplot_ctype <- mcmc_areas(coefsamples) +
                  xlab('complexification')

coefnames <- fixef(fit3_numcomp) %>% rownames()
coefsamples <- fixef(fit3_numcomp, summary=FALSE, pars=coefnames) %>% 
  as_tibble() %>% 
  rename("1 comp"="numcomp1",
         "2 comp"="numcomp2", 
         "3 comp"="numcomp3", 
         "4 comp"="numcomp4", 
         "5 comp"="numcomp5")

postplot_numcomp <- mcmc_areas(coefsamples) +
                  xlab('complexification')

posteriors <- postplot_ctype + postplot_numcomp + 
    plot_annotation(tag_levels = c('a'), tag_suffix = ')')

ggsave(here("analysis","output", "bayesreg_ctype.pdf"), plot = posteriors, width = 7.08661,height = 3.5)

```

# Supporting information

## Alternative complexity measures

```{r alternativecomplexity, include=TRUE}
dpath_pixcomp <- here("data", "hanziyuan_pix_complexity.csv")
dpath_cd<- here("data", "character_dataset.csv")
dpath_gc<- here("data", "gif_complexities.csv")

d_pixcomp <- read_csv(dpath_pixcomp) %>%
    rename(Traditional=traditional,  Simplified=simplified) %>% 
    pivot_longer(cols=periods, names_to = "period", values_to = "pixcomplexity") %>% 
  select(character, period, pixcomplexity) %>% 
  group_by(character, period) %>%
  dplyr::summarize(pixcomplexity=mean(pixcomplexity), .groups="drop")  %>% 
  right_join(dsumm, by = c("character", "period"))

d_cd <- read_csv(dpath_cd) %>%
  select(character, "stroke count") %>% 
  mutate(period="Simplified") %>% 
  right_join(dsumm, by=c("character","period")) %>% 
  mutate(pixcomplexity = `stroke count`)

d_gc <- read_csv(dpath_gc) %>%
  mutate(period="Simplified") %>% 
  right_join(dsumm, by=c("character","period")) %>% 
  mutate(pixcomplexity = `gif complexity`)

myscatter <- function(d) {
  pixcomplexity <- d %>% 
    ggscatter(x = "complexity", y = "pixcomplexity",
    size = 0.5,
    add = "reg.line",  
    add.params = list(color = "blue", fill = "lightgray"), 
    ggtheme = mytheme
  ) +
  stat_cor(
    aes(label = ..r.label..),
    label.x = 3,
    cor.coef.name='r'
  )  + #+ coord_fixed(ratio=2)
  xlab('perimetric complexity')

  return(pixcomplexity)
}

pixcomplexity  <- myscatter(d_pixcomp) +
  ylab('pixel-based complexity')

pixcomplexity 
strokecount<- myscatter(d_cd) +
  ylab('stroke count') 
strokecount

gifcomplexity<- myscatter(d_gc) +
  ylab('writing time')

altcomplex <- pixcomplexity + strokecount + gifcomplexity +
  plot_annotation(tag_levels = c('a'), tag_suffix = ')')

ggsave(here("analysis","output", "altcomplexitymeasures.pdf"), plot = altcomplex,  width=7.08661,height = 2.8)

```

### 2021-01-28 docs/resources/CogSci_paper/paper.tex
minor
added brms prior

@@
 
-To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms}.
+To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms} with default priors\footnote{By default brms uses improper flat priors on the coefficients of fixed effects, and a half student-t prior with 3 degrees of freedom, a location of 0 and a scale of 2.5 on the intercept and on the standard deviations of the random effects.}.
 For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded.
@@
 The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence.
-Error bars show the standard error of the mean estimated using XXX bootstrap samples.}
+Error bars show the standard error of the mean estimated using 100 bootstrap samples.}
 \label{fig:expt1}
@@
 
-\textbf{Participants}.
-100 participants were recruited via Amazon Mechanical Turk and paid \$1 for a 6 minute experiment.
-Two did not submit a valid completion code, leaving 98 sets of responses for analysis.
+\textbf{Participants}. 100 participants were recruited via Amazon Mechanical Turk and paid \$1 for a 6 minute experiment. 
+%Two did not submit a valid completion code, leaving 98 sets of responses for analysis. NO LONGER TRUE
 
@@
 The posterior mean of the intercept variable was 2.0 (95\% credible interval $[1.4, 2.65]$), indicating that the continuous branch was chosen more often than the delayed branch.
-The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
+The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
 Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2.
@@
 For example, by introducing preventive causal links, we can explore case of double prevention (e.g.\ cases where an active detector that would have prevented the activation of the GCM is itself inactivated by another detector).
-\citeA{glymourstone} point out that current work on actual causation is based largely on discussions of an ``infinitesimal fraction'' of the set of possible cases, but our particle detector task can support a comprehensive survey of people's judgments about actual causation.
+\citeA{glymourstone} point out that current work on actual causation is based largely on discussions of an ``infinitesimal fraction'' of the set of possible cases, but our paradigm can potentially be used to carry out a comprehensive survey of judgments about actual causation.
 

### 2021-01-31 analysis/statistical_analyses.Rmd
aurelien edits

@@
 d1 <- read_csv(dpath_exp1) #%>%
+d2 <- read_csv(dpath_exp2) #%>%
 
@@
 
-```{r exp1stats, include=TRUE, fig.height=4, fig.width=6}
-
-sdif_coding <- function(x) {
-  x <- as.factor(x)
-  contrasts(x) <- MASS::contr.sdif(levels(x))
-  x
-}
+```{r exp1stats, eval=FALSE, fig.height=4, fig.width=6}
 
@@
                         TRUE ~ NA_character_ 
-                        )),
+                        ), levels = c("immediate", "root")),
          cd = factor(case_when(response=="RC_IC" ~  "continuous",
@@
                         TRUE ~ NA_character_ 
-                        )),
-         #lengthc = sdif_coding(factor(length))
-         lengthc = contr.sum(factor(length))  # use sum coding
+                        ), levels = c("delayed", "continuous")),
+         length = factor(length, levels = c("short", "medium", "long")),
+         state_or_event= factor(state_or_event, levels = c("event", "state")),
         )
+
+d1_stats1ri <- d1_stats1 %>% 
+  drop_na(ri)
+
+d1_stats1cd <- d1_stats1 %>% 
+  drop_na(cd)
+
+# use sum coding 
+#options(contrasts=c("contr.sum", "contr.poly"))
+options(contrasts=c("contr.treatment", "contr.poly"))
 
@@
 nchain <- 1 # for testing
+niter <- 500 # for testing
 
-fit1_ri <- brm( ri ~ state_or_event + lengthc + (1|participantnum), family=bernoulli(), data = d1_stats1, chains = nchain)
+nchain <- 4 # for real run
+niter <- 2000 # for real run
 
-fit1_cd <- brm( cd ~ state_or_event + lengthc + (1|participantnum), family=bernoulli(), data = d1_stats1, chains = nchain)
+fit1_ri <- brm( ri ~ state_or_event + length + (1|participantnum), family=bernoulli(), data = d1_stats1ri, chains = nchain, iter = niter )
+
+fit1_cd <- brm( cd ~ state_or_event + length + (1|participantnum), family=bernoulli(), data = d1_stats1cd,  chains = nchain, iter = niter )
+
+cd_tab <- fixef(fit1_cd) %>% as_tibble(rownames="NA")
+
+mcdtableout<- xtable(cd_tab, caption = "Coefficients for cd regression, Expt 1" )
+print(mcdtableout, file = here("analysis", "output", paste0("coeftable_cd_exp1.tex")))
+
+ri_tab <- fixef(fit1_ri) %>% as_tibble(rownames="NA")
+
+mritableout<- xtable(ri_tab, caption = "Coefficients for ri regression, Expt 1" )
+print(mritableout, file = here("analysis", "output", paste0("coeftable_ri_exp1.tex")), include.rownames = FALSE)
+
+```
+
+### Fig 2
+```{r exp2plot, include=TRUE, fig.height=4, fig.width=6}
+
+d2plot <- d2 %>% 
+  filter(stimulustype != "chain") %>% 
+  group_by(order, response) %>% 
+  summarize(count = n()) %>% 
+  mutate(response = factor(response, levels = c("RC", "RD", "IC", "ID", "other")))
+
+plot2 <- d2plot %>% 
+  ggplot(aes(x = response, y=count)) +
+  geom_bar(stat="identity") +
+  facet_wrap(~order)
+plot2
+
+```
+
+### Stats for Experiment 2
+
+```{r exp2stats, eval=TRUE, fig.height=4, fig.width=6}
+
+d2_stats1 <- d2 %>% 
+  mutate(ri = factor(case_when(response=="RC" ~     "root",
+                               response=="RD" ~     "root",
+                               response=="IC" ~     "immediate",
+                               response=="ID" ~     "immediate",
+                               TRUE ~ NA_character_ 
+                               ), levels = c("immediate", "root")),
+         cd = factor(case_when(response=="RC" ~     "continuous",
+                               response=="RD" ~     "delayed",
+                               response=="IC" ~     "continuous",
+                               response=="ID" ~     "delayed",
+                               TRUE ~ NA_character_ 
+                               ), levels = c("delayed", "continuous")),
+         order = factor(order, levels = c("RC_RD_ID", "RD_RC_ID", "RD_ID_RC"))
+        )
+
+d2_stats1ri <- d2_stats1 %>% 
+  drop_na(ri)
+
+d2_stats1cd <- d2_stats1 %>% 
+  drop_na(cd)
+
+# use sum coding 
+#options(contrasts=c("contr.sum", "contr.poly"))
+options(contrasts=c("contr.treatment", "contr.poly"))
+
+m0 <- glmer(ri ~ order + (1|participantnum), data = d2_stats1, 
+            family = binomial(link = "logit"))
+
+nchain <- 1 # for testing
+niter <- 500 # for testing
+
+nchain <- 4 # for real run
+niter <- 2000 # for real run
+
+fit2_ri <- brm( ri ~ order + (1|participantnum), family=bernoulli(), data = d2_stats1ri, chains = nchain, iter = niter )
+
+fit2_cd <- brm( cd ~ order + (1|participantnum), family=bernoulli(), data = d2_stats1cd,  chains = nchain, iter = niter )
+
+cd_tab <- fixef(fit2_cd) %>% as_tibble(rownames="NA")
+
+mcdtableout<- xtable(cd_tab, caption = "Coefficients for cd regression, Expt 2" )
+print(mcdtableout, file = here("analysis", "output", paste0("coeftable_cd_exp2.tex")))
+
+ri_tab <- fixef(fit2_ri) %>% as_tibble(rownames="NA")
+
+mritableout<- xtable(ri_tab, caption = "Coefficients for ri regression, Expt 2" )
+print(mritableout, file = here("analysis", "output", paste0("coeftable_ri_exp2.tex")), include.rownames = FALSE)
 

### 2021-01-31 docs/resources/CogSci_paper/paper.tex
figures modified
aurelien edits

@@
 
-Psychological theories of actual causation aim to characterize which of multiple causes of an event is singled out as the primary cause.
-We present one such theory called the continuity account of actual causation.
+Psychological accounts of actual causation aim to characterize which of multiple causes of an event is singled out as the primary cause.
+We present one such account called the continuity account.
 The continuity account treats events as changes of state in continuous time and traces a sequence of stage changes backwards through time from an event to its primary cause.
@@
 The following sections introduce the continuity account of actual causation and present two experiments designed to test core predictions of the account.
-Among previous theories of actual causation the most natural comparison is Spellman's probability raising account, which proposes that the actual cause of an effect is the cause that increased its probability to the greatest extent~\cite{spellman97}.
+Among previous theories of actual causation a natural comparison is Spellman's probability raising account, which proposes that the actual cause of an effect is the cause that increased its probability to the greatest extent~\cite{spellman97}.
 Our second experiment directly compares the continuity account with the probability raising account and we find that the continuity account provides the better account of our data.
@@
 \begin{figure*}
+\captionsetup[subfigure]{labelformat=empty}
 \begin{center}
 \null\hfill
-\subfloat[\label{subfig:1a}]{%
-\includegraphics{figures/fig1a.eps}
-}
+\subfloat[]{\includegraphics{figures/fig1a.eps}\label{subfig:1a}}
 \hfill
-\subfloat[\label{subfig:1b}]{%
-\includegraphics{figures/fig1b.eps}
-}
+\subfloat[]{\includegraphics{figures/fig1b.eps}\label{subfig:1b}}
 \hfill\null
@@
 Because both of these general accounts of causation make no clear predictions, it seemed possible that people's inferences about the networks in our experiments would be highly variable and would reveal no clear trends.
-Our first experiment therefore explored simple cases analogous to the smoke detector example in Figure~\ref{fig:intro} with the goal of establishing whether the basic experimental procedure was viable.
+Our first experiment therefore explored simple cases analogous to the example in Figure~\ref{fig:intro}a with the goal of establishing whether the basic experimental procedure was viable.
 The second experiment focused on more elaborate dual-branch cases that aimed to distinguish between the continuity account and Spellman's probability raising account~\cite{spellman97}.
@@
 Excluding the GCM, each chain and each branch of each dual network had 7 detectors.
+Within each dual branch sequence, the root causes on the two branches (\ev{R}{D} and \ev{R}{C}) were equidistant from the GCM, but these distances varied across sequences.
 We refer to the sequences as \emph{short}, \emph{medium} or \emph{long} based on the distance between the root causes and the GCM.
-Excluding the GCM, the short, medium and long chain sequences showed 1, 3 and 6 active detectors respectively at the end of the sequence.
-Within each dual branch sequence, the root causes on the two branches (\ev{R}{D} and \ev{R}{C}) were equidistant from the GCM, but these distances varied across sequences.
-The sequence in Figure~\ref{fig:intro}a is a medium dual branch sequence (3 active detectors in each branch excluding the GCM), and short and long versions had 1 and 6 active detectors respectively per branch.
+Excluding the GCM, the short, medium and long sequences showed 1, 3 and 6 active detectors respectively per branch at the end of the sequence.
+The sequence in Figure~\ref{fig:intro}a is a medium dual branch sequence (3 active detectors in each branch excluding the GCM).
 
@@
 %\includegraphics[width=8cm]{results_E2}
-\includegraphics{figures/expt2.eps}
-%\includegraphics{figures/expt2test.eps}
+%\includegraphics{figures/expt2.eps}
+\includegraphics{figures/expt2test.eps}
 %\includegraphics{figures/results_E2.eps}
@@
 The counterfactual account can exploit fine-grained temporal information if each event is supplemented with a timestamp.
-If the effect to be explained in Figure~\ref{fig:intro} ``the activation of the GCM at exactly 10 sec after the start of the trial,'' then the counterfactual account suggests that \ev{R}{C} is a cause of the effect but \ev{R}{D} is not, because if \ev{R}{D} had been different (e.g.\ by occurring slightly earlier or later) then the effect would still have occurred at the same instant.
+If the effect to be explained in Figure~\ref{fig:intro} is ``the activation of the GCM at exactly 10 sec after the start of the trial,'' then the counterfactual account suggests that \ev{R}{C} is a cause of the effect but \ev{R}{D} is not, because if \ev{R}{D} had been different (e.g.\ by occurring slightly earlier or later) then the effect would still have occurred at the same instant.
 
 As suggested earlier, standard formulations of the physical process view treat both \ev{R}{C} and \ev{R}{D} as causes of the effect in Figure~\ref{fig:intro}.
-Our data are consistent with the possibility that people view \ev{R}{C} and \ev{R}{D} as equally valid causes, but pick \ev{R}{D} when forced to break the tie for reasons that may be relatively superficial.
-For example, perhaps \ev{R}{D} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch.
+Our data are consistent with the possibility that people view \ev{R}{C} and \ev{R}{D} as equally valid causes, but pick \ev{R}{C} when forced to break the tie for reasons that may be relatively superficial.
+For example, perhaps \ev{R}{C} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch.
 We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a task in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.

### 2021-02-01 analysis/statistical_analyses.Rmd
fit to 6 pages

@@
 library(brms)
+library(xtable)
 ```
@@
 
-```{r exp1stats, eval=FALSE, fig.height=4, fig.width=6}
+```{r exp1stats, eval=TRUE, fig.height=4, fig.width=6}
 
@@
                         response=="ID" ~     "delayed",
+                        response=="otherC" ~     "continuous",
+                        response=="otherD" ~     "delayed",
                         TRUE ~ NA_character_ 
@@
 
-nchain <- 4 # for real run
-niter <- 2000 # for real run
+nchain <- 8 # for real run
+niter <- 4000 # for real run
 
@@
                                response=="ID" ~     "delayed",
+                               response=="otherC" ~     "continuous",
+                               response=="otherD" ~     "delayed",
                                TRUE ~ NA_character_ 
@@
 ```
-
-### Fig 2a
-
-```{r fig2a, include=TRUE, fig.height=4, fig.width=6}
-
-#dsumm <- d %>% 
-#  select(character, Oracle, Bronze, Seal, simplified, traditional, `first period`) %>% 
-#  pivot_longer(cols=all_of(periods), names_to="period", values_to = "complexity" ) %>% 
-#  mutate(period = factor(period, levels=periods, order=TRUE)) %>% 
-#  rename(first = `first period`)
-
-fig2a <- dsumm %>% 
-  ggplot(aes(x=period, y=complexity, group=character))+
-  geom_line(alpha = 0.05) 
-fig2a
-```
-
-### Fig 2b
-
-```{r fig2b, include=TRUE, fig.height=4, fig.width=6}
-
-dmeans <- dsumm %>%
-  group_by(period, first) %>%
-  dplyr::summarize( n=sum(!is.na(complexity)), sd = sd(complexity, na.rm=TRUE), se=sd/sqrt(n), complexity=mean(complexity, na.rm=TRUE), .groups="drop")  %>%
-  group_by(first) %>%
-  mutate(n = mean(n))
-
-fig2b <- dmeans %>%
-  #filter(first!="traditional") %>%
-  ggplot(aes(x=period, y=complexity, group=first, color=first,size=n))+
-  geom_line(alpha = 0.7) +
-  geom_point(alpha = 1.0, size=1) +
-  geom_errorbar(aes(ymin=complexity-se, ymax=complexity+se), size = 0.5 , width=0.1) +
-  scale_color_manual(values = period_col, guide='none') +
-  scale_size_continuous(name="count") +
-  theme(legend.position = c(0.15, 0.82))
-
-fig2b
-ggsave(here("analysis","output", "avcomplexityperfirstperiod.pdf"), plot = fig2b, width = 3.4252,height = 2.8)
-
-```
-## Fig 3
-
-```{r fig3, include=TRUE, fig.height=4, fig.width=6}
-d_trad_vs_ob <- dsumm %>%
-  select(-distinctiveness) %>% 
-  pivot_wider(names_from=period, values_from=complexity) %>%
-  mutate(complexification= Traditional - Oracle) %>%
-  select(character, C1Type, complexification, numcomp)
-
-fig3a <- d_trad_vs_ob %>%
-  filter(numcomp <= 5) %>%
-  ggplot(aes(x=numcomp, y=complexification, group=numcomp))+
-  geom_violin(fill="gray80", draw_quantiles = c(0.25, 0.5, 0.75)) +
-  geom_hline(yintercept=0,linetype="dashed") +
-  xlab("Number of components")
-
-fig3a
-fig3b <- d_trad_vs_ob %>%
-  ggplot(aes(x=C1Type, y=complexification, group=C1Type))+
-  geom_violin(fill="gray80", draw_quantiles = c(0.25, 0.5, 0.75)) +
-  geom_hline(yintercept=0,linetype="dashed") +
-  xlab("Character type")
-fig3b
-
-patchr <- fig3b / fig3a
-
-patchrprint <- patchr + 
-  plot_annotation(tag_levels = c('a'), tag_suffix = ')')
-
-ggsave(here("analysis","output", "changebyclasses.pdf"), plot = patchrprint, width = 3.5, height = 2.8)
-
-patchall <- fig2b + 
-  patchr + 
-  plot_layout(widths=c(1.7,2)) + 
-  plot_annotation(tag_levels = c('a'), tag_suffix = ')')
-
-patchall
-ggsave(here("analysis","output", "complexitychange.pdf"), plot = patchall, width = 7.08661,height = 2.8)
-
-```
-
-##  Frequency
-```{r frequency, include=TRUE}
-
-dsumm_bin <- dsumm %>% 
-  filter(first == 'Oracle') %>% 
-  mutate(freqbin = -floor(log10(frequency))) %>% 
-  group_by(period, freqbin) %>% 
-  dplyr::summarize(complexity = mean(complexity)) %>% 
-  drop_na()
-
-wireframe(complexity ~ freqbin * period, 
-          data = dsumm_bin,
-          screen = list(z = 45, x=-60),
-          )
-
-freqplot <- dsumm_bin %>% 
-  mutate(freqbin = -freqbin) %>% 
-  ggplot(aes(x=period, y=complexity, group=freqbin, color=freqbin)) +
-  geom_line()
-freqplot
-
-```  
-##  Distinctiveness
-
-```{r distinct, include=TRUE}
-dmeans_distinct <- dsumm %>%
-  group_by(period, first) %>%
-  dplyr::summarize( n=sum(!is.na(distinctiveness)), sd = sd(distinctiveness, na.rm=TRUE), se=sd/sqrt(n), complexity=mean(distinctiveness, na.rm=TRUE), .groups="drop")  %>%
-  group_by(first) %>%
-  mutate(n = mean(n))
-
-fig2b_distinct <- fig2b %+% dmeans_distinct +
-  ylab('distinctiveness') +
-  theme(legend.position = c(0.15, 0.76))
-
-fig2b_distinct
-ggsave(here("analysis","output", "avdistinctivenessperfirstperiod.pdf"), plot = fig2b_distinct, width = 3.4252,height = 2.8)
-
-fig_complexvsdistinct <- dsumm %>% 
-  ggscatter(x = "complexity", y = "distinctiveness",
-    size = 0.5,
-    add = "reg.line",  
-    add.params = list(color = "blue", fill = "lightgray"), 
-    ggtheme = mytheme
-  ) +
-  stat_cor(
-    aes(label = ..r.label..),
-    label.x = 3,
-    cor.coef.name='r'
-  ) +
-  coord_fixed(ratio=2)
-
-fig_complexvsdistinct
-
-patchall <- fig_complexvsdistinct +
-  fig2b_distinct + 
-  #plot_layout(widths=c(1.7,2)) + 
-  plot_annotation(tag_levels = c('a'), tag_suffix = ')')
-
-ggsave(here("analysis","output", "distinctivenesschange.pdf"), plot = patchall, width = 7.08661,height = 2.8)
-
-#  ggplot(aes(x=complexity, y=discriminability)) +
-#  geom_point() +
-#  geom_smooth(method=lm) +
-#  stat_cor(method="pearson")
-
-```
-
-##  Mixed effects regression.
-
-For simplicity we treat period as a numeric variable. We run the same approach with complexity and distinctiveness as dependent variables.
-
-```{r frequentist, include=TRUE}
-
-# Model comparison:
-
-# M0: Complexity ~ 1 + (1+Period | Character)
-# M1: Complexity ~ Period + (1+Period | Character)
-
-dsumm_numeric <- dsumm %>% 
-  mutate(period = as.numeric(period))
-
-dc_numeric <- d_c %>% 
-  mutate(period = as.numeric(period))
-
-dvs <- c("complexity")
-
-for (dv in dvs) {
-  dforstats <- dsumm_numeric #%>% 
-    #filter(!is.na(!!as.symbol(dv)))
-
-  m0 <- lmer(as.formula(paste0(dv, "~ 1 + (1+period|character)")), data = dforstats, REML = FALSE)  
-  m1 <- lmer(as.formula(paste0(dv, "~ period + (1+period|character)")), data = dforstats, REML = FALSE)
-
-  beta_period <- round(coef(summary(m1))[2,"Estimate"], 2)
-  ci_period <- confint(m1)
-
-  a <- anova(m1, m0)
-  # pull out p value and chi-squared statistic
-  p_value <-  signif(a$"Pr(>Chisq)"[2] , digits=3)
-  #chi_sq <- signif(a$Chisq[2], digits=3)
-  chi_sq <- a$Chisq[2]
-  df <- a$Df[2]
-
-  summ <- sprintf( "The likelihood ratio test reveals a statistically significant relationship between period and %s ($\\beta$ = %.2f, 95%% CI =  [ %.2f, %.2f], $\\chi^2$(%d) = %.2f, p =  %s).
-The coefficient for period indicates that moving from one period to the next tends to increase %s by around %.2f units", dv, beta_period, round(ci_period[6,1], 2), round(ci_period[6,2], 2), df, chi_sq, as.character(p_value), dv, beta_period)
-
-  print(summ)
-
-  aprint <- a %>% 
-    select(AIC, BIC, Chisq, Df, `Pr(>Chisq)`) %>% 
-    rownames_to_column() %>% 
-    rename(model=rowname) %>% 
-    mutate(model = recode(model, "m0"="1 + (1+period|character)", 
-                                 "m1"= "period + (1+period|character)")) %>% 
-    mutate( `Pr(>Chisq)` = as.character(signif(`Pr(>Chisq)`, digits=3)))
-  mstable <- xtable(aprint, caption = paste0("Comparison between models with and without period as a predictor. The dependent variable is ", dv, "."), label = paste0("tab:regtable_linear_", dv))
-  print(mstable, file = here("analysis", "output", paste0("regtable_linear_", dv, ".tex")), include.rownames = FALSE)
-}
-
-```
-
-The likelihood ratio test reveals a statistically significant relationship between period and complexity ($\beta$ = `r beta_period`, 95% CI =  [ `r round(ci_period[6,1], 2)`, `r round(ci_period[6,2], 2)`], $\chi^2$(`r df`) = `r chi_sq `, p =  `r p_value `).
-The coefficient for period indicates that moving from one period to the next tends to increase complexity by around `r beta_period` units.
-
-```{r frequentistfreqdata, include=TRUE}
-
-dsumm_numeric_freq <- dsumm_numeric %>% 
-  filter(first=="Oracle") %>%   # only characters in Oracle stream
-  filter(period<5) %>%   # ignore simplified b/c singular fit when we include it
-  filter(!is.na(logfrequency)) %>% 
-  mutate(periodscale = scale(period)[,1]) %>% 
-  #mutate(periodscale = period) %>% 
-  mutate(rarity= -logfrequencyscale) %>% 
-  mutate(rarityraw= -logfrequencyrawscale)
-
-suffix<- "raw"
-
-m0 <- lmer(complexity ~ period + (1+period|character), data = dsumm_numeric_freq, REML = FALSE)  
-m1 <- lmer(formula(paste0("complexity ~ rarity", suffix, " + period + (1+period|character)")), data = dsumm_numeric_freq, REML = FALSE)  
-m2 <- lmer(formula(paste0("complexity ~ rarity", suffix, " + period + rarity", suffix, "*period + (1+period|character)")), data = dsumm_numeric_freq, REML = FALSE)
-
-beta_freq <- round(coef(summary(m2))[2,"Estimate"], 2)
-beta_period <- round(coef(summary(m2))[3,"Estimate"], 2)
-beta_product<- round(coef(summary(m2))[4,"Estimate"], 2)
-
-ci_period <- confint(m2)
-
-ci_periodtab <- ci_period %>%
-  as_tibble() %>% 
-  tail(4) %>% 
-  rename(lint = "2.5 %", rint = "97.5 %") %>% 
-  mutate(lint = paste0("[", round(lint, digits=1), ","), rint = paste0(round(rint, digits=1), "]")) %>% 
-  unite(`95 % CI`, lint, rint, sep=" ")
-
-m2table <- tidy(m2) %>% 
-  select(term, estimate) %>% 
-  rename(coefficient = "term") %>% 
-  head(4) %>% 
-  cbind(ci_periodtab)
-
-m2tableout <- xtable(m2table, caption = "Coefficients for a mixed-effects regression model with with formula complexity $\\sim$ rarity + frequency + rarity*frequency + (1 + period$|$ character).
-Both rarity and frequency were standardized so the coefficients for these variables indicate the increase in complexity expected when the predictor increases by one standard deviation.", label = paste0("tab:coeftable_freq", suffix))
-print(m2tableout, file = here("analysis", "output", paste0("coeftable_freq", suffix, ".tex")), include.rownames = FALSE)
-
-a <- anova(m2, m1, m0)
-
-aprint <- a %>% 
-  select(AIC, BIC, Chisq, Df, `Pr(>Chisq)`) %>% 
-  rownames_to_column() %>% 
-  rename(model=rowname) %>% 
-  mutate(model = recode(model, "m0"="period + (1+period|character)", 
-                               "m1"= "rarity + period + (1+period|character)",
-                               "m2"= "rarity + period + rarity * period + (1+period|character)")) %>% 
-  mutate( `Pr(>Chisq)` = as.character(signif(`Pr(>Chisq)`, digits=3)))
-mstable <- xtable(aprint, caption = "Comparison between models that add rarity and an interaction between rarity and period  to a null model.", label = paste0("tab:regtable_freq", suffix))
-print(mstable, file = here("analysis", "output", paste0("regtable_freq", suffix, ".tex")), include.rownames = FALSE)
-
-# pull out p value and chi-squared statistic
-p_value <-  signif(a$"Pr(>Chisq)"[2] , digits=3)
-chi_sq <- signif(a$Chisq[2], digits=3)
-df <- a$Df[2]
-```
-
-Here we've scaled both `period` and `logfrequency`, and for ease of interpretation we've inverted `logfrequency` to create a variable called `rarity` (ie bigger values denote rarity).
-The likelihood ratio tests indicate that the best model includes both a fixed effect of rarity and an interaction between rarity and period.
-All coefficients are positive, suggesting that complexity increases as period increases, complexity increases as rarity increases, and that the complexity increases faster with period as rarity increases.
-
-## Bayesian analysis with brms
-
-```{r bayesian, include=FALSE}
-
-# Model comparison:
-
-# M0: Complexity ~   1 + (1 + mo(Period) | Character)
-# M1: Complexity ~   mo(Period) + (1 + mo(Period) | Character)
-
-# Based in part on
-#https://github.com/paul-buerkner/monotonic-effects-paper/blob/master/monotonic_effects.Rmd
-
-sdif_coding <- function(x) {
-  x <- as.factor(x)
-  contrasts(x) <- MASS::contr.sdif(levels(x))
-  x
-}
-
-dsumm_brm <- dsumm %>% 
-  mutate(periodl = as.numeric(period)) %>%  
-  mutate(periodc = sdif_coding(factor(period))) %>% 
-  select(character, period, periodl, periodc, complexity, distinctiveness, first, logfrequency, logfrequencyscale, logfrequencyraw, logfrequencyrawscale)
-
-nchain <- 1 # for testing
-
-# Takes a while to run
-#fit1 <- brm( complexity ~ mo(period) + (1 + mo(period) | character), data = dsumm_brm, chains = nchain)
-#plot(conditional_effects(fit1))
-
-#fit2 <- brm( complexity ~ periodl + (1 + period| character), data = dsumm_brm, chains = nchain)
-
-fit3 <- brm( complexity ~ periodc + (1 + period| character), data = dsumm_brm, chains = nchain)
-
-#pme1 <- plot(conditional_effects(fit1), plot = FALSE, ask = FALSE)
-#pme2 <- plot(conditional_effects(fit2), plot = FALSE, ask = FALSE)
-pme3 <- plot(conditional_effects(fit3), plot = FALSE, ask = FALSE)
-
-#pmeall <- pme1[[1]] + pme2[[1]] + pme3[[1]]
-#loo(fit1, fit2, fit3)
-#model_weights(fit1, fit2, fit3, weights = "loo")
-
-coefnames <- fixef(fit3) %>% rownames()
-coefsamples <- fixef(fit3, summary=FALSE, pars=coefnames[-1]) %>% 
-  as_tibble() %>% 
-  rename("Bronze-Oracle"=periodcBronzeMOracle, "Seal-Bronze"=periodcSealMBronze, "Traditional-Seal"=periodcTraditionalMSeal, "Simplified-Traditional"=periodcSimplifiedMTraditional)
-postplot <- mcmc_areas(coefsamples)
-
-ggsave(here("analysis","output", "bayesreg_period.pdf"), plot = postplot,  width=4,height = 2.8)
-
-#comp2 <- loo_compare(loo(fit1), loo(fit2), loo(fit3)) %>%
-#  as.data.frame() %>%
-#  select(elpd_loo, elpd_diff, se_diff, p_loo, looic) %>%
-#  rename(
-#    "ELPD-Diff" = elpd_diff, "SE-Diff" = se_diff, 
-#    "ELPD-LOO" = elpd_loo, "P-LOO" = p_loo, "LOOIC" = looic
-#  )
-#mw2 <- model_weights(fit1, fit2, fit3, weights = "loo")[rownames(comp2)]
-#comp2 %>%
-#  cbind("Akaike-Weight" = mw2) %>%
-#  apa_table(
-#    format = "latex", booktabs = TRUE, digits = 2,
-#    caption = "Comparison of models fit4 to fit6 based on approximate leave-one-out cross-validation.",
-#    note =  "ELPD-LOO = expected log posterior predictive density (higher is better); ELPD-DIFF = difference in ELPD values compared to the best model.
-SE-DIFF = standard error of the ELPD difference.
-P-LOO = effective number of model parameters (lower is better); LOOIC: leave-one-out information criterion (lower is better); Akaike-Weight = Model weight based on the LOOIC values (higher is better).",
-#    align = c("l", rep("r", 6))
-#  )
-
-```
-
-```{r bayesianfreqdata, include=FALSE}
-dsumm_brm_freq <- dsumm_brm %>% 
-  filter(first=="Oracle") %>%   # only characters in Oracle stream
-  filter(period<"Simplified") %>%   # ignore simplified b/c singular fit when we include it
-  filter(!is.na(logfrequency)) %>% 
-  mutate(rarity= -logfrequencyscale) %>% 
-  mutate(rarityraw= -logfrequencyrawscale) %>% 
-  mutate(periodc = sdif_coding(factor(period)))
-
-suffix <- "raw"
-
-dsumm_forstats <- dsumm_brm_freq %>% 
-  mutate(rarity = !!as.symbol(paste0("rarity", suffix)))
-
-fit3_freq <- brm( complexity ~ periodc + rarity + periodc*rarity + (1 + period| character), data = dsumm_forstats, chains = nchain)
-
-coefnames <- fixef(fit3_freq) %>% rownames()
-coefsamples <- fixef(fit3_freq, summary=FALSE, pars=coefnames[-1]) %>% 
-  as_tibble() %>% 
-  rename("Bronze-Oracle"=periodcBronzeMOracle, 
-         "Seal-Bronze"=periodcSealMBronze, 
-         "Traditional-Seal"=periodcTraditionalMSeal, 
-         "(Bronze-Oracle):rarity"="periodcBronzeMOracle:rarity", 
-         "(Seal-Bronze):rarity"="periodcSealMBronze:rarity", 
-         "(Traditional-Seal):rarity"="periodcTraditionalMSeal:rarity")
-
-postplot <- mcmc_areas(coefsamples)
-
-ggsave(here("analysis","output", paste0("bayesreg_freq", suffix, ".pdf")), plot = postplot,  width=5,height = 5)
-
-```
-
-```{r bayesianpictographic, include=FALSE}
-
-fit3_ctype <- brm( complexification ~ 0+C1Type, data = d_trad_vs_ob, chains = nchain)
-
-d_trad_vs_ob_smallcomp <- d_trad_vs_ob %>%
-  filter(numcomp <= 5) %>% 
-  mutate(numcomp = factor(numcomp))
-
-fit3_numcomp <- brm( complexification ~ 0+numcomp, data = d_trad_vs_ob_smallcomp, chains = nchain)
-
-coefnames <- fixef(fit3_ctype) %>% rownames()
-coefsamples <- fixef(fit3_ctype, summary=FALSE, pars=coefnames) %>% 
-  as_tibble() %>% 
-  rename("pictographic"="C1Typepictographic",
-         "pictologic"="C1Typepictologic", 
-         "pictosyn"="C1Typepictosyn.", 
-         "pictophon"="C1Typepictophon.", 
-         "other"="C1Typeother")
-
-postplot_ctype <- mcmc_areas(coefsamples) +
-                  xlab('complexification')
-
-coefnames <- fixef(fit3_numcomp) %>% rownames()
-coefsamples <- fixef(fit3_numcomp, summary=FALSE, pars=coefnames) %>% 
-  as_tibble() %>% 
-  rename("1 comp"="numcomp1",
-         "2 comp"="numcomp2", 
-         "3 comp"="numcomp3", 
-         "4 comp"="numcomp4", 
-         "5 comp"="numcomp5")
-
-postplot_numcomp <- mcmc_areas(coefsamples) +
-                  xlab('complexification')
-
-posteriors <- postplot_ctype + postplot_numcomp + 
-    plot_annotation(tag_levels = c('a'), tag_suffix = ')')
-
-ggsave(here("analysis","output", "bayesreg_ctype.pdf"), plot = posteriors, width = 7.08661,height = 3.5)
-
-```
-
-# Supporting information
-
-## Alternative complexity measures
-
-```{r alternativecomplexity, include=TRUE}
-dpath_pixcomp <- here("data", "hanziyuan_pix_complexity.csv")
-dpath_cd<- here("data", "character_dataset.csv")
-dpath_gc<- here("data", "gif_complexities.csv")
-
-d_pixcomp <- read_csv(dpath_pixcomp) %>%
-    rename(Traditional=traditional,  Simplified=simplified) %>% 
-    pivot_longer(cols=periods, names_to = "period", values_to = "pixcomplexity") %>% 
-  select(character, period, pixcomplexity) %>% 
-  group_by(character, period) %>%
-  dplyr::summarize(pixcomplexity=mean(pixcomplexity), .groups="drop")  %>% 
-  right_join(dsumm, by = c("character", "period"))
-
-d_cd <- read_csv(dpath_cd) %>%
-  select(character, "stroke count") %>% 
-  mutate(period="Simplified") %>% 
-  right_join(dsumm, by=c("character","period")) %>% 
-  mutate(pixcomplexity = `stroke count`)
-
-d_gc <- read_csv(dpath_gc) %>%
-  mutate(period="Simplified") %>% 
-  right_join(dsumm, by=c("character","period")) %>% 
-  mutate(pixcomplexity = `gif complexity`)
-
-myscatter <- function(d) {
-  pixcomplexity <- d %>% 
-    ggscatter(x = "complexity", y = "pixcomplexity",
-    size = 0.5,
-    add = "reg.line",  
-    add.params = list(color = "blue", fill = "lightgray"), 
-    ggtheme = mytheme
-  ) +
-  stat_cor(
-    aes(label = ..r.label..),
-    label.x = 3,
-    cor.coef.name='r'
-  )  + #+ coord_fixed(ratio=2)
-  xlab('perimetric complexity')
-
-  return(pixcomplexity)
-}
-
-pixcomplexity  <- myscatter(d_pixcomp) +
-  ylab('pixel-based complexity')
-
-pixcomplexity 
-strokecount<- myscatter(d_cd) +
-  ylab('stroke count') 
-strokecount
-
-gifcomplexity<- myscatter(d_gc) +
-  ylab('writing time')
-
-altcomplex <- pixcomplexity + strokecount + gifcomplexity +
-  plot_annotation(tag_levels = c('a'), tag_suffix = ')')
-
-ggsave(here("analysis","output", "altcomplexitymeasures.pdf"), plot = altcomplex,  width=7.08661,height = 2.8)
-
-```

### 2021-02-01 docs/resources/CogSci_paper/paper.tex
fit to 6 pages

@@
 \newcommand{\ev}[2]{$#1_#2$}
+\renewcommand\bibliographytypesize{\footnotesize}
 
@@
 
-Psychological accounts of actual causation aim to characterize which of multiple causes of an event is singled out as the primary cause.
-We present one such account called the continuity account.
+Psychological theories of actual causation aim to characterize which of multiple causes of an event is singled out as the primary cause.
+We present one such theory called the continuity account of actual causation.
 The continuity account treats events as changes of state in continuous time and traces a sequence of stage changes backwards through time from an event to its primary cause.
@@
 
-Although we focus on processes unfolding in continuous time, it will be convenient to divide up the temporal dimension into intervals brief enough that there is at most one event per interval.
+Although we focus on processes unfolding in continuous time, it will be convenient to divide up the temporal dimension into intervals brief enough that at most one event occurs per interval.
 The effect to be explained is an event that happens within one of these intervals: for example, the GCM's change of state from inactive to active.
@@
 If we focus only on root causes, again there is no clear preference between \ev{R}{C} and \ev{R}{D}.
-After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM which only ``unlocks`` the detector once \ev{I}{C} is also active.
+After the activation of \ev{I}{D}, one can think of this detector as continuously sending activation towards the GCM which only ``unlocks'' the detector once \ev{I}{C} is also active.
 It follows that both \ev{R}{C} and \ev{R}{D} are connected to the effect by physical processes.
@@
 We included a random intercept for participant to allow for statistical dependencies between multiple responses from the same participant.
-The posterior mean of the intercept variable was 10.54 and the 95\% credible interval was $[2.30, 26.54]$, which supports the conclusion that root causes were chosen more often than immediate causes.
-The credible intervals for the other coefficients both included 0 (state: -2.75, $[-10.03, 1.14]$ ; long: 1.85, $[-1.78, 7.33]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.
+The posterior mean of the intercept variable was 10.8 and the 95\% credible interval was $[2.4, 26.8]$, which supports the conclusion that root causes were chosen more often than immediate causes.
+The credible intervals for the other coefficients both included 0 (state: -2.7, $[-9.2, 1.2]$ ; long: 2.1, $[-1.7, 9.1]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.
 
 The second analysis was very similar except that the binary dependent variable now indicated whether participants chose a cause on the continuous branch or the delayed branch.
-The posterior mean of the intercept variable was 2.42 (95\% credible interval $[0.33, 4.83]$), indicating that the continuous branch was chosen more often than the delayed branch.
-The credible intervals for the state variable also excluded 0 (1.39, $[0.32, 2.51]$), suggesting that state sequences led to a stronger preference for the continuous branch than did event sequences.
-Credible intervals for both length variables included 0 (medium: 0.17, $[-0.91,1.31]$ ; long: 0.11, $[-1.01,1.21]$, where the reference level is small), suggesting that the preference for the continuous branch was not affected by network length.
+The posterior mean of the intercept variable was 1.7 but the 95\% credible interval included zero ($[-0.07, 3.8]$), suggesting only weak support for the conclusion that the continuous branch was chosen more often than the delayed branch.
+The credible intervals for the state variable excluded 0 (1.0, $[0.05, 1.9]$), suggesting that state sequences led to a stronger preference for the continuous branch than did event sequences.
+Credible intervals for both length variables included 0 (medium: 0.4, $[-0.5,1.4]$ ; long: 0.5, $[-0.5,1.4]$, where the reference level is small), suggesting that the preference for the continuous branch was not strongly affected by network length.
 
@@
 \textbf{Results}.  
-
 Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.
@@
 The first analysis used a logistic regression in which order (i.e.\ the temporal order of the key events) was included as a predictor of a binary dependent variable (root or immediate).
-The posterior mean of the intercept variable was 5.75 and the 95\% credible interval $[2.44, 10.52]$, which supports the conclusion that root causes were chosen more often than immediate causes.
-The credible intervals for both order variables included 0 (\{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: 2.11, $[-1.53,1.21]$; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.38 $[-1.78,0.90]$, where the reference level was \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\}), suggesting that the preference for root causes was fairly consistent across the three orders.
+The posterior mean of the intercept variable was 5.2 and the 95\% credible interval $[2.0, 9.4]$, which supports the conclusion that root causes were chosen more often than immediate causes.
+The credible intervals for both order variables included 0 (\{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.2, $[-1.4,1.2]$; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.4 $[-1.7,0.9]$, where the reference level was \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\}), suggesting that the preference for root causes was fairly consistent across the three orders.
 
 The second and more critical analysis used a binary dependent variable that indicated whether participants chose a cause on the continuous branch or the delayed branch.
-The posterior mean of the intercept variable was 2.0 (95\% credible interval $[1.4, 2.65]$), indicating that the continuous branch was chosen more often than the delayed branch.
-The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
+The posterior mean of the intercept variable was 1.9 (95\% credible interval $[1.3, 2.5]$), indicating that the continuous branch was chosen more often than the delayed branch.
+The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.4, $[-1,0.2]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.5, $[-1,0.1]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
+
 Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2.
@@
 \bibliographystyle{apacite}
-
 \setlength{\bibleftmargin}{.125in}

### 2021-05-04 docs/resources/CogSci_paper/paper.tex
explain prob raising prediction

@@
 \usepackage{multirow}
+\usepackage{balance}
 
-%\cogscifinalcopy % Uncomment this line for the final submission
+\cogscifinalcopy % Uncomment this line for the final submission
 
@@
 \newcommand{\ev}[2]{$#1_#2$}
-\renewcommand\bibliographytypesize{\footnotesize}
+%\renewcommand\bibliographytypesize{\footnotesize}
 
@@
 
-\author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\
-  Department of Psychology, 1202 W. Johnson Street \\
-  Madison, WI 53706 USA
-  \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\
-  Department of Educational Psychology, 1025 W. Johnson Street \\
-  Madison, WI 53706 USA}
+\author{{\large \bf Aur\'{e}lien Fermo (XXX)} \\
+  XXX \\
+  XXX 
+  \AND {\large \bf Charles Kemp (c.kemp@unimelb.edu.au)} \\
+  School of Psychological Sciences, University of Melbourne\\
+  Parkville, Victoria 3010, Australia}
 
@@
 Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}.
-Psychologists and philosophers have explored these judgments in detail~and have developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.
+Psychologists and philosophers have explored these judgments in detail~and have developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg21}.
 Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.
@@
 One such sequence is shown in Figure~\ref{fig:intro}b.
-In cases like this, \ev{R}{D} guarantees the occurrence of the effect, and therefore qualifies as the primary cause according to the probability raising account.
-The continuity account, however, still treats \ev{R}{C} as the primary cause.
+In cases like this, the continuity account still treats \ev{R}{C} as the primary cause, but the probability raising account now treats \ev{R}{D} as the primary cause.
+To see why, let $b$ (for baserate) be the probability of activation along along a given branch during a given trial.
+At the start of a dual branch trial, the probability that the GCM will activate is $b^2$, because independent activations are required along both branches.
+After \ev{R}{C} the probability increases to $b$, and after \ev{R}{D} the probability increases to 1.
+The two probability increments are therefore $b - b^2$ and $1 - b$, and it is straightforward to show that $b - b^2 \le 1 - b$ for all values of $b$, with equality attained only when $b = 1$.
+As a result, the probability raising account chooses \ev{R}{D} for virtually all values of the baserate, and if $b=1$ the account is indifferent between \ev{R}{C} and \ev{R}{D}.
 
-The materials and procedure for Experiment 2 are similar to those for Experiment 1, and we highlight only the few points of difference.
+The materials and procedure for Experiment 2 are similar to those for Experiment 1, and we will highlight only the few points of difference.
+As for Experiment 1, no explicit information about the baserate $b$ was provided.
+For each network shown, activation eventually occurred along each branch, and some participants may have therefore assumed that $b=1$.
+Most participants, however, probably implicitly assumed that $b < 1$, which allows for the possibility of observation periods during which no activations occurred.
+We therefore assume that $b < 1$ when discussing the predictions of the probability raising account.
 
@@
 
+\section{Acknowledgments}
+
+AF carried out this work at the University of Melbourne while enrolled in a Master's degree at the \'{E}cole Normale Sup\'{e}rieure. 
+% add financial support?
+
 \bibliographystyle{apacite}
@@
 
+\balance
 \bibliography{uber}

### 2021-05-05 analysis/statistical_analyses.Rmd
revised entire paper

@@
          length = factor(length, levels = c("short", "medium", "long")),
-         state_or_event= factor(state_or_event, levels = c("event", "state")),
+         state_or_event = factor(state_or_event, levels = c("event", "state")),
+         position = factor(case_when(is.na(position) ~  "middle",
+                                     TRUE ~ position
+                        ), levels = c("top", "bottom", "middle")),
+         delay = factor(case_when(is.na(delay) ~  "none",
+                                     TRUE ~ as.character(delay) 
+                        ), levels = c("none", "2", "4", "6"))
         )
@@
 d1_stats1ri <- d1_stats1 %>% 
-  drop_na(ri)
+  drop_na(ri) %>% 
+  drop_na(state_or_event) # drop chain
 
 d1_stats1cd <- d1_stats1 %>% 
-  drop_na(cd)
+  drop_na(cd) %>% 
+  drop_na(state_or_event)  # drop chain
 
@@
 
-fit1_ri <- brm( ri ~ state_or_event + length + (1|participantnum), family=bernoulli(), data = d1_stats1ri, chains = nchain, iter = niter )
+fit1_ri <- brm( ri ~ state_or_event + length + (1|participantnum), family=bernoulli(), data = d1_stats1ri, chains = nchain, iter = niter, control = list(adapt_delta = 0.99) )
 

### 2021-05-05 docs/resources/CogSci_paper/paper.tex
revised entire paper

@@
 
-Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}.
-Psychologists and philosophers have explored these judgments in detail~and have developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg21}.
+Identifying the primary cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}.
+Psychologists and philosophers have explored actual causation judgments in detail~and have developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg21}.
 Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.
-On this account, the tripping of relay Q29BD is the main cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.
+On this account, the tripping of relay Q29BD is the primary cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.
+Identifying this primary cause may involve two steps: the first is to assemble a set of causes (e.g.\ a set that includes all of the state changes just mentioned), and the second is to select the primary cause from among this set~\cite{stephanmw20e}.
+Our general approach has implications for both steps, but in our setting the second step (often called causal selection) is the more challenging of the two, and we focus on it throughout.
 
@@
 The physical process view suggests that the relay tripping is a cause of the elevator stopping because the two are linked by a physical process involving the transmission of force or energy.
-Both accounts need additional machinery in order to specify which among the many causes of an event is singled out as the main cause.
+Both accounts need additional machinery in order to specify which among the many causes of an event is singled out as the primary cause.
 Our continuity account fits most naturally with the physical process approach, and can be viewed as an attempt to bring out some implications of this approach for actual causation.
@@
 State changes are embedded in continuous time, and a change that begins at a specific moment is typically the direct result of an event that occurred an instant before.
-Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause.
+Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its primary cause.
 Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09,davisbr20}.
@@
 \end{center}
+\vspace{-0.3in}
 \caption{(a)(i) A network of particle detectors.
@@
 The effect to be explained is an event that happens within one of these intervals: for example, the GCM's change of state from inactive to active.
-Given this setup, our theory can be formulated using a procedure that starts with the effect and steps backwards through time until the actual cause of the effect is identified.
+The continuity account suggests that there can be at most one event that triggered a continuous sequence of state changes that culminated in the effect, and proposes that this triggering event (if it exists) is singled out as the primary cause.
+The triggering event can be identified using a procedure that starts with the effect and steps backwards through time until the primary cause is identified.
 
@@
 The procedure terminates once we arrive at an event that has no immediate cause within the system of interest,
-and this cause is declared to be the main cause of the effect.  In Figure~\ref{fig:intro}a the main cause is the activation of \ev{R}{C}.
+and this cause is selected as the primary cause of the effect.  In Figure~\ref{fig:intro}a the primary cause is the activation of \ev{R}{C}.
 
-In most cases, the backward-tracing procedure just described will identify a single main cause of the effect.
-If the effect has no immediate cause within the system, then no main cause will be identified.
-Because there is at most one event per interval, the procedure can never identify more than one main cause.
+In most cases, the backward-tracing procedure just described will identify a single primary cause of the effect.
+If the effect has no immediate cause within the system, then no primary cause will be identified.
+Because there is at most one event per interval, the procedure can never identify more than one primary cause.
 
@@
 The boundaries of the causal system under consideration will depend on context.
-For simplicity, our discussion of Figure~\ref{fig:intro}a has focused only on events internal to the particle detection network, and has attemped to characterize which of these events is best viewed as the main cause of the GCM's activation.
+For simplicity, our discussion of Figure~\ref{fig:intro}a has focused only on events internal to the particle detection network, and has attemped to characterize which of these events is best viewed as the primary cause of the GCM's activation.
 The particle detection network, however, could also be considered part of a broader causal system that includes both the network and the network's surroundings.
-For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom could be identified as the immediate cause of \ev{R}{C} and the main cause of the GCM's activation.
+For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom could be identified as the immediate cause of \ev{R}{C} and the primary cause of the GCM's activation.
 
@@
 
+To apply the probability raising account to the dual branch stimuli, let $b$ (for base rate) be the probability of activation along along a given branch during a given trial.
+At the start of a dual branch trial, the probability that the GCM will activate is $b^2$, because independent activations are required along both branches.
+After \ev{R}{D} the probability increases to $b$, and after \ev{R}{C} the probability increases to 1.
+The two probability increments are therefore $b - b^2$ and $1 - b$, and it is straightforward to show that $b - b^2 \le 1 - b$ for all $b$, with equality attained only when $b = 1$.
+As a result, the probability raising account selects \ev{R}{C} if $b<1$, and is indifferent between \ev{R}{C} and \ev{R}{D} if $b = 1$.
+
+For each dual branch network shown, activation eventually occurred along each branch, and some participants may have therefore assumed that $b=1$.
+Most participants, however, probably implicitly assumed that $b < 1$, which allows for the possibility of trials during which a branch never activated.
+We will therefore assume that $b < 1$ when considering the predictions of the probability raising account.
+
 \textbf{Participants}. 30 participants were recruited via Amazon Mechanical Turk and paid \$3 for an 18 minute experiment.
@@
 
-Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the main cause.
+Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the primary cause.
 This result can be separated into two general conclusions.
@@
 The second conclusion is that for dual branch sequences, the root cause on the continuous branch (\ev{R}{C}) is preferred to the root cause on the delayed branch (\ev{R}{D}).
-This result is consistent with the continuity account but also consistent with the probability raising account, because \ev{R}{C} guarantees the occurrence of the effect but \ev{R}{D} does not.
+As explained earlier, this result is consistent with both the continuity and probability raising accounts.
 
-To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms} with default priors\footnote{By default brms uses improper flat priors on the coefficients of fixed effects, and a half student-t prior with 3 degrees of freedom, a location of 0 and a scale of 2.5 on the intercept and on the standard deviations of the random effects.}.
-For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded.
+To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms} with default priors.\footnote{By default brms uses improper flat priors on the coefficients of fixed effects, and a half student-t prior with 3 degrees of freedom, a location of 0 and a scale of 2.5 on the intercept and on the standard deviations of the random effects.} For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded.
 We then ran a logistic regression in which stativity (i.e.\ state vs event) and network length (long vs medium) were included as predictors of a binary dependent variable (root vs immediate).
 We included a random intercept for participant to allow for statistical dependencies between multiple responses from the same participant.
-The posterior mean of the intercept variable was 10.8 and the 95\% credible interval was $[2.4, 26.8]$, which supports the conclusion that root causes were chosen more often than immediate causes.
-The credible intervals for the other coefficients both included 0 (state: -2.7, $[-9.2, 1.2]$ ; long: 2.1, $[-1.7, 9.1]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.
+The posterior mean of the intercept variable was 11.0 and the 95\% credible interval was $[2.6, 28.7]$, which supports the conclusion that root causes were chosen more often than immediate causes.
+The credible intervals for the other coefficients both included 0 (state: -2.7, $[-9.3, 1.1]$ ; long: 2.1, $[-1.7, 9.0]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.
 
 The second analysis was very similar except that the binary dependent variable now indicated whether participants chose a cause on the continuous branch or the delayed branch.
-The posterior mean of the intercept variable was 1.7 but the 95\% credible interval included zero ($[-0.07, 3.8]$), suggesting only weak support for the conclusion that the continuous branch was chosen more often than the delayed branch.
+The posterior mean of the intercept variable was 1.7 but the 95\% credible interval included zero ($[-0.06, 3.7]$), suggesting only weak support for the conclusion that the continuous branch was chosen more often than the delayed branch.
 The credible intervals for the state variable excluded 0 (1.0, $[0.05, 1.9]$), suggesting that state sequences led to a stronger preference for the continuous branch than did event sequences.
@@
 \end{center}
+\vspace{-0.1in}
 \caption{Experiment 1.
 (a) Order of the four key events for dual branch sequences.
-The probability raising and continuity accounts both single out \ev{R}{C} as the main cause.
+The probability raising and continuity accounts both single out \ev{R}{C} as the primary cause.
 (b)  Responses for dual branch sequences.
@@
 \end{center}
+\vspace{-0.1in}
 \caption{Experiment 2.
@@
 In cases like this, the continuity account still treats \ev{R}{C} as the primary cause, but the probability raising account now treats \ev{R}{D} as the primary cause.
-To see why, let $b$ (for baserate) be the probability of activation along along a given branch during a given trial.
-At the start of a dual branch trial, the probability that the GCM will activate is $b^2$, because independent activations are required along both branches.
-After \ev{R}{C} the probability increases to $b$, and after \ev{R}{D} the probability increases to 1.
-The two probability increments are therefore $b - b^2$ and $1 - b$, and it is straightforward to show that $b - b^2 \le 1 - b$ for all values of $b$, with equality attained only when $b = 1$.
-As a result, the probability raising account chooses \ev{R}{D} for virtually all values of the baserate, and if $b=1$ the account is indifferent between \ev{R}{C} and \ev{R}{D}.
+For dual branch stimuli, the argument presented earlier shows that the probability raising account identifies the second of the two root causes (i.e.\ \ev{R}{C} in Experiment 1 and \ev{R}{D} for the sequence in Figure~\ref{fig:intro}b) as the primary cause if the base rate $b < 1$.
+If $b=1$ the account is unable to select any event as the primary cause.
 
 The materials and procedure for Experiment 2 are similar to those for Experiment 1, and we will highlight only the few points of difference.
-As for Experiment 1, no explicit information about the baserate $b$ was provided.
-For each network shown, activation eventually occurred along each branch, and some participants may have therefore assumed that $b=1$.
-Most participants, however, probably implicitly assumed that $b < 1$, which allows for the possibility of observation periods during which no activations occurred.
-We therefore assume that $b < 1$ when discussing the predictions of the probability raising account.
 
@@
 The first analysis used a logistic regression in which order (i.e.\ the temporal order of the key events) was included as a predictor of a binary dependent variable (root or immediate).
-The posterior mean of the intercept variable was 5.2 and the 95\% credible interval $[2.0, 9.4]$, which supports the conclusion that root causes were chosen more often than immediate causes.
-The credible intervals for both order variables included 0 (\{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.2, $[-1.4,1.2]$; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.4 $[-1.7,0.9]$, where the reference level was \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\}), suggesting that the preference for root causes was fairly consistent across the three orders.
+The posterior mean of the intercept variable was 5.0 and the 95\% credible interval $[2.0, 8.9]$, which supports the conclusion that root causes were chosen more often than immediate causes.
+The credible intervals for both order variables included 0 (\{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.2, $[-1.5,1.2]$; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.4 $[-1.7,0.9]$, where the reference level was \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\}), suggesting that the preference for root causes was fairly consistent across the three orders.
 
@@
 The posterior mean of the intercept variable was 1.9 (95\% credible interval $[1.3, 2.5]$), indicating that the continuous branch was chosen more often than the delayed branch.
-The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.4, $[-1,0.2]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.5, $[-1,0.1]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
+The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: -0.4, $[-0.9,0.1]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.5, $[-1,0.1]$) suggesting that the preference for the continuous branch was of similar strength across all three orders.
 
@@
 We presented an account of actual causation that highlights the role of temporal continuity and described experiments that support two of its predictions.
-First, people tend to identify root causes rather than immediate causes as the main cause of an effect.
+First, people tend to identify root causes rather than immediate causes as the primary cause of an effect.
 Second, when an effect is produced by the convergence of multiple causal pathways, participants tend to identify the pathway that has temporal continuity with the effect as the primary cause.
 
-Because the continuity account considers changes of state in continuous time, it can exploit a ``no coincidences'' principle to identify a single main cause of an effect.
+Because the continuity account considers changes of state in continuous time, it can exploit a ``no coincidences'' principle to identify a single primary cause of an effect.
 This principle applies broadly to causation in physical systems, but is less applicable to social scenarios such as voting scenarios~\cite{livengood13}.
@@
 For example, perhaps \ev{R}{C} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch.
-We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a task in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.
+We have planned a future study that asks whether the preference for \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a task in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.
 
@@
 
-AF carried out this work at the University of Melbourne while enrolled in a Master's degree at the \'{E}cole Normale Sup\'{e}rieure. 
+AF carried out this work at the University of Melbourne as part of a Master's degree at the \'{E}cole Normale Sup\'{e}rieure. 
 % add financial support?

### 2021-05-07 analysis/generating_figures.py
first commit

@@ -1,0 +1,199 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: Aurélien Fermo
+"""
+
+#%%
+import os
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+root_dir = os.path.dirname(os.getcwd())
+
+#%%
+# Importing data from experiment 1 and experiment 2.
+
+data_E1 = pd.read_csv("{}/data/E1_responses.csv".format(root_dir))
+for col in data_E1.columns[1:]:
+    data_E1[col] = data_E1[col].astype('category')
+
+data_E2 = pd.read_csv("{}/data/E2_responses.csv".format(root_dir))
+for col in data_E2.columns[1:]:
+    data_E2[col] = data_E2[col].astype('category')
+
+#%%
+# Functions for bootstraping and generating the two figures of the paper.
+
+def bootstrap(stim,experiment):
+    df_mean_std = {}
+    data = stim.response.to_numpy()
+    stim_type = stim.stimulustype.cat.remove_unused_categories().cat.categories[0]
+    if experiment == 'E1':
+        length = stim.length.cat.remove_unused_categories().cat.categories[0]
+        pop_size = 30
+        if stim_type == 'dualbranch' and length == 'short':
+            labels = ['otherC','otherD','RC_IC','RD_ID','GCM']
+        elif stim_type == 'dualbranch' and length != 'short':
+            labels = ['otherC','otherD','RC','RD','IC','ID','GCM']
+        elif stim_type == 'chain' and length == 'short':
+            labels = ['otherC','RC_IC','GCM']
+        elif stim_type == 'chain' and length != 'short':
+            labels = ['otherC','RC','IC','GCM']
+    elif experiment == 'E2':
+        pop_size = 100
+        if stim_type == 'dualbranch':
+            labels = ['RC','RD','IC','ID','otherC','otherD','GCM']
+        elif stim_type == 'chain':
+            labels = ['RC','IC','otherC','GCM']
+            
+    mean_per_cat = []
+    std_per_cat = []    
+
+    sample_mean = {}
+    for category in labels:
+            sample_mean[category] = []
+    for _ in range(100):
[…]

### 2021-05-07 experiments/Experiment D/100ms_definitive/source/js/plugins/jspsych-causal-graph1.js

@@ -1,376 +1,0 @@
-jsPsych.plugins['causal-graph1'] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'causal-graph1',
-		description: '',
-		parameters: {
-			stimulus: {
-			type: jsPsych.plugins.parameterType.IMAGE,
-			pretty_name: 'Stimulus',
-			default: undefined,
-			array: true,
-			description: 'Image to be displayed.'
-			},
-			stim_height: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus height',
-			default: 100,
-			description: 'Height of images in pixels.'
-			},
-			stim_width: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus width',
-			default: 100,
-			description: 'Width of images in pixels'
-			},
-			gif_duration: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'GIF duration',
-			default: null,
-			description: 'How long the animation lasts.'
-			},
-			status: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'The status of the plugin',
-			default: null,
-			description: 'Whether the causal graph is active or not.'
-			},
-			A_detectors_type: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Type of detectors in A',
-			default: null,
-			description: 'Whether the detectors in A are round or square.'
-			},
-			A_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Coordinates of the square detectors',
-			default: null,
-			array: true,
-			description: 'Coordinates of the round detectors.'
-			},
-			B_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Coordinates of the square detectors',
-			default: null,
-			array: true,
-			description: 'Coordinates of the square detectors.'
-			},
-			prompt: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Prompt',
-			default: null,
-			array: true,
-			description: 'It can be used to provide a reminder about the action the subject is supposed to take.'
-			},
-			button_label_again: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Button label',
-			default:  null,
[…]

### 2021-05-07 experiments/Experiment D/100ms_definitive/source/js/plugins/jspsych-causal-graph3.js

@@ -1,428 +1,0 @@
-jsPsych.plugins['causal-graph3'] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'causal-graph1',
-		description: '',
-		parameters: {
-			stim_GIF_A: {
-			type: jsPsych.plugins.parameterType.IMAGE,
-			pretty_name: 'Stimulus',
-			default: undefined,
-			description: 'Image to be displayed.'
-			},
-			stim_GIF_B: {
-			type: jsPsych.plugins.parameterType.IMAGE,
-			pretty_name: 'Stimulus',
-			default: null,
-			array: true,
-			description: 'Image to be displayed.'
-			},
-			stim_GIF_C: {
-			type: jsPsych.plugins.parameterType.IMAGE,
-			pretty_name: 'Stimulus',
-			default: undefined,
-			description: 'Image to be displayed.'
-			},
-			loop_state : {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Loop state',
-			default: null,
-			description: 'Whether the loop is in the state condition or not.'
-			},
-			stim_height: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus height',
-			default: 100,
-			description: 'Height of images in pixels.'
-			},
-			stim_width: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus width',
-			default: 100,
-			description: 'Width of images in pixels'
-			},
-			GIF_B_duration: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'GIF duration',
-			default: null,
-			array: true,
-			description: 'How long the animation lasts.'
-			},
-			status: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'The status of the plugin',
-			default: 'active',
-			description: 'Whether the causal graph is active or not.'
-			},
-			A_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Nodes coordinates',
-			default: null,
-			array: true,
-			description: 'Coordinates of the round detectors.'
-			},
-			B_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Nodes coordinates',
-			default: null,
-			array: true,
-			description: 'Coordinates of the square detectors.'
-			},
-			prompt: {
[…]

### 2021-05-07 experiments/Experiment1/500ms_temporary/source/js/plugins/jspsych-causal-graph1.js

@@ -1,315 +1,0 @@
-jsPsych.plugins['causal-graph1'] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'causal-graph1',
-		description: '',
-		parameters: {
-			stimulus: {
-			type: jsPsych.plugins.parameterType.IMAGE,
-			pretty_name: 'Stimulus',
-			default: undefined,
-			array: true,
-			description: 'Image to be displayed.'
-			},
-			stim_height: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus height',
-			default: 100,
-			description: 'Height of images in pixels.'
-			},
-			stim_width: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus width',
-			default: 100,
-			description: 'Width of images in pixels'
-			},
-			gif_duration: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'GIF duration',
-			default: null,
-			description: 'How long the animation lasts.'
-			},
-			status: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'The status of the plugin',
-			default: 'active',
-			description: 'Whether the causal graph is active or not.'
-			},
-			nodes_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Nodes coordinates',
-			default: null,
-			array: true,
-			description: 'Coordinates of the top-left of each node'
-			},
-			prompt: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Prompt',
-			default: null,
-			array: true,
-			description: 'It can be used to provide a reminder about the action the subject is supposed to take.'
-			},
-			button_label_again: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Button label',
-			default:  null,
-			description: 'The text that appears on the button to run again the current trial.'
-			},
-			button_label_next: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Button label',
-			default:  ['Run','Next network'],
-			array: true,
-			description: 'The text that appears on the button to continue to the next trial.'
-			},
-			prompt_error: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Prompt error',
-			default: null,
[…]

### 2021-05-07 experiments/Experiment2/js/plugins/jspsych-causal-graph1.js

@@ -1,0 +1,376 @@
+jsPsych.plugins['causal-graph1'] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'causal-graph1',
+		description: '',
+		parameters: {
+			stimulus: {
+			type: jsPsych.plugins.parameterType.IMAGE,
+			pretty_name: 'Stimulus',
+			default: undefined,
+			array: true,
+			description: 'Image to be displayed.'
+			},
+			stim_height: {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'Stimulus height',
+			default: 100,
+			description: 'Height of images in pixels.'
+			},
+			stim_width: {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'Stimulus width',
+			default: 100,
+			description: 'Width of images in pixels'
+			},
+			gif_duration: {
+			type: jsPsych.plugins.parameterType.INT,
+			pretty_name: 'GIF duration',
+			default: null,
+			description: 'How long the animation lasts.'
+			},
+			status: {
+			type: jsPsych.plugins.parameterType.STRING,
+			pretty_name: 'The status of the plugin',
+			default: null,
+			description: 'Whether the causal graph is active or not.'
+			},
+			A_detectors_type: {
+			type: jsPsych.plugins.parameterType.STRING,
+			pretty_name: 'Type of detectors in A',
+			default: null,
+			description: 'Whether the detectors in A are round or square.'
+			},
+			A_coord: {
+			type: jsPsych.plugins.parameterType.OBJECT,
+			pretty_name: 'Coordinates of the square detectors',
+			default: null,
+			array: true,
+			description: 'Coordinates of the round detectors.'
+			},
+			B_coord: {
+			type: jsPsych.plugins.parameterType.OBJECT,
+			pretty_name: 'Coordinates of the square detectors',
+			default: null,
+			array: true,
+			description: 'Coordinates of the square detectors.'
+			},
+			prompt: {
+			type: jsPsych.plugins.parameterType.STRING,
+			pretty_name: 'Prompt',
+			default: null,
+			array: true,
+			description: 'It can be used to provide a reminder about the action the subject is supposed to take.'
+			},
+			button_label_again: {
+			type: jsPsych.plugins.parameterType.STRING,
+			pretty_name: 'Button label',
+			default:  null,
[…]

### 2021-05-07 experiments/ExperimentB/500ms_temporary/source/backend.py

@@ -1,39 +1,0 @@
-import os
-import urllib
-import jinja2
-import webapp2
-
-from google.appengine.api import users
-from google.appengine.ext import ndb
-
-JINJA_ENVIRONMENT = jinja2.Environment(
-    loader=jinja2.FileSystemLoader(os.path.dirname(__file__)),
-    extensions=['jinja2.ext.autoescape'])
-
-
-class DataObject(ndb.Model):
-    """Models an individual Guestbook entry with author, content, and date."""
-    content = ndb.TextProperty(required=True) # defaults to non-indexed
-    date = ndb.DateTimeProperty(auto_now_add=True, indexed=False)
-
-
-class MainPage(webapp2.RequestHandler):
-
-    def get(self):
-        template = JINJA_ENVIRONMENT.get_template('index.html')
-        self.response.write(template.render())
-
-class WriteDataObject(webapp2.RequestHandler):
-
-    def post(self):
-        data = DataObject()
-
-        data.content = self.request.get('content')
-        data.exp = self.request.get('exp')
-        data.put()
-
-
-application = webapp2.WSGIApplication([
-    ('/', MainPage),
-    ('/submit', WriteDataObject),
-], debug=True)

### 2021-05-07 experiments/ExperimentB/500ms_temporary/source/js/plugins/jspsych-causal-familiarisation.js

@@ -1,186 +1,0 @@
-jsPsych.plugins['causal-familiarisation'] = (function() {
-
-  var plugin = {};
-
-  plugin.info = {
-    name: 'causal-familiarisation',
-    description: '',
-    parameters: {
-      stimulus: {
-        type: jsPsych.plugins.parameterType.IMAGE,
-        pretty_name: 'Stimulus',
-        default: undefined,
-        description: 'Image to be displayed.'
-      },
-      stim_height: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Stimulus height',
-        default: 100,
-        description: 'Height of images in pixels.'
-      },
-      stim_width: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Stimulus width',
-        default: 100,
-        description: 'Width of images in pixels'
-      },
-      nodes_coord: {
-        type: jsPsych.plugins.parameterType.OBJECT,
-        pretty_name: 'Nodes coordinates',
-        default: null,
-        array: true,
-        description: 'Coordinates of the top-left of each node'
-      },
-      prompt: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt',
-        default: null,
-        array: true,
-        description: 'It can be used to provide a reminder about the action the subject is supposed to take.'
-      },
-      button_label_next: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button label',
-        default:  ['Continue'],
-        description: 'The text that appears on the button to continue to the next trial.'
-      }
-    }
-  };
-
-  plugin.trial = function(display_element, trial) {
-    
-    var start_time = performance.now();
-    
-  	var IMG = $("<img>",{
-      id: "Graphics",
-      src: trial.stimulus,
-      useMap:"#gra"
-	  });
-	  IMG.addClass("center");
-
-	  var MAP1 = $("<map>",{
-  	  name: "gra"
-	  });
-
-    var html_IMG = '<div id="IMG_container"</div>';
-    var next_btn = '<button id="jspsych-causal-familiarisation-next-btn" class="jspsych-btn" style="margin-left: 5px">'+trial.button_label_next+'</button>';
-
-    display_element.innerHTML = html_IMG;
-    IMG.appendTo("#IMG_container");
-    MAP1.appendTo("#IMG_container");
[…]

### 2021-05-07 experiments/ExperimentB/500ms_temporary/source/js/plugins/jspsych-causal-graph2.js

@@ -1,362 +1,0 @@
-jsPsych.plugins['causal-graph2'] = (function() {
-
-  var plugin = {};
-
-  plugin.info = {
-    name: 'causal-graph2',
-    description: '',
-    parameters: {
-      stimulus: {
-        type: jsPsych.plugins.parameterType.IMAGE,
-        pretty_name: 'Stimulus',
-        default: undefined,
-        array: true,
-        description: 'Image to be displayed.'
-      },
-      stim_height: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Stimulus height',
-        default: 100,
-        description: 'Height of images in pixels.'
-      },
-      stim_width: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Stimulus width',
-        default: 100,
-        description: 'Width of images in pixels'
-      },
-      gif_duration: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'GIF duration',
-        default: null,
-        description: 'How long the animation lasts.'
-      },
-      status: {
-      	type: jsPsych.plugins.parameterType.STRING,
-      	pretty_name: 'The status of the plugin',
-      	default: 'active',
-      	description: 'Whether the causal graph is active or not.'
-      },
-      nodes_coord: {
-        type: jsPsych.plugins.parameterType.OBJECT,
-        pretty_name: 'Nodes coordinates',
-        default: null,
-        array: true,
-        description: 'Coordinates of the top-left of each node'
-      },
-      prompt: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt',
-        default: null,
-        array: true,
-        description: 'It can be used to provide a reminder about the action the subject is supposed to take.'
-      },
-      button_label_again: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button label',
-        default:  null,
-        description: 'The text that appears on the button to run again the current trial.'
-      },
-      button_label_next: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button label',
-        default:  ['Run','Next network'],
-        array: true,
-        description: 'The text that appears on the button to continue to the next trial.'
-      },
-      prompt_error: {
-      	type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt error',
-        default: null,
-        description: 'The message displayed in case of error.'
[…]

### 2021-05-07 experiments/ExperimentC/100ms_definitive/source/js/plugins/jspsych-causal-graph1.js

@@ -1,376 +1,0 @@
-jsPsych.plugins['causal-graph1'] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'causal-graph1',
-		description: '',
-		parameters: {
-			stimulus: {
-			type: jsPsych.plugins.parameterType.IMAGE,
-			pretty_name: 'Stimulus',
-			default: undefined,
-			array: true,
-			description: 'Image to be displayed.'
-			},
-			stim_height: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus height',
-			default: 100,
-			description: 'Height of images in pixels.'
-			},
-			stim_width: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'Stimulus width',
-			default: 100,
-			description: 'Width of images in pixels'
-			},
-			gif_duration: {
-			type: jsPsych.plugins.parameterType.INT,
-			pretty_name: 'GIF duration',
-			default: null,
-			description: 'How long the animation lasts.'
-			},
-			status: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'The status of the plugin',
-			default: null,
-			description: 'Whether the causal graph is active or not.'
-			},
-			A_detectors_type: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Type of detectors in A',
-			default: null,
-			description: 'Whether the detectors in A are round or square.'
-			},
-			A_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Coordinates of the square detectors',
-			default: null,
-			array: true,
-			description: 'Coordinates of the round detectors.'
-			},
-			B_coord: {
-			type: jsPsych.plugins.parameterType.OBJECT,
-			pretty_name: 'Coordinates of the square detectors',
-			default: null,
-			array: true,
-			description: 'Coordinates of the square detectors.'
-			},
-			prompt: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Prompt',
-			default: null,
-			array: true,
-			description: 'It can be used to provide a reminder about the action the subject is supposed to take.'
-			},
-			button_label_again: {
-			type: jsPsych.plugins.parameterType.STRING,
-			pretty_name: 'Button label',
-			default:  null,
[…]

### 2021-05-07 experiments/ExperimentC/100ms_definitive/source/js/plugins/jspsych-survey-multi-choice.js

@@ -1,198 +1,0 @@
-/**
- * jspsych-survey-multi-choice
- * a jspsych plugin for multiple choice survey questions
- *
- * Shane Martin
- *
- * documentation: docs.jspsych.org
- *
- */
-
-
-jsPsych.plugins['survey-multi-choice'] = (function() {
-  var plugin = {};
-
-  plugin.info = {
-    name: 'survey-multi-choice',
-    description: '',
-    parameters: {
-      questions: {
-        type: jsPsych.plugins.parameterType.COMPLEX,
-        array: true,
-        pretty_name: 'Questions',
-        nested: {
-          prompt: {
-            type: jsPsych.plugins.parameterType.STRING,
-            pretty_name: 'Prompt',
-            default: undefined,
-            description: 'The strings that will be associated with a group of options.'
-          },
-          options: {
-            type: jsPsych.plugins.parameterType.STRING,
-            pretty_name: 'Options',
-            array: true,
-            default: undefined,
-            description: 'Displays options for an individual question.'
-          },
-          required: {
-            type: jsPsych.plugins.parameterType.BOOL,
-            pretty_name: 'Required',
-            default: false,
-            description: 'Subject will be required to pick an option for each question.'
-          },
-          horizontal: {
-            type: jsPsych.plugins.parameterType.BOOL,
-            pretty_name: 'Horizontal',
-            default: false,
-            description: 'If true, then questions are centered and options are displayed horizontally.'
-          },
-          name: {
-            type: jsPsych.plugins.parameterType.STRING,
-            pretty_name: 'Question Name',
-            default: '',
-            description: 'Controls the name of data values associated with this question'
-          }
-        }
-      },
-      randomize_question_order: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Randomize Question Order',
-        default: false,
-        description: 'If true, the order of the questions will be randomized'
-      },
-      preamble: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Preamble',
-        default: null,
-        description: 'HTML formatted string to display at the top of the page above all the questions.'
-      },
-      button_label: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button label',
-        default:  'Continue',
-        description: 'Label of the button.'
-      }
-    }
-  }
[…]

### 2021-05-07 analysis/generating_figures.py
last update

@@ -13,14 +13,15 @@ root_dir = os.path.dirname(os.getcwd())
 
 #%%
 # Importing data from experiment 1 and experiment 2.
-
-data_E1 = pd.read_csv("{}/data/E1_responses.csv".format(root_dir))
+data_E1 = pd.read_csv(os.path.join(root_dir,'data','E1_responses.csv'))
 for col in data_E1.columns[1:]:
     data_E1[col] = data_E1[col].astype('category')
+print('\n Dataframe of the results of Experiment 1:\n\n',data_E1)
 
-data_E2 = pd.read_csv("{}/data/E2_responses.csv".format(root_dir))
+data_E2 = pd.read_csv(os.path.join(root_dir,'data','E2_responses.csv'))
 for col in data_E2.columns[1:]:
     data_E2[col] = data_E2[col].astype('category')
+print('\n Dataframe of the results of Experiment 2:\n\n',data_E2)
 
 #%%
 # Functions for bootstraping and generating the two figures of the paper.
@@ -165,6 +166,7 @@ def graph_E2(x):
                           index=categories)   
     df_mean.plot.bar(ax=axis,yerr=df_std,capsize=1,ecolor='black',rot=0,error_kw={'elinewidth':0.5,'capthick':0.5},fontsize=7)
     axis.legend(fontsize=7)
+    plt.show()
 
 #%%
 ShortState = data_E1[(data_E1['stimulustype']=='dualbranch')&(data_E1['length']=='short')&(data_E1['state_or_event']=='state')]

### 2021-05-11 analysis/statistical_analyses.Rmd
first commit
update analysis script

@@
 library(ggthemes)
-library(lme4)
 library(brms)
@@
 ```{r data, include=TRUE}
-dpath_exp1 <- here("data", "responsesOnly_E1_100def.csv")
-dpath_exp2 <- here("data", "responsesOnly_E2_100def.csv")
+dpath_exp1 <- here("data", "E1_responses.csv")
+dpath_exp2 <- here("data", "E2_responses.csv")
 
@@
 
+# For simplicity run analyses on dual networks only
+
 d1_stats1ri <- d1_stats1 %>% 
@@
 
-# use sum coding 
-#options(contrasts=c("contr.sum", "contr.poly"))
 options(contrasts=c("contr.treatment", "contr.poly"))
 
-m0 <- glmer(ri ~ state_or_event + length + (1|participantnum), data = d1_stats1, 
-            family = binomial(link = "logit"))
-
-nchain <- 1 # for testing
-niter <- 500 # for testing
-
-nchain <- 8 # for real run
-niter <- 4000 # for real run
+nchain <- 8 
+niter <- 4000
 
@@
 
-# use sum coding 
-#options(contrasts=c("contr.sum", "contr.poly"))
 options(contrasts=c("contr.treatment", "contr.poly"))
@@
 
-nchain <- 1 # for testing
-niter <- 500 # for testing
-
-nchain <- 4 # for real run
-niter <- 2000 # for real run
+nchain <- 8 
+niter <- 4000
 

### 2021-05-12 analysis/statistical_analyses.Rmd
made pass through statistical_analyses notebook

@@
 ```
+
+This notebook uses `brms` to analyze data from the two experiments reported in Fermo, A. & Kemp, C. (2021). Temporal Continuity and the Judgment of Actual Causation.
 
@@
 # Experiment 1 data frame 
-d1 <- read_csv(dpath_exp1) #%>%
-d2 <- read_csv(dpath_exp2) #%>%
+d1 <- read_csv(dpath_exp1) 
+d2 <- read_csv(dpath_exp2)
 
@@
 
-### Plot data
+### Plot Experiment 1 Data
 
@@
 
-### Stats for Expt 1
+### Statistical analyses for Expt 1
 
@@
 
-# For simplicity run analyses on dual networks only
+# the state_or_event variable doesn't apply to chains, so for simplicity we run analyses on dual networks only
 
@@
 
-### Fig 2
+### Plot Experiment 2 data
+
 ```{r exp2plot, include=TRUE, fig.height=4, fig.width=6}
